---
title: "Wrangling FishMIP Fishing Effort"
author: "Kieran Murphy"
date: "Sys.Date()"
output: rmdformats::readthedown
---

Taking fishing effort and catch from FishMIP output located at:

http://portal.sf.utas.edu.au/thredds/catalog/gem/fishmip/catalog.html

Follow InputData/ -> fishing/ -> histsoc/ then use the regional files filter to the Prydz Bay regional model domain

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Packages


```{r install-packages, include = FALSE}
# source('./R/install_required_packages.R') #run this line to install the required packages
```

```{r load-packages, warning=FALSE, message=FALSE}
# library(palmerpenguins) #penguins data
library(tidyverse) #wrangling
# library(tidymodels) #modeling and machine learning
# library(vip) #variable importance
# library(fastshap) #SHapley Additive exPlanations
library(janitor) #cleaning variable names
library(lubridate) #manipulating dates
library(tictoc) #timing computations
# library(ggcorrplot) #correlation plots
```

## The FishMIP data

Load in and inspect the master dataframes for historical catch and effort

```{r FishMIP_hist_data, echo=TRUE}
df_FishMIP_catch <- read.csv("FishMIP_fishing_data/calibration_catch_histsoc_1850_2004_regional_models.csv")
df_FishMIP_effort <- read.csv("FishMIP_fishing_data/effort_histsoc_1841_2010_regional_models.csv")
```

And let's take a glimpse at the two datasets:

```{r glimpse-FishMIP_data}
dplyr::glimpse(df_FishMIP_catch)
dplyr::glimpse(df_FishMIP_effort)
```
Check how many regions are included and identify the region of interest, Prydz Bay
```{r unique-region}
unique(df_FishMIP_catch$region)
```
Great, there it is. It's called `Prydz.Bay`


Filter dataframe so now I have the Prydz Bay data only

```{r filter-catch}
df_Prydz_catch <- df_FishMIP_catch %>% 
  dplyr::filter(region == "Prydz.Bay") %>% 
  dplyr::group_by(Year,Sector,SAUP, FGroup) %>% 
  dplyr::mutate(total_catch = sum(Reported,IUU,Discards),
                FishMIP_yield = sum(Reported,Discards))

head(df_Prydz_catch)
```

```{r}
# saveRDS(df_Prydz_catch, "FishMIP_fishing_data/calibration_catch_histsoc_1850_2004_Prydz_Bay.csv")
```


```{r filter-effort}
df_Prydz_effort <- df_FishMIP_effort %>% 
  dplyr::filter(region == "Prydz.Bay")

head(df_Prydz_effort)
dplyr::glimpse(df_Prydz_effort)
summary(df_Prydz_effort)
```

```{r}
# saveRDS(df_Prydz_effort, "FishMIP_fishing_data/effort_histsoc_1841_2010_Prydz_Bay.csv")
```



Time series of catch by functional group
```{r}
df_Prydz_catch %>% 
  ggplot(aes(x = Year, y = total_catch, colour = FGroup)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(~FGroup)
```

Time series of effort by functional group
```{r}
df_Prydz_effort %>% 
  ggplot(aes(x = Year, y = NomActive, colour = FGroup)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(~FGroup)
```

Aggregate `FGroup` so that they align with functional groups in the Prydz Bay size spectrum model

```{r unique_FGroup}
unique(df_Prydz_catch$FGroup)
```


```{r}
params <- readRDS("params/params_latest_xx.RDS")
params@species_params$species
params@species_params$w_max
```


"demersal<30cm" = "shelf and coastal fishes"

"rays<90cm" = NA - not represented in the current functional groups

"benthopelagic30-90cm" = "shelf and coastal fishes"

"benthopelagic>=90cm" = "toothfishes"

"krill" = "antarctic krill"

"pelagic30-90cm" = "shelf and coastal fishes"

"bathydemersal>=90cm" = "toothfishes"

"pelagic<30cm" = "shelf and coastal fishes": When I extracted data at the Marine Ecosystem (ME) level on sea around us for East Antarctic Dronning Maud Land, East Antarctic Enderby Land, and East Antarctic Wilkes Land (which describe the model region better than using the RFMO selection of CCAMLR), all the reported catch fit into the "shelf and coastal fishes" group. At the CCAMLR level, there are significant lanternfish catches, i.e., "mesopelagic fishes"

"lobsterscrab" = NA - not represented in the current functional groups

"cephalopods" = "squids"

"demersal30-90cm" = "shelf and coastal fishes"

"bathypelagic<30cm" = "bathypelagic fishes"

"bathydemersal30-90cm" = "shelf and coastal fishes"

```{r tidy-penguins}
df_Prydz_catch_prepared <- df_Prydz_catch %>%
  # janitor::clean_names() %>%
  # dplyr::select(c(species, island, date_egg, culmen_length_mm:delta_13_c_o_oo)) %>%
  dplyr::mutate(species = case_when(FGroup == "demersal<30cm" ~ "shelf and coastal fishes",
                             FGroup == "rays<90cm" ~ "rays<90cm",
                             FGroup == "benthopelagic30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "benthopelagic>=90cm" ~ "toothfishes",
                             FGroup == "krill" ~ "antarctic krill",
                             FGroup == "pelagic30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "bathydemersal>=90cm" ~ "toothfishes",
                             FGroup == "pelagic<30cm" ~ "shelf and coastal fishes",
                             FGroup == "lobsterscrab" ~ "lobsterscrab",
                             FGroup == "cephalopods" ~ "squids",
                             FGroup == "demersal30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "bathypelagic<30cm" ~ "bathypelagic fishes",
                             FGroup == "bathydemersal30-90cm" ~ "shelf and coastal fishes"
                             )) %>% 
  filter(!FGroup=="rays<90cm" & !FGroup=="lobsterscrab") %>% 
  group_by(Year, species) %>% 
  # summarise(catch = sum(FishMIP_yield), # not including IUU value
  summarise(catch = sum(total_catch),
            catch_g = catch*1e6)

            # total_catch = sum(total_catch),
            # reported = sum(Reported),
            # discards = sum(Discards),
            # IUU = sum(IUU))
  # dplyr::mutate(species = factor(species),
  #               island = factor(island),
  #               sex = factor(sex),
  #               year = lubridate::year(date_egg)) %>%
  # dplyr::rename(bill_length_mm = culmen_length_mm,
  #               bill_depth_mm = culmen_depth_mm,
  #               delta_15_n = delta_15_n_o_oo,
  #               delta_13_c = delta_13_c_o_oo) %>%
  # tidyr::drop_na()

glimpse(df_Prydz_catch_prepared)
head(df_Prydz_catch_prepared)
summary(df_Prydz_catch_prepared)

# saveRDS(df_Prydz_catch_prepared, "FishMIP_fishing_data/catch_Prydz_Bay_1950_2004_resolution_v1.RDS") # total catch, all 3 categories
# saveRDS(df_Prydz_catch_prepared, "FishMIP_fishing_data/catch_Prydz_Bay_1950_2004_resolution_v2.RDS") # just reported and discards, as IUU catch possibly not accounted for in the effort * Need to confirm this
```

```{r}
# df_Prydz_catch_prepared <- readRDS("FishMIP_fishing_data/catch_Prydz_Bay_1950_2004_resolution_v1.RDS")
# Load required packages
library(tidyr)
library(dplyr)

df_catch <- df_Prydz_catch_prepared[,-3]

# Assume your original dataframe is named summary_df

# Pivot the dataframe to create the desired array format
reshaped_df <- df_catch %>%
  pivot_wider(names_from = species, values_from = catch_g, values_fill = 0)

# Rename columns for cleaner names
# colnames(reshaped_df)[-1] <- gsub(" ", "_", colnames(reshaped_df)[-1]) # Replace spaces with underscores
# colnames(reshaped_df)[-1] <- gsub("[^[:alnum:]_]", "", colnames(reshaped_df)[-1]) # Remove non-alphanumeric characters

# Display the reshaped dataframe
print(reshaped_df)
```

# Read in IWC catch

```{r}
df_IWC_catch <- readRDS("IWC_data/catch_timeseries_BanzareBank_1930_2019_CPUE.rds")

glimpse(df_IWC_catch)
```

```{r}
unique(df_IWC_catch$Species)
```
I need to add some code cleaning notes here:

```{r}
df_IWC_catch_prepared <- df_IWC_catch %>%
  # janitor::clean_names() %>%
  dplyr::select(c(Year, Species, total_catch_kg)) %>%
  dplyr::mutate(species = case_when(Species == "Baleen" ~ "baleen whales",
                             Species == "Sperm" ~ "sperm whales",
                             Species == "Antarctic Minke" ~ "minke whales",
                             Species == "Killer" ~ "orca")) %>% 
  dplyr::select(c(Year, species, total_catch_kg)) %>%
  mutate(catch_g = total_catch_kg * 1000) %>% 
  dplyr::select(c(Year, species, catch_g))

head(df_IWC_catch_prepared)
```

```{r}
df_IWC_catch_prepared
df_FishMIP_catch_prepared <- df_Prydz_catch_prepared[,-3]

head(df_IWC_catch_prepared)
head(df_FishMIP_catch_prepared)
```

```{r}
df_combined <- bind_rows(df_IWC_catch_prepared, df_FishMIP_catch_prepared)
```


```{r}

# Assume your original dataframe is named summary_df

# Pivot the dataframe to create the desired array format
reshaped_df <- df_combined %>%
  pivot_wider(names_from = species, values_from = catch_g, values_fill = 0)

# Rename columns for cleaner names
# colnames(reshaped_df)[-1] <- gsub(" ", "_", colnames(reshaped_df)[-1]) # Replace spaces with underscores
# colnames(reshaped_df)[-1] <- gsub("[^[:alnum:]_]", "", colnames(reshaped_df)[-1]) # Remove non-alphanumeric characters

# Display the reshaped dataframe
print(reshaped_df)
```


```{r}
# Order the combined dataframe by year
ordered_df <- reshaped_df %>%
  arrange(Year)

ordered_df
```


```{r}
# List of extra species names
extra_species <- c("mesozooplankton", "other krill", "other macrozooplankton", "salps",
                   "mesopelagic fishes", "flying birds", "small divers", "leopard seals",
                   "medium divers", "large divers")

# Loop through the extra species and add them with 0 values
for (species in extra_species) {
  ordered_df[[species]] <- 0
}

# Print the extended dataframe
print(ordered_df)
```

```{r}
# List of desired species order
desired_species_order <- params@species_params$species
# desired_species_order <- c("orca", "sperm whales", "minke whales", "baleen whales", "toothfishes", "shelf and coastal fishes", "new_species1", "new_species2", "new_species3")

# Reorder the columns based on the desired order
reordered_df <- ordered_df %>%
  dplyr::select(Year, all_of(desired_species_order))

# Print the reordered dataframe
print(reordered_df)
summary(reordered_df)
```


```{r}
# Create a sequence of years covering the entire range
all_years <- seq(min(reordered_df$Year), max(reordered_df$Year), by = 1)


# Create a new dataframe with the filled years
filled_df <- data.frame(Year = all_years)

# Merge the filled data with the original dataframe
merged_df <- merge(filled_df, reordered_df, by = "Year", all.x = TRUE)

# write.csv(merged_df, "catch_timeseries.csv")
```



```{r}
# Create a sequence of years from 1930 to 2019
all_years <- seq(1930, 2019)

existing_years <- ordered_df$Year

# Find the missing years
missing_years <- setdiff(all_years, existing_years)

# Print the missing years
print(missing_years)
```




```{r}

# Assume your original dataframe is named summary_df

# Pivot the dataframe to create the desired array format
reshaped_df_IWC <- df_IWC_catch_prepared %>%
  pivot_wider(names_from = species, values_from = catch_g, values_fill = 0)

# Rename columns for cleaner names
# colnames(reshaped_df)[-1] <- gsub(" ", "_", colnames(reshaped_df)[-1]) # Replace spaces with underscores
# colnames(reshaped_df)[-1] <- gsub("[^[:alnum:]_]", "", colnames(reshaped_df)[-1]) # Remove non-alphanumeric characters

# Display the reshaped dataframe
print(reshaped_df_IWC)
```


```{r}
params@species_params$species

# Replace NA values with 0
merged_df_clean <- merged_df %>%
  replace_na(list("mesozooplankton" = 0,
                  "other krill" = 0,
                  "other macrozooplankton" = 0,
                  'antarctic krill' = 0,
                  "salps" = 0,
                  "mesopelagic fishes" = 0,
                  'bathypelagic fishes' = 0,
                  'shelf and coastal fishes' = 0,
                  "flying birds" = 0,
                  "small divers" = 0,
                  "squids" = 0,
                  "toothfishes" = 0,
                  "leopard seals" = 0,
                  "medium divers" = 0,
                  "large divers" = 0,
                  'minke whales' = 0,
                  "orca" = 0,
                  'sperm whales' = 0,
                  'baleen whales' = 0))


merged_df_clean

# write.csv(merged_df_clean, "yield_observed_timeseries.csv")
```




```{r}
# Combine the dataframes while keeping the correct order
combined_df <- bind_rows(reshaped_df, reshaped_df_IWC)

# Replace NA values with 0
combined_df <- combined_df %>%
  replace_na(list('shelf and coastal fishes' = 0,
                  toothfishes = 0,
                  'antarctic krill' = 0,
                  squids = 0,
                  'bathypelagic fishes' = 0,
                  'baleen whales' = 0,
                  'sperm whales' = 0,
                  'minke whales' = 0,
                  orca = 0))


# Group by Year and summarize to merge values for all species
merged_df <- combined_df %>%
  group_by(Year) %>%
  summarize(across('shelf and coastal fishes':orca, sum), .groups = "keep")

# Replace NA values with 0
merged_df <- merged_df %>%
  replace_na(list('shelf and coastal fishes' = 0,
                  toothfishes = 0,
                  'antarctic krill' = 0,
                  squids = 0,
                  'bathypelagic fishes' = 0,
                  'baleen whales' = 0,
                  'sperm whales' = 0,
                  'minke whales' = 0,
                  orca = 0))


ordered_df <- combined_df %>%
  arrange(Year)

head(ordered_df)
```




Time series of catch by functional group
```{r}
df_Prydz_catch_prepared %>% 
  ggplot(aes(x = Year, y = catch, colour = species)) +
  geom_point() +
  scale_y_log10() +
  facet_wrap(~species)
```

The other option is to remove effort by functional group and just use effort  gear types

Select the main gear types and filter the region effort based on these
"Trawl_Bottom" = ice fish (shelf & coastal fish + toothfish) each gear can have more than one species (see gearParams())
Midwater/pelagic trawl = krill
Longline = toothfish

```{r unique_Gear}
unique(df_Prydz_effort$Gear)
```
The gears don't seem to match up well with what should actually be catching what, i.e., "Lines_Handlines_and_poles", "Lines_Longlines", "Pots_and_Traps" etc catching krill 


```{r tidy-penguins}
df_Prydz_effort_prepared <- df_Prydz_effort %>%
  # janitor::clean_names() %>%
  # dplyr::select(c(species, island, date_egg, culmen_length_mm:delta_13_c_o_oo)) %>%
  dplyr::mutate(species = case_when(FGroup == "demersal<30cm" ~ "shelf and coastal fishes",
                             FGroup == "rays<90cm" ~ "rays<90cm",
                             FGroup == "benthopelagic30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "benthopelagic>=90cm" ~ "toothfishes",
                             FGroup == "krill" ~ "antarctic krill",
                             FGroup == "pelagic30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "bathydemersal>=90cm" ~ "toothfishes",
                             FGroup == "pelagic<30cm" ~ "shelf and coastal fishes",
                             FGroup == "lobsterscrab" ~ "lobsterscrab",
                             FGroup == "cephalopods" ~ "squids",
                             FGroup == "demersal30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "bathypelagic<30cm" ~ "bathypelagic fishes",
                             FGroup == "bathydemersal30-90cm" ~ "shelf and coastal fishes"
                             )) %>% 
  filter(!FGroup=="rays<90cm" & !FGroup=="lobsterscrab") %>% 
  group_by(Year, species) %>% 
  summarise(effort = sum(NomActive)) %>% 
  filter(!effort < 1e-15) %>%
  group_by(species) %>% 
  mutate(max_effort = max(effort)) %>%
  ungroup() %>% 
  group_by(species,Year) %>%
  mutate(effort_standard = effort/max_effort) %>% 
  ungroup() %>% 
  # dplyr::mutate(species = factor(species),
  #               island = factor(island),
  #               sex = factor(sex),
  #               year = lubridate::year(date_egg)) %>%
  # dplyr::rename(bill_length_mm = culmen_length_mm,
  #               bill_depth_mm = culmen_depth_mm,
  #               delta_15_n = delta_15_n_o_oo,
  #               delta_13_c = delta_13_c_o_oo) %>%
  tidyr::drop_na()

glimpse(df_Prydz_effort_prepared)
head(df_Prydz_effort_prepared)
summary(df_Prydz_effort_prepared)

# saveRDS(df_Prydz_effort_prepared, "FishMIP_fishing_data/effort_Prydz_Bay_1950_2010_resolution_v1.RDS")
```

Time series of effort by functional group
```{r}
df_Prydz_effort_prepared %>% 
  ggplot(aes(x = Year, y = effort, colour = species)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(~species)
```


Time series of effort by functional group
```{r}
p_effort <- df_Prydz_effort_prepared %>% 
  ggplot(aes(x = Year, y = effort_standard, colour = species)) +
  geom_point() +
  geom_line() +
  # scale_y_log10() +
  ylab("Normalised Effort") +
  facet_wrap(~species) +
  theme_classic()

# ggsave("plots/FishMIP_Effort_Normalised_Prydz_Bay_Annual_Means.tiff", plot = p_effort, units="cm", width=24, height=14, dpi =300)
```

