---
title: "Prydz Bay therMizer"
output: html_notebook
---

```{r}
# remotes::install_github("sizespectrum/mizerHowTo")
# remotes::install_github("sizespectrum/mizerExperimental")
# remotes::install_github("sizespectrum/therMizer")
```


```{r}
library(mizer)
library(therMizer)
library(tidyverse)
library(mizerExperimental)
#library(mizerHowTo)
library(pbapply)
library(openair)
```

```{r}
params <- readRDS("params_for_use.RDS")
```

Check initial effort

```{r}
initial_effort(params)
```

Inspect some key parameters
```{r}
species_params(params) |> select(erepro, R_max)
```

# Update yield_observed

Make sure to use yield from 2001-2010

```{r}
species_params(params) |> select(species, yield_observed)
```


Inspect ppmr and feeding kernel parameters
```{r}
species_params(params) |> select(pred_kernel_type, beta, ppmr_min, ppmr_max)
```

Update with latest depth resolved temperature from FishMIP input explorer: https://rstudio.global-ecosystem-model.cloud.edu.au/shiny/FishMIP_Input_Explorer/

```{r}
time_steps <- 1961:2010

# tos <- readRDS("FishMIP_Temperature_Forcing/tos.rds")
# t500m <- readRDS("FishMIP_Temperature_Forcing/t500m.rds")
# t1000m <- readRDS("FishMIP_Temperature_Forcing/t1000m.rds")
# t1500m <- readRDS("FishMIP_Temperature_Forcing/t1500m.rds")
# tob <- readRDS("FishMIP_Temperature_Forcing/tob.rds")

tos <- readRDS("tos_annual.rds")
t500m <- readRDS("t500m_annual.rds")
t1000m <- readRDS("t1000m_annual.rds")
t1500m <- readRDS("t1500m_annual.rds")
tob <- readRDS("tob_annual.rds")


# tos <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/tos_annual_WAO_corrected.rds")
# t500m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t500m_annual_WAO_corrected.rds")
# t1000m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t1000m_annual_WAO_corrected.rds")
# t1500m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t1500m_annual_WAO_corrected.rds")
# tob <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/tob_annual_WAO_corrected.rds")

ocean_temp <- abind::abind(tos,t500m,t1000m, t1500m, tob, along = 2)

realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")

colnames(ocean_temp) <- realm_names
rownames(ocean_temp) <- time_steps
```

Let's run a simulation and inspect the dynamics
```{r}
sim <- project(params, t_max = 500)
plot(sim)
plotlyBiomass(sim)
plotlyBiomassRelative(sim)
plotlySpectra(sim, power = 2, total = T)
```

Can we over write mammal scalars to 1?

We want metabolism and encounter for mammals to be 1

```{r}
#plotTherPerformance(params_pre_temp)
```


```{r}
plot(sim)
plotBiomass(sim)
plotFeedingLevel(sim, include_critical = T)
plotSpectra(sim,power=2)
plotBiomassObservedVsModel(sim,ratio=T)
plotYieldObservedVsModel(sim,ratio=T)
```


Set up time steps
```{r}
time_steps <- 1961:2010
```

## plankton forcing data
Set up the resource:
```{r}
isimip_plankton <- read.table("GFDL_resource_spectra_annual.dat") # log10 abundance
isimip_plankton <- as(isimip_plankton, "matrix")
sizes <- names(params@initial_n_pp)

n_pp_array <- array(NA, dim = c(length(time_steps), length(sizes)), dimnames = list(time = time_steps, w = sizes))

# rownames(isimip_plankton) <- time_steps
# colnames(isimip_plankton) <- sizes # signif(sizes, 6)

# dim(isimip_plankton)
# dim(n_pp_array)
```

Fill array
- Need an extra time step preceding the simulation
- IMPORTANT NOTE: CORRECTION TO  N_PP_ARRAY HERE

```{r}
n_pp_array[1,] <- (isimip_plankton[1,] - colMeans(isimip_plankton[,])) + log10(params@initial_n_pp*params@dw_full)
for (t in seq(1,length(time_steps) - 1,1)) {
  n_pp_array[t + 1,] <- (isimip_plankton[t,]  - colMeans(isimip_plankton[,])) +  log10(params@initial_n_pp*params@dw_full)
}

# saveRDS(n_pp_array,"n_pp_array_26_08_2024.RDS")
```


## Fishing effort forcing

Repeat this step for yield_observed

```{r}
time_steps_effort <- 1930:2010
isimip_effort <- read.csv("effort_array.csv")

years <- isimip_effort[,1]

isimip_effort <- isimip_effort[,-1]
rownames(isimip_effort) <- years

for (i in 1:ncol(isimip_effort)) {
  colnames(isimip_effort)[i] <- "knife_edge_gear"
}

head(isimip_effort)
glimpse(isimip_effort)
dim(isimip_effort)

# effort_2001_2010 <- isimip_effort[-c(1:71,82:90),]
effort_1930_2010 <- isimip_effort[-c(82:90),]

isimip_effort_array <- as(effort_1930_2010, "matrix")
# isimip_effort_array_1930_2010 <- as(effort_1930_2010, "matrix")

gear_name = colnames(isimip_effort)


effort_array <- array(NA, c(length(time_steps_effort), length(gear_name)), dimnames = list(time = time_steps_effort, gear = gear_name))
effort_array[1,] = isimip_effort_array[1,]
for (t in seq(1,length(time_steps_effort) - 1, 1)) {
  effort_array[t + 1,] <- isimip_effort_array[t + 1,]
}

effort_array[is.na(effort_array)] <- 0

head(effort_array)

# saveRDS(effort_array,"effort_array_26_08_2024.RDS")
# saveRDS(effort_array_full,"effort_array_full_26_08_2024.RDS")
```

```{r}
# effort_array <- effort_array_full

dim(effort_array)
head(effort_array)
effort_array[1,]

# gear_params(params_v2)$gear <- gear_params(params_v2)$species
gear_params(params)$gear <- gear_params(params)$species
# gear_params(params_steady_F)$gear <- gear_params(params_steady_F)$species

# gear_name = gear_params(params_steady_F)$gear
# gear_name = gear_params(params_v2)$gear
gear_name = gear_params(params)$gear
colnames(effort_array) <- gear_name
head(effort_array)
# 
# gear_params(params)$gear <- gear_name
# gear_params(params)$gear
# head(effort_array)

# saveRDS(effort_array, "effort_array_therMizer.RDS")
```

# Recalibrate model with Thermizer
- THIS PART IS FOR THE STEADY STATE ONLY

For this step, we re-estimate the reproduction/biomass parameters using time-averaged forcing data

First, reset initial n_pp to be time averaged value of plankton.

```{r}
# years_to_use_effort <- as.character(1951:1960)
years_to_use<-as.character(2001:2010) # period used for the steady-state calibration


initialNResource(params)[which(w_full(params) <100)]<-10^(colMeans(n_pp_array[years_to_use,])[which(w_full(params) < 100)])/params@dw_full[which(w_full(params) < 100)]
# #need to cut resource at 100g
# initialNResource(params)[which(w_full(params) >=10)]<-0.0

```

```{r}
# gear_name = gear_params(params)$gear
# colnames(effort_array) <- gear_name
# colnames(effort_array_full) <- gear_name
# head(effort_array)

# same with initial effort

initial_effort(params)<-colMeans(effort_array[years_to_use,])

# # use pre-collapse plank and temperatures
# params_pre <- upgradeTherParams(params, ocean_temp_array = ocean_temp[years_to_use,],n_pp_array = n_pp_array[years_to_use,],
#       # vertical_migration_array = vertical_migration_array,
#                             # exposure_array = ESS_exposure, 
#                             aerobic_effect = FALSE, metabolism_effect = FALSE)

params_pre_temp <- upgradeTherParams(params, ocean_temp_array = ocean_temp[years_to_use,],n_pp_array = n_pp_array[years_to_use,],
      # vertical_migration_array = vertical_migration_array,
                            # exposure_array = ESS_exposure, 
                            aerobic_effect = FALSE, metabolism_effect = TRUE)


plot(params)
# plot(params_pre)
plot(params_pre_temp)

# trim the effort array
# effort_array_pre <-effort_array_full[years_to_use,]

# saveRDS(params_pre,"params_for_use.RDS")

# plotFeedingLevel(params_pre, include_critical = T)
```
Try using calibrateBiomass to provide the new system with enough food, but then you must update the n_pp_array to factor the anomaly based on the initial_n_pp

Check initial_n_pp between the temp affects and not. 

```{r}
head(initialNResource(params))
# head(initialNResource(params_pre))
head(initialNResource(params_pre_temp))

# head(params_pre@initial_n_pp)

```


```{r}
params_pre_temp@species_params$species
```

[1] "mesozooplankton"          "other krill"              "other macrozooplankton"   "antarctic krill"         
 [5] "salps"                    "mesopelagic fishes"       "bathypelagic fishes"      "shelf and coastal fishes"
 [9] "flying birds"             "small divers"             "squids"                   "toothfishes"             
[13] "leopard seals"            "medium divers"            "large divers"             "minke whales"            
[17] "orca"                     "sperm whales"             "baleen whales"        

```{r}
# Add a new column to species_params to indicate thermal sensitivity
#params_pre_temp@species_params$is_ectothermic <- c(TRUE, TRUE, TRUE, TRUE,
#                                                    TRUE, TRUE, TRUE, TRUE,
#                                                    FALSE, FALSE, TRUE, TRUE,
#                                                    FALSE, FALSE, FALSE, FALSE,
#                                                    FALSE, FALSE,FALSE)
# 
# 
# # Create a function to conditionally apply temperature effects
# apply_temp_scaling <- function(params, temp_at_t) {
#     # Only apply scaling to ectothermic species
#     is_ecto <- species_params(params)$is_ectothermic
# 
#     # Calculate unscaled temperature effect
#     unscaled_temp_effect <-
#         temp_at_t * (temp_at_t - (species_params(params)$temp_min + 273)) *
#         ((species_params(params)$temp_max + 273) - temp_at_t)^(1/2)
# 
#     # Zero out temperature effects for endothermic species
#     unscaled_temp_effect[!is_ecto] <- 1
# 
#     return(unscaled_temp_effect)
# }


```



```{r}

sim <- project(params_pre_temp, t_max = 200)

plot(sim)
plotlyBiomass(sim)
```


```{r}
initial_effort(params_pre_temp)
params_pre_temp@gear_params$catchability

params_temp <- params_pre_temp %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady()

params_temp <- params_temp %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady()

plot(params_temp)
```


Try adjusting catchability to increase minke whale reproduction parameters
```{r}
species_params(params_temp) |> select(species, yield_observed)
gear_params(params_temp) |> select(gear, catchability)

params_temp@species_params$species
params_temp@species_params$yield_observed

plotYield(params_temp)

plotYieldObservedVsModel(params_temp)
#Test ratio
ratio<- plotYieldObservedVsModel(params_temp, return_data = TRUE)$ratio

gear_params(params_temp)$catchability[16]<-gear_params(params_temp)$catchability[16]/ratio[3] #minke 
gear_params(params_temp)$catchability[19]<-gear_params(params_temp)$catchability[19]/ratio[4] #baleen whales

# gear_params(params)$catchability[c(7,8,12,16,19)]<-gear_params(params)$catchability[c(7,8,12,16,19)]/ratio
```


```{r}

params_temp_v2 <- params_temp %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady()


plotYieldObservedVsModel(params_temp_v2)
```

```{r}
ratio<- plotYieldObservedVsModel(params_temp_v2, return_data = TRUE)$ratio

gear_params(params_temp_v2)$catchability[16]<-gear_params(params_temp_v2)$catchability[16]/ratio[3] #minke 
# gear_params(params_temp)$catchability[19]<-gear_params(params_temp)$catchability[19]/ratio[4] #baleen whales
```


```{r}
params_temp_v3 <- params_temp_v2 %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady() %>% matchBiomasses() %>%  steady()


plotYieldObservedVsModel(params_temp_v3)
plotFeedingLevel(params_temp_v3)

#params_temp_v3 <- tuneParams(params_temp_v3)

params_temp_v3@species_params$erepro
```


```{r}
sim_temp <- project(params_temp, t_max = 100)

plot(sim_temp)
plotBiomassObservedVsModel(sim_temp)
plotBiomassRelative(sim_temp)

sim_temp@params@species_params$erepro
```


```{r}

params_pre_v2 <- params_pre %>% calibrateBiomass() 

params_pre_v3 <- tuneParams(params_pre)
```


```{r}
head(initialNResource(params_pre))
head(initialNResource(params_pre_v2))
```


Nice test to see that things are working correctly
```{r}
plotTherPerformance(params_pre)
```

Load catch data
```{r}
yield_timeseries <- read.csv("FishMIP_fishing_data/catch_timeseries.csv")

rownames(yield_timeseries) <- yield_timeseries$Year

yield_timeseries <- yield_timeseries[,-1]

species_names <- species_params(params)$species

colnames(yield_timeseries) <- species_names

yield_array <- as(yield_timeseries, "matrix")

# saveRDS(yield_array, "yield_array_ts.RDS")
```


Attempt to achieve a pre-1961 steady-state using the yield from 1961

```{r}
yield_ts <- readRDS("yield_array_ts.RDS")

species_params(params_pre)$yield_observed <- colMeans(yield_ts[years_to_use_effort,])


```


```{r}
params_pre <- steady(params_pre, t_max = 100, tol = 0.001)

params_pre <- params_pre|> 
  matchYields() |> steady(t_max = 1000, tol = 0.001) |> matchYields()|> steady(t_max = 1000, tol = 0.001) |>
  matchYields() |> steady(t_max = 1000, tol = 0.001) |> matchYields()|> steady(t_max = 1000, tol = 0.001) |>
  matchYields() |> steady(t_max = 1000, tol = 0.001) |> matchYields()|> steady(t_max = 1000, tol = 0.001) |>
  matchYields() |> steady(t_max = 1000, tol = 0.001) |> matchYields()|> steady(t_max = 1000, tol = 0.001) |>
  matchYields() |> steady(t_max = 1000, tol = 0.001) |> matchYields()|> steady(t_max = 1000, tol = 0.001)

```




```{r}
plot(params_pre)
```


























