---
title: "09 Monte Carlo parameter exploration"
subtitle: "Preliminary uncertainty analysis for the Prydz Bay therMizer model"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE,
                      fig.width = 12, fig.height = 8)
```

# Overview

When the model calibrated to the contemporary (2010–2020) ecosystem is run through
the historical period with whaling and fishing (1841–2010), several groups, notably
the whales, end far from their observed contemporary biomass. This notebook explores
which values of three poorly constrained parameter groups are consistent with the
model ending near the observations:

| Parameter group | Perturbed | Multiplier applies to |
|---|---|---|
| Gear catchability | the 9 fished gears | each gear |
| Search-volume coefficient (gamma) | all 19 groups | each group |
| Initial abundance | all 19 groups | each group |

Each ensemble member draws log-normal multipliers, is re-equilibrated with
`steady()`, spun up without fishing and tested for stability, then run from 1841 to
2010 with historical climate and fishing. Members whose mean 2000–2010 biomass lies
within the observational bounds for all but at most three groups are accepted
(rejection sampling). **Stage 1** samples widely around the reference model
(`params_sel_adj.rds`); **stage 2** samples again around the centre of the stage-1
accepted multipliers. The stage-2 centre, relative to the reference model, is the
parameter set carried forward.

The step-by-step description, and every function used here, is in `mc_functions.R`.

## Relation to the original 09 notebook

This notebook keeps the sampling scheme (the same member id and base model give the
same random draws as the original code), the parameter bounds, the steady-state and
spin-up settings, the stability test, the acceptance rule and the use of the mean
of accepted multipliers as the centre. Changes:

1. The first spin-up cycle no longer passes `initial_n = NULL` to `project()`, which
   errors in mizer 2.4–2.5 (every member failed with "replacement has length zero").
2. `steady()` non-convergence (reported by `message()`, not `warning()`, in mizer 2.5)
   and erepro values raised by `steady()` are recorded for every member. Rejecting
   them is optional and off by default, as in the original; the duplicate `steady()`
   call that followed any warning is gone.
3. The pre-check uses `maturity()`; the original called `getMaturity()`, which does
   not exist in mizer 2.5, so that check was silently skipped.
4. Multipliers are also stored relative to the reference model, so the stage-2
   centre includes the stage-1 shift. The original applied each round's centre to
   `params_sel_adj`, which dropped the previous round's shift after the first round.
5. Member ids must not overlap between batches or stages. The original restarted at
   1 in every batch, so batches and rounds reused the same random draws.
6. Results are saved per block as biomass, yield, multipliers and diagnostics
   (full `MizerSim` objects optional) and interrupted runs resume.
7. The original stage 1 perturbed `params_sel_adj_v2` (hand-set catchabilities);
   later steps used `params_sel_adj.rds`. This notebook uses `params_sel_adj.rds`
   throughout.
8. The optimiser, adaptive-search, best-member and species-wise hybrid sections are
   not included.

# 1. Setup

```{r libraries}
suppressPackageStartupMessages({
  library(mizer)       # developed with mizer 2.5.0
  library(therMizer)
  library(ggplot2)
})
source("mc_functions.R")
```

```{r switches}
paths <- list(
  base_params = "params_sel_adj.rds",                  # reference model (therMizer, 1841-2010 forcing)
  effort      = "effort_array_1841_2010.rds",          # historical effort [year x gear]
  yield_obs   = "yield_observed_timeseries_tidy.RDS",  # observed catches: Year, Species, Yield (g/yr)
  out_dir     = file.path("Output_large_files", "monte_carlo_results", "consolidated")
)

run_stage1 <- FALSE   # TRUE: run (or resume) stage 1; FALSE: load saved results
run_stage2 <- FALSE   # TRUE: run (or resume) stage 2; FALSE: load saved results
quick_test <- FALSE   # TRUE: 6 members per stage with short, relaxed settings (checks the code only)
n_cores <- max(1, parallel::detectCores() - 1)

if (quick_test) paths$out_dir <- paste0(paths$out_dir, "_quicktest")
```

# 2. Inputs

```{r inputs}
base_params <- readRDS(paths$base_params)
effort      <- readRDS(paths$effort)
yield_obs   <- readRDS(paths$yield_obs)

species_order <- species_params(base_params)$species

# The historical run's years come from the effort array
range(as.numeric(dimnames(effort)[[1]]))
```

Observed biomass (2010–2020 means, McCormack et al. 2020) and acceptance bounds in
grams. Data-derived ranges are used where available; elsewhere the bounds are 0.5–1.5
times the observed mean.

```{r observations}
obs_biomass <- data.frame(
  Species = species_order,
  ObsBiomass = c(1.297420e+13, 2.801248e+12, 1.474341e+13, 5.897364e+12, 9.612703e+11,
                 1.769209e+12, 1.769209e+12, 4.027900e+12, 4.423023e+09, 2.358946e+10,
                 2.211512e+11, 1.105756e+12, 2.948682e+09, 3.907004e+11, 1.621775e+10,
                 2.064077e+10, 8.846046e+09, 1.621775e+10, 1.872413e+11)
)
bounds <- mc_biomass_bounds(
  obs_biomass,
  lower_g = c(NA, NA, NA, 3.24355e12, NA, NA, NA, NA, NA, 2.3589456e10, NA, NA,
              1.474341e9, 2.41674e11, 1.1794728e10, 1.9166433e10, 7.371705e9,
              1.0320387e10, 4.5704571e10),
  upper_g = c(NA, NA, NA, 1.0173e13, NA, NA, NA, NA, NA, 2.6538138e10, NA, NA,
              5.897364e9, 5.20251e11, 3.0961161e10, 2.2115115e10, 1.0320387e10,
              1.6217751e10, 2.99291e11),
  assumed_fraction = 0.5
)
knitr::kable(transform(bounds, ObsBiomass = signif(ObsBiomass / 1e6, 3),
                       Lower_t = signif(Lower_g / 1e6, 3), Upper_t = signif(Upper_g / 1e6, 3))[,
             c("Species", "ObsBiomass", "Lower_t", "Upper_t", "RangeSource")],
             caption = "Observed biomass and acceptance bounds (tonnes)")
```

# 3. Settings

Stage 1 uses the defaults in `mc_settings()`, which reproduce the last stage-1 run of
the original notebook. Stage 2 uses the refinement settings of the original rerun.

```{r settings}
settings_stage1 <- mc_settings()
settings_stage2 <- mc_settings(catchability_sd = 1, abundance_sd = 4, gamma_sd = 2,
                               steady_tol = 0.002, steady_t_max = 1200,
                               spinup_cycles = 3, spinup_min_cycles = 2)
ids_stage1 <- 1:1200             # further stage-1 batches: continue from 1201
ids_stage2 <- 100000 + 1:1000    # stage-2 ids never overlap stage-1 ids

if (quick_test) {
  test_overrides <- list(steady_t_max = 50, steady_tol = 0.05, stability_cv_max = 100,
                         stability_trend_rel_slope_max = 100, accept_max_outside = 19)
  settings_stage1 <- modifyList(settings_stage1, test_overrides)
  settings_stage2 <- modifyList(settings_stage2, test_overrides)
  ids_stage1 <- 1:6
  ids_stage2 <- 100000 + 1:6
  n_cores <- min(n_cores, 2)
}

knitr::kable(mc_settings_table(stage1 = settings_stage1, stage2 = settings_stage2),
             caption = "Monte Carlo settings")
```

# 4. Stage 1: broad exploration

Expect roughly 1–2.5 minutes per member per core with the full settings.

```{r stage1-run}
dir_stage1 <- file.path(paths$out_dir, "stage1")
if (run_stage1) {
  mc_run_stage(base_params, effort, ids_stage1, settings_stage1, dir_stage1, n_cores = n_cores)
}
stage1 <- mc_load_stage(dir_stage1)
```

```{r stage1-accept}
acc1 <- mc_accept(stage1, bounds)
ids_acc1 <- acc1$accepted_ids
knitr::kable(mc_funnel(stage1, acc1), caption = "Stage 1: members at each step")
knitr::kable(mc_outside_by_group(acc1), caption = "Stage 1: members outside the bounds, by group")
```

```{r stage1-parameters, fig.height = 9}
mc_plot_multipliers(stage1, ids_acc1, "catchability")
mc_plot_multipliers(stage1, ids_acc1, "gamma", group_order = species_order)
mc_plot_multipliers(stage1, ids_acc1, "abundance", group_order = species_order)
knitr::kable(mc_multiplier_summary(stage1, ids_acc1, scale = "stage"), digits = 3,
             caption = "Stage 1: multipliers, all members run vs accepted")
```

```{r stage1-envelopes}
mc_plot_biomass(stage1, ids_acc1, bounds, species_order = species_order)
mc_plot_yield(stage1, ids_acc1, yield_obs, species_order = species_order)
```

Centre of the accepted multipliers (arithmetic mean, as in the original analysis;
`stat = "geometric_mean"` or `"median"` are the alternatives for log-normal
multipliers). Applying it to the reference model gives the stage-2 base.

```{r stage1-centre}
centre1 <- mc_centre(stage1, ids_acc1, stat = "mean")
write.csv(mc_centre_table(centre1, base_params),
          file.path(dir_stage1, "centre_stage1_mean.csv"), row.names = FALSE)
params_centre1 <- mc_apply_centre(base_params, centre1)
knitr::kable(mc_centre_table(centre1, base_params), digits = 3,
             caption = "Stage 1 centre (multipliers relative to the reference model)")
```

# 5. Stage 2: refinement around the stage-1 centre

Stage-2 multipliers are drawn relative to the stage-1 centre (`"stage"` scale) and
also stored relative to the reference model (`"reference"` scale).

```{r stage2-run}
dir_stage2 <- file.path(paths$out_dir, "stage2")
if (run_stage2) {
  mc_run_stage(params_centre1, effort, ids_stage2, settings_stage2, dir_stage2,
               reference_params = base_params, n_cores = n_cores)
}
stage2 <- mc_load_stage(dir_stage2)
```

```{r stage2-accept}
acc2 <- mc_accept(stage2, bounds)
ids_acc2 <- acc2$accepted_ids
knitr::kable(mc_funnel(stage2, acc2), caption = "Stage 2: members at each step")
knitr::kable(mc_outside_by_group(acc2), caption = "Stage 2: members outside the bounds, by group")
```

```{r stage2-parameters, fig.height = 9}
mc_plot_multipliers(stage2, ids_acc2, "catchability")
mc_plot_multipliers(stage2, ids_acc2, "gamma", group_order = species_order)
mc_plot_multipliers(stage2, ids_acc2, "abundance", group_order = species_order)
```

```{r stage2-envelopes}
mc_plot_biomass(stage2, ids_acc2, bounds, species_order = species_order)
mc_plot_yield(stage2, ids_acc2, yield_obs, species_order = species_order)
```

# 6. Final parameter set

The mean of the stage-2 accepted multipliers on the reference scale, applied to the
reference model without re-running `steady()` (as for the original
`params_mean_from_rerun_results.rds`).

```{r final-params}
centre2 <- mc_centre(stage2, ids_acc2, stat = "mean", scale = "reference")
write.csv(mc_centre_table(centre2, base_params),
          file.path(dir_stage2, "centre_stage2_mean_reference.csv"), row.names = FALSE)
params_mc_centre <- mc_apply_centre(base_params, centre2)
saveRDS(params_mc_centre, file.path(paths$out_dir, "params_mc_centre_stage2.rds"))
knitr::kable(mc_centre_table(centre2, base_params), digits = 3,
             caption = "Final centre (multipliers relative to the reference model)")
```

# 7. Numbers for the Methods

```{r methods-numbers}
f1 <- mc_funnel(stage1, acc1)
f2 <- mc_funnel(stage2, acc2)
m1 <- mc_members(stage1); m1a <- m1[m1$member_id %in% ids_acc1, ]
m2 <- mc_members(stage2); m2a <- m2[m2$member_id %in% ids_acc2, ]
pct <- function(x) sprintf("%.1f%%", 100 * mean(x, na.rm = TRUE))
```

- Stage 1 ran `r f1$n[1]` members; `r f1$n[6]` completed every step and `r f1$n[7]`
  (`r sprintf("%.1f%%", 100 * f1$n[7] / f1$n[1])`) met the biomass criterion.
- Stage 2 ran `r f2$n[1]` members; `r f2$n[6]` completed every step and `r f2$n[7]`
  (`r sprintf("%.1f%%", 100 * f2$n[7] / f2$n[1])`) met the biomass criterion.
- Accepted members whose steady state converged within the tolerance: stage 1
  `r pct(m1a$steady_converged)`, stage 2 `r pct(m2a$steady_converged)`.
- Accepted members with any erepro above 1 after `steady()`: stage 1
  `r pct(m1a$max_erepro > 1)`, stage 2 `r pct(m2a$max_erepro > 1)`.

```{r methods-tables}
knitr::kable(mc_multiplier_summary(stage2, ids_acc2, scale = "reference"), digits = 3,
             caption = "Stage 2 accepted multipliers relative to the reference model")
```

# 8. Session information

```{r session}
stage1$info$versions
sessionInfo()
```
