---
title: "R Notebook"
output: html_notebook
---

# Uncertainty Analysis for Prydz Bay Mizer Model

This document implements a comprehensive uncertainty analysis for our Prydz Bay ecosystem model to assess how parameter uncertainty affects model projections. Understanding uncertainty in complex marine ecosystem models is essential for robust ecosystem-based management and provides context for interpreting historical and future projections.

Our approach combines Monte Carlo simulations with biomass-based rejection criteria to:

1. Quantify uncertainty in key model parameters (catchability, search rates, initial abundance)
2. Assess parameter distributions before and after applying observation-based constraints
3. Generate uncertainty bounds for model projections of biomass and yield
4. Evaluate model performance against historical observations

We systematically vary parameters with ecological significance, including fishing catchability coefficients (gear efficiency), search rates (foraging behavior), and initial abundances (population sizes). The Monte Carlo framework allows us to explore parameter space comprehensively while the rejection criteria ensure biological realism by filtering out parameter combinations that produce unrealistic ecosystem states relative to observed biomass data.

The resulting uncertainty bounds provide critical context for interpreting model projections and identifying which species groups have well-constrained versus highly uncertain dynamics in our Prydz Bay ecosystem model.

Do I need contant phytoplankton and temperature forcings for any spin-up, like if we need to use steady() in a loop testing parameter uncertaitny?

Also need to make sure initial effort is zero for the repreated forcing during a spin-up like above 

# Core Functions for Uncertainty Analysis

The following functions provide the core functionality for Monte Carlo uncertainty analysis,
parameter distribution assessment, and biomass-based rejection criteria.

```{r core-functions}
# Function to run enhanced combined uncertainty simulations
run_enhanced_combined_uncertainty_sims <- function(params,
                                                   n_sims = 10,
                                                   catchability_sd = 1,
                                                   abundance_sd = 1,
                                                   gamma_sd = 1,
                                                   marine_mammal_species = NULL,
                                                   effort_scen,
                                                   enhanced_steady_tol = 0.002,
                                                   enhanced_steady_t_max = 1000,
                                                   spinup_years = 118,
                                                   t_start = 1841,
                                                   sim_years = 170,
                                                   preserve_erepro = TRUE,
                                                   max_erepro = 1,
                                                   verbose = FALSE) {
  
  # Default marine mammal species if not specified
  if (is.null(marine_mammal_species)) {
    marine_mammal_species <- c("minke whales", "orca", "sperm whales", "baleen whales")
  }
  
  # Progress message
  if (verbose) {
    cat("=== ENHANCED COMBINED UNCERTAINTY ANALYSIS ===\n")
    cat("Running", n_sims, "simulations with enhanced protocols\n")
    cat("  - Catchability SD:", catchability_sd, "\n")
    cat("  - Abundance SD:", abundance_sd, "\n")
    cat("  - Gamma SD:", gamma_sd, "\n")
    cat("  - Spinup years:", spinup_years, "(unfished)\n")
    cat("  - Biomass stability test: CV < 10% over final 40 years\n")
  }
  
  # Storage for results
  sim_results <- list()
  param_combinations <- list()
  failed_sims <- list()
  successful_count <- 0
  
  # Run simulations
  for(i in 1:n_sims) {
    if (verbose && i %% 5 == 0) {
      cat("  Processing simulation", i, "of", n_sims, "\n")
    }
    
    tryCatch({
      # Create copy of params
      rand_params <- params
      
      # Step 1: Randomize catchability for fished species
      gear_df <- gear_params(rand_params)
      original_catchability <- gear_df$catchability
      
      if(nrow(gear_df) > 0) {
        for(j in 1:nrow(gear_df)) {
          if(gear_df$catchability[j] > 0) {
            noise <- rnorm(1, mean = 0, sd = catchability_sd)
            gear_df$catchability[j] <- gear_df$catchability[j] * exp(noise)
            gear_df$catchability[j] <- min(1, max(0, gear_df$catchability[j]))
          }
        }
        gear_params(rand_params) <- gear_df
      }
      
      # Step 2: Randomize gamma (search rate) for all species
      sp_params <- species_params(rand_params)
      original_gamma <- sp_params$gamma
      
      for(j in 1:nrow(sp_params)) {
        gamma_noise <- rnorm(1, mean = 0, sd = gamma_sd)
        sp_params$gamma[j] <- sp_params$gamma[j] * exp(gamma_noise)
        sp_params$gamma[j] <- min(1000, max(0.01, sp_params$gamma[j]))
      }
      species_params(rand_params) <- sp_params
      
      # Step 3: Randomize initial abundance for all species
      all_species <- sp_params$species
      scaling_factors <- exp(rnorm(length(all_species), mean = 0, sd = abundance_sd))
      
      for (k in seq_along(all_species)) {
        rand_params@initial_n[k,] <- rand_params@initial_n[k,] * scaling_factors[k]
      }
      
      # Step 4: Run to steady state with randomized parameters
      if (preserve_erepro) {
        rand_params <- steady(rand_params, tol = enhanced_steady_tol, 
                             t_max = enhanced_steady_t_max, preserve = c("erepro"))
      } else {
        rand_params <- steady(rand_params, tol = enhanced_steady_tol, 
                             t_max = enhanced_steady_t_max)
        
        # Check erepro constraint
        current_erepro <- species_params(rand_params)$erepro
        if (any(current_erepro > max_erepro, na.rm = TRUE)) {
          next  # Skip this simulation
        }
      }
      
      # Step 5: Run unfished spinup to test biomass stability
      sim_spinup <- project(rand_params, 
                           t_start = t_start, 
                           t_max = spinup_years,
                           effort = 0)  # No fishing during spinup
      
      # Step 6: Test biomass stability over spinup period
      biomass_matrix <- getBiomass(sim_spinup)
      check_years <- min(40, spinup_years)  # Use final 40 years or all years if spinup < 40
      
      if (nrow(biomass_matrix) < check_years) {
        next  # Skip if insufficient data
      }
      
      # Get final years for stability check
      final_years_data <- tail(biomass_matrix, check_years)
      
      # Calculate CV for each species over the check period
      cv_by_species <- apply(final_years_data, 2, function(x) {
        if (mean(x) == 0) return(Inf)
        sd(x) / mean(x)
      })
      
      # Check if all species have CV < 10% (0.1)
      stability_threshold <- 0.1
      all_stable <- all(cv_by_species < stability_threshold, na.rm = TRUE)
      
      if (!all_stable) {
        if (verbose) {
          unstable_species <- names(cv_by_species)[cv_by_species >= stability_threshold]
          cat("    Sim", i, ": Failed stability test. Unstable:", 
              paste(unstable_species, collapse = ", "), "\n")
        }
        next  # Skip this simulation if biomass not stable
      }
      
      if (verbose) {
        max_cv <- max(cv_by_species, na.rm = TRUE)
        cat("    Sim", i, ": Passed stability test. Max CV:", round(max_cv * 100, 1), "%\n")
      }
      
      # Step 7: Run main simulation using final spinup state as initial condition
      final_initial_n <- sim_spinup@n[spinup_years,,]
      
      sim <- project(rand_params,
                    initial_n = final_initial_n,
                    t_start = t_start,
                    t_max = sim_years,
                    effort = effort_scen)
      
      # Store results
      successful_count <- successful_count + 1
      sim_results[[successful_count]] <- sim
      param_combinations[[successful_count]] <- list(
        sim_id = i,
        catchability = gear_df$catchability,
        catchability_change = gear_df$catchability / original_catchability,
        gamma_values = sp_params$gamma,
        gamma_change = sp_params$gamma / original_gamma,
        abundance_scaling = scaling_factors,
        species_names = all_species,
        max_cv = max(cv_by_species, na.rm = TRUE),
        stable = TRUE
      )
      
    }, error = function(e) {
      failed_sims[[length(failed_sims) + 1]] <- list(
        sim_id = i,
        error = e$message
      )
      if (verbose) {
        cat("    Sim", i, ": Error -", e$message, "\n")
      }
    })
  }
  
  if (verbose) {
    cat("Completed", successful_count, "successful simulations out of", n_sims, "attempts\n")
    if (length(failed_sims) > 0) {
      cat("Failed simulations:", length(failed_sims), "\n")
    }
  }
  
  return(list(
    simulations = sim_results,
    parameters = param_combinations,
    failed = failed_sims,
    n_successful = length(sim_results),
    n_attempts = n_sims
  ))
}

# Function to extract and standardize simulation outputs
extract_simulation_data <- function(sim_list, metric_type = "yield") {
  
  if (metric_type == "yield") {
    data_list <- lapply(seq_along(sim_list), function(i) {
      sim <- sim_list[[i]]
      yield_matrix <- getYield(sim)
      yield_data <- reshape2::melt(yield_matrix)
      names(yield_data) <- c("Year", "Species", "Yield")
      yield_data$sim_id <- i
      return(yield_data)
    })
  } else if (metric_type == "biomass") {
    data_list <- lapply(seq_along(sim_list), function(i) {
      sim <- sim_list[[i]]
      biomass_matrix <- getBiomass(sim)
      biomass_data <- reshape2::melt(biomass_matrix)
      names(biomass_data) <- c("Year", "Species", "Biomass")
      biomass_data$sim_id <- i
      return(biomass_data)
    })
  }
  
  # Combine all simulations
  combined_data <- do.call(rbind, data_list)
  return(combined_data)
}

# Function to plot parameter distributions
plot_parameter_distributions <- function(param_combinations, param_type = "gamma", 
                                       title_suffix = "") {
  require(ggplot2)
  require(tidyr)
  require(dplyr)
  
  if (param_type == "gamma") {
    # Extract gamma values for all species
    gamma_data <- do.call(rbind, lapply(seq_along(param_combinations), function(i) {
      params <- param_combinations[[i]]
      data.frame(
        sim_id = i,
        species = params$species_names,
        gamma = params$gamma_values,
        gamma_change = params$gamma_change
      )
    }))
    
    p <- ggplot(gamma_data, aes(x = gamma_change)) +
      geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7, color = "black") +
      facet_wrap(~species, scales = "free") +
      theme_bw() +
      labs(
        title = paste("Gamma (Search Rate) Parameter Distribution", title_suffix),
        x = "Gamma Change Factor (relative to original)",
        y = "Frequency"
      ) +
      geom_vline(xintercept = 1, linetype = "dashed", color = "red", alpha = 0.7)
    
  } else if (param_type == "catchability") {
    # Extract catchability values
    catch_data <- do.call(rbind, lapply(seq_along(param_combinations), function(i) {
      params <- param_combinations[[i]]
      if (length(params$catchability) > 0) {
        data.frame(
          sim_id = i,
          gear = 1:length(params$catchability),
          catchability = params$catchability,
          catchability_change = params$catchability_change
        )
      } else {
        NULL
      }
    }))
    
    if (!is.null(catch_data)) {
      p <- ggplot(catch_data, aes(x = catchability_change)) +
        geom_histogram(bins = 30, fill = "darkgreen", alpha = 0.7, color = "black") +
        facet_wrap(~gear, scales = "free") +
        theme_bw() +
        labs(
          title = paste("Catchability Parameter Distribution", title_suffix),
          x = "Catchability Change Factor (relative to original)",
          y = "Frequency"
        ) +
        geom_vline(xintercept = 1, linetype = "dashed", color = "red", alpha = 0.7)
    } else {
      p <- ggplot() + labs(title = "No catchability data available")
    }
    
  } else if (param_type == "abundance") {
    # Extract abundance scaling values
    abund_data <- do.call(rbind, lapply(seq_along(param_combinations), function(i) {
      params <- param_combinations[[i]]
      data.frame(
        sim_id = i,
        species = params$species_names,
        abundance_scaling = params$abundance_scaling
      )
    }))
    
    p <- ggplot(abund_data, aes(x = abundance_scaling)) +
      geom_histogram(bins = 30, fill = "coral", alpha = 0.7, color = "black") +
      facet_wrap(~species, scales = "free") +
      theme_bw() +
      labs(
        title = paste("Initial Abundance Scaling Distribution", title_suffix),
        x = "Abundance Scaling Factor (relative to original)",
        y = "Frequency"
      ) +
      geom_vline(xintercept = 1, linetype = "dashed", color = "red", alpha = 0.7)
  }
  
  return(p)
}

# Function to apply biomass rejection criteria
apply_biomass_rejection_criteria <- function(sim_results, param_combinations, 
                                           obs_biomass_data, 
                                           tolerance = 0.25,
                                           comparison_years = 2000:2010,
                                           verbose = TRUE) {
  
  if (verbose) {
    cat("=== APPLYING BIOMASS REJECTION CRITERIA ===\n")
    cat("Tolerance: ±", tolerance * 100, "%\n")
    cat("Comparison period:", min(comparison_years), "-", max(comparison_years), "\n")
    cat("Total simulations to assess:", length(sim_results), "\n")
  }
  
  # Calculate bounds for observed biomass
  obs_bounds <- obs_biomass_data %>%
    mutate(
      lower_bound = ObsBiomass * (1 - tolerance),
      upper_bound = ObsBiomass * (1 + tolerance)
    )
  
  # Check each simulation
  accepted_indices <- c()
  rejection_details <- list()
  
  for (i in seq_along(sim_results)) {
    sim <- sim_results[[i]]
    
    # Extract biomass for comparison years
    biomass_matrix <- getBiomass(sim)
    year_indices <- which(as.numeric(rownames(biomass_matrix)) %in% comparison_years)
    
    if (length(year_indices) == 0) {
      if (verbose) cat("Simulation", i, ": No data for comparison years\n")
      next
    }
    
    # Calculate mean biomass over comparison period
    mean_biomass <- colMeans(biomass_matrix[year_indices, , drop = FALSE])
    
    # Check if all species fall within bounds
    species_results <- sapply(names(mean_biomass), function(sp) {
      if (sp %in% obs_bounds$Species) {
        obs_row <- obs_bounds[obs_bounds$Species == sp, ]
        modeled_value <- mean_biomass[sp]
        within_bounds <- modeled_value >= obs_row$lower_bound && 
                        modeled_value <= obs_row$upper_bound
        return(within_bounds)
      } else {
        return(TRUE)  # Accept if no observed data available
      }
    })
    
    # Accept simulation only if ALL species are within bounds
    all_within_bounds <- all(species_results, na.rm = TRUE)
    
    if (all_within_bounds) {
      accepted_indices <- c(accepted_indices, i)
    }
    
    # Store rejection details
    rejection_details[[i]] <- list(
      sim_id = i,
      accepted = all_within_bounds,
      species_results = species_results,
      mean_biomass = mean_biomass
    )
  }
  
  # Filter results
  accepted_sims <- sim_results[accepted_indices]
  accepted_params <- param_combinations[accepted_indices]
  
  if (verbose) {
    cat("Accepted simulations:", length(accepted_sims), "out of", length(sim_results), 
        "(", round(length(accepted_sims)/length(sim_results)*100, 1), "%)\n")
    
    # Show rejection reasons
    rejected_details <- rejection_details[setdiff(1:length(sim_results), accepted_indices)]
    if (length(rejected_details) > 0) {
      cat("\nRejection summary by species:\n")
      species_rejection_count <- table(unlist(lapply(rejected_details, function(x) {
        names(x$species_results)[!x$species_results]
      })))
      print(species_rejection_count)
    }
  }
  
  return(list(
    accepted_simulations = accepted_sims,
    accepted_parameters = accepted_params,
    accepted_indices = accepted_indices,
    rejection_details = rejection_details,
    obs_bounds = obs_bounds,
    n_accepted = length(accepted_sims),
    n_total = length(sim_results),
    acceptance_rate = length(accepted_sims) / length(sim_results)
  ))
}
```


```{r setup-libraries}
library(therMizer)
library(mizer)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(reshape2)
library(viridis)
```


# Load params
Best version of Pyrdz Bay mizer model parameters at steady-state for the contemporary ecosystem 2010 - 2020
```{r}
params_new_v4 <- readRDS("params_steady_state_2011_2020_tol_0.00025.RDS")

combined_effort_array <- readRDS("effort_array_1841_2010.rds")
```

# Load climate forcings

The temperature and phytoplankton climate forcings for the ISIMIP3a protocol for Prydz Bay region
In some cases we need repeated constant values of the first year of these forcings, either for spin-ups or to mimin 'no climate change'. These repeated forcings are created as starting in 1961 and a version that starts in 1841

```{r}
# Temperature
extended_ocean_temp <- readRDS("temperature_forcing_1841_2010.rds")
constant_array_temp <- readRDS("constant_array_temp.RDS")
constant_array_temp_1841 <- readRDS("constant_array_temp_1841.RDS")

# Phytoplankton
extended_n_pp_array <- readRDS("phytoplankton_forcing_1841_2010.rds")
constant_array_n_pp <- readRDS("constant_array_n_pp.RDS")
constant_array_n_pp_1841 <- readRDS("constant_array_n_pp_1841.RDS")
```

# Setup params

```{r setup-params}
params_1841_2010_climate_only <- upgradeTherParams(params_new_v4, 
                                                   ocean_temp_array = extended_ocean_temp,
                                                   n_pp_array = extended_n_pp_array,
                                                   # vertical_migration_array = vertical_migration_array,
                                                   # exposure_array = ESS_exposure,
                                                   aerobic_effect = FALSE, metabolism_effect = TRUE)

params_1841_2010_fishing_only <- upgradeTherParams(params_new_v4, 
                                                   ocean_temp_array = constant_array_temp_1841,
                                                   n_pp_array = constant_array_n_pp_1841,
                                                   # vertical_migration_array = vertical_migration_array,
                                                   # exposure_array = ESS_exposure,
                                                   aerobic_effect = FALSE, metabolism_effect = TRUE)
```

# Run baseline simulations

```{r run-baseline-simulations}
# Spinup simulation (unfished)
sim_1841_2010_climate_only_spinup <- project(params_1841_2010_climate_only, 
                                             t_start = 1841, 
                                             t_max = 118,  
                                             effort = 0)

# Climate only simulation
sim_1841_2010_climate_only <- project(params_1841_2010_climate_only,
                                      initial_n=sim_1841_2010_climate_only_spinup@n[118,,], 
                                      t_start = 1841, 
                                      t_max = 170,  
                                      effort = 0)

# Fishing only simulation  
sim_1841_2010_fishing_only <- project(params_1841_2010_fishing_only,
                                      initial_n=sim_1841_2010_climate_only_spinup@n[118,,],
                                      t_start = 1841, 
                                      t_max = 170,  
                                      effort = combined_effort_array)

# Climate and fishing simulation
sim_1841_2010_climate_and_fishing <- project(params_1841_2010_climate_only,
                                             initial_n=sim_1841_2010_climate_only_spinup@n[118,,],
                                             t_start = 1841, 
                                             t_max = 170,  
                                             effort = combined_effort_array)
```


params_1841_2010_fishing_only <- upgradeTherParams(params_new_v4, 
                                                   ocean_temp_array = constant_array_temp_1841,
                                                   n_pp_array = constant_array_n_pp_1841,
                                                   # vertical_migration_array = vertical_migration_array,
                                                   # exposure_array = ESS_exposure,
                                                   aerobic_effect = FALSE, metabolism_effect = TRUE)


```

# Run simulations

```{r}
sim_1841_2010_climate_only_spinup <- project(params_1841_2010_climate_only, 
                                             t_start = 1841, 
                                             t_max = 118,  
                                             effort = 0)
```


```{r}
sim_1841_2010_climate_only <- project(params_1841_2010_climate_only,
                                      initial_n=sim_1841_2010_climate_only_spinup@n[118,,], 
                                      t_start = 1841, 
                                      t_max = 170,  
                                      effort = 0)
```


```{r}
sim_1841_2010_fishing_only <- project(params_1841_2010_fishing_only,
                                      initial_n=sim_1841_2010_climate_only_spinup@n[118,,],
                                      t_start = 1841, 
                                      t_max = 170,  
                                      effort = combined_effort_array)
```


```{r}
sim_1841_2010_climate_and_fishing <- project(params_1841_2010_climate_only,
                                             initial_n=sim_1841_2010_climate_only_spinup@n[118,,],
                                             t_start = 1841, 
                                             t_max = 170,  
                                             effort = combined_effort_array)
```





# Size spectrum

```{r}
mizerExperimental::plotSpectra2(sim_1841_2010_climate_only, sim_1841_2010_fishing_only, time_range = 1841:1860)
mizerExperimental::plotSpectra2(sim_1841_2010_climate_only, sim_1841_2010_fishing_only, time_range = 1940:1950)
mizerExperimental::plotSpectra2(sim_1841_2010_climate_only, sim_1841_2010_fishing_only, time_range = 2000:2010)

plotSpectraRelative(sim_1841_2010_climate_only, sim_1841_2010_fishing_only, time_range = 1920:1930)
plotSpectraRelative(sim_1841_2010_climate_only, sim_1841_2010_fishing_only, time_range = 1930:1940)
plotSpectraRelative(sim_1841_2010_climate_only, sim_1841_2010_fishing_only, time_range = 1940:1950)
plotSpectraRelative(sim_1841_2010_climate_only, sim_1841_2010_fishing_only, time_range = 2000:2010)
```


# Initial Model Assessment

Let's first examine the baseline model performance before uncertainty analysis.

```{r baseline-assessment}
# Load observed data
yield_ts_tidy <- readRDS("yield_observed_timeseries_tidy.RDS")

# Observed biomass data (2010-2020 means from McCormack et al. 2020)
obs_biomass_data <- data.frame(
  Species = c("mesozooplankton","other krill","other macrozooplankton","antarctic krill",
              "salps","mesopelagic fishes","bathypelagic fishes","shelf and coastal fishes",
              "flying birds","small divers","squids","toothfishes","leopard seals",
              "medium divers","large divers","minke whales","orca","sperm whales","baleen whales"),
  ObsBiomass = c(1.297420e+13, 2.801248e+12, 1.474341e+13, 5.897364e+12, 9.612703e+11, 
                 1.769209e+12, 1.769209e+12, 4.027900e+12, 4.423023e+09, 2.358946e+10, 
                 2.211512e+11, 1.105756e+12, 2.948682e+09, 3.907004e+11, 1.621775e+10, 
                 2.064077e+10, 8.846046e+09, 1.621775e+10, 1.872413e+11)
)

# Basic plotting function for biomass comparison
plot_baseline_biomass <- function(sim_object, obs_data, obs_years = 2010:2020) {
  require(ggplot2)
  require(dplyr)
  
  # Extract modeled biomass
  mod_biomass_matrix <- getBiomass(sim_object)
  df_mod_biomass <- reshape2::melt(mod_biomass_matrix)
  names(df_mod_biomass) <- c("Year", "Species", "Biomass")
  
  # Create time series version for observation years
  obs_biomass_ts <- obs_data %>%
    mutate(Lower = ObsBiomass * 0.9, Upper = ObsBiomass * 1.1) %>%
    crossing(Year = obs_years)
  
  # Plot
  ggplot() +
    geom_line(data = df_mod_biomass, aes(x = Year, y = Biomass, color = Species), linewidth = 1.2) +
    geom_point(data = obs_biomass_ts, aes(x = Year, y = ObsBiomass, color = Species), size = 0.8) +
    geom_point(data = obs_biomass_ts, aes(x = Year, y = ObsBiomass), shape = 1, size = 0.8, colour = "black") +
    geom_ribbon(data = obs_biomass_ts, aes(x = Year, ymin = Lower, ymax = Upper), alpha = 0.2) +
    facet_wrap(~Species, scales = "free_y") +
    theme_bw() +
    theme(legend.position = "none", strip.text = element_text(face = "bold")) +
    labs(x = "Year", y = "Biomass", title = "Baseline Model: Biomass vs Observations")
}

# Basic plotting function for yield comparison
plot_baseline_yield <- function(sim_object, obs_yield_data) {
  require(ggplot2)
  
  p_yield <- plotYield(sim_object) + 
    facet_wrap(~Species, scales = "free_y") +
    geom_point(data = obs_yield_data, aes(x = Year, y = Yield, colour = Species), size = 1) +
    geom_point(data = obs_yield_data, aes(x = Year, y = Yield), shape = 1, size = 1, colour = "black") +
    geom_vline(aes(xintercept = 1961), linetype = "dashed") + 
    geom_vline(aes(xintercept = 2010), linetype = "dashed") +
    theme_bw() +
    theme(legend.position = "none", strip.text = element_text(face = "bold")) +
    labs(title = "Baseline Model: Yield vs Observations")
  
  return(p_yield)
}

# Assess baseline model
plot_baseline_biomass(sim_1841_2010_fishing_only, obs_biomass_data)
plot_baseline_yield(sim_1841_2010_fishing_only, yield_ts_tidy)
```


# Monte Carlo Uncertainty Analysis

This section implements a comprehensive Monte Carlo uncertainty analysis that varies:
1. **Catchability** - fishing gear efficiency 
2. **Gamma** - species search rates (foraging efficiency)
3. **Initial abundance** - starting biomass for all species

The analysis includes parameter distribution assessment and biomass-based rejection criteria.

## Step 1: Run Monte Carlo Simulations

```{r run-monte-carlo-analysis}
# Run enhanced uncertainty analysis
cat("Running Monte Carlo uncertainty analysis...\n")

# Set random seed for reproducibility
set.seed(123)

monte_carlo_results <- run_enhanced_combined_uncertainty_sims(
  params = params_1841_2010_climate_only,
  n_sims = 10,  # Small number for testing - increase for production runs
  catchability_sd = 0.4,  # 40% coefficient of variation for catchability
  abundance_sd = 0.8,     # 80% coefficient of variation for abundance
  gamma_sd = 0.5,         # 50% coefficient of variation for gamma (search rate)
  marine_mammal_species = c("minke whales", "orca", "sperm whales", "baleen whales"),
  effort_scen = combined_effort_array,
  enhanced_steady_tol = 0.002,
  enhanced_steady_t_max = 1000,
  spinup_years = 118,
  t_start = 1841,
  sim_years = 170,
  preserve_erepro = TRUE,
  max_erepro = 10,
  verbose = TRUE
)

cat("Monte Carlo analysis completed!\n")
cat("Successful simulations:", length(monte_carlo_results$simulations), "out of 10\n")
```

```{r run-monte-carlo-analysis}
# Run enhanced uncertainty analysis
cat("Running Monte Carlo uncertainty analysis...\n")

# Set random seed for reproducibility
set.seed(123)

monte_carlo_results <- run_enhanced_combined_uncertainty_sims(
  params = params_1841_2010_climate_only,
  n_sims = 50,  # Using 50 for testing - increase for production runs
  catchability_sd = 0.4,  # 40% coefficient of variation for catchability
  abundance_sd = 0.8,     # 80% coefficient of variation for abundance
  gamma_sd = 0.5,         # 50% coefficient of variation for gamma (search rate)
  marine_mammal_species = c("minke whales", "orca", "sperm whales", "baleen whales"),
  effort_scen = combined_effort_array,
  enhanced_steady_tol = 0.002,
  enhanced_steady_t_max = 1000,
  spinup_years = 118,
  t_start = 1841,
  sim_years = 170,
  preserve_erepro = TRUE,
  max_erepro = 10,
  verbose = TRUE
)

cat("Monte Carlo analysis completed!\n")
cat("Successful simulations:", monte_carlo_results$n_successful, "\n")
cat("Success rate:", round(monte_carlo_results$n_successful/monte_carlo_results$n_attempts*100, 1), "%\n")
```

## Step 2: Assess Parameter Distributions (Pre-Rejection)

Before applying any rejection criteria, let's examine the distributions of the parameters we varied.

```{r assess-parameter-distributions}
cat("=== PARAMETER DISTRIBUTION ASSESSMENT ===\n")

# Plot gamma parameter distributions
p_gamma_pre <- plot_parameter_distributions(
  monte_carlo_results$parameters, 
  param_type = "gamma",
  title_suffix = "(Pre-Rejection)"
)
print(p_gamma_pre)

# Plot catchability parameter distributions  
p_catch_pre <- plot_parameter_distributions(
  monte_carlo_results$parameters,
  param_type = "catchability", 
  title_suffix = "(Pre-Rejection)"
)
print(p_catch_pre)

# Plot abundance scaling distributions
p_abund_pre <- plot_parameter_distributions(
  monte_carlo_results$parameters,
  param_type = "abundance",
  title_suffix = "(Pre-Rejection)"
)
print(p_abund_pre)

# Print summary statistics
cat("\n=== PARAMETER SUMMARY STATISTICS (PRE-REJECTION) ===\n")

# Gamma statistics
gamma_summary <- do.call(rbind, lapply(monte_carlo_results$parameters, function(x) {
  data.frame(species = x$species_names, gamma_change = x$gamma_change)
}))
cat("Gamma change factors:\n")
print(gamma_summary %>% group_by(species) %>% 
      summarise(mean = mean(gamma_change), sd = sd(gamma_change), 
                min = min(gamma_change), max = max(gamma_change)))

# Abundance scaling statistics  
abund_summary <- do.call(rbind, lapply(monte_carlo_results$parameters, function(x) {
  data.frame(species = x$species_names, abundance_scaling = x$abundance_scaling)
}))
cat("\nAbundance scaling factors:\n")
print(abund_summary %>% group_by(species) %>% 
      summarise(mean = mean(abundance_scaling), sd = sd(abundance_scaling), 
                min = min(abundance_scaling), max = max(abundance_scaling)))
```



## Step 3: Apply Biomass-Based Rejection Criteria

Now we apply the strict rejection criteria: remove any parameter sets where the modeled biomass 
(averaged over 2000-2010) falls outside ±25% of the observed biomass (2010-2020) for ANY species.

```{r apply-biomass-rejection}
cat("=== APPLYING BIOMASS REJECTION CRITERIA ===\n")

# Apply rejection criteria
rejection_results <- apply_biomass_rejection_criteria(
  sim_results = monte_carlo_results$simulations,
  param_combinations = monte_carlo_results$parameters,
  obs_biomass_data = obs_biomass_data,
  tolerance = 0.25,  # ±25% tolerance
  comparison_years = 2000:2010,  # Use 2000-2010 from simulations
  verbose = TRUE
)

cat("\nRejection completed!\n")
cat("Acceptance rate:", round(rejection_results$acceptance_rate * 100, 1), "%\n")
```

## Step 4: Assess Parameter Distributions (Post-Rejection)

Let's examine how the rejection criteria affected our parameter distributions.

```{r assess-post-rejection-parameters}
if (rejection_results$n_accepted > 0) {
  cat("=== PARAMETER DISTRIBUTIONS AFTER REJECTION ===\n")
  
  # Plot gamma parameter distributions (post-rejection)
  p_gamma_post <- plot_parameter_distributions(
    rejection_results$accepted_parameters, 
    param_type = "gamma",
    title_suffix = "(Post-Rejection)"
  )
  print(p_gamma_post)
  
  # Plot catchability parameter distributions (post-rejection)
  p_catch_post <- plot_parameter_distributions(
    rejection_results$accepted_parameters,
    param_type = "catchability", 
    title_suffix = "(Post-Rejection)"
  )
  print(p_catch_post)
  
  # Plot abundance scaling distributions (post-rejection)
  p_abund_post <- plot_parameter_distributions(
    rejection_results$accepted_parameters,
    param_type = "abundance",
    title_suffix = "(Post-Rejection)"
  )
  print(p_abund_post)
  
  # Compare parameter distributions before and after rejection
  cat("\n=== PARAMETER COMPARISON: BEFORE vs AFTER REJECTION ===\n")
  
  # Gamma statistics (post-rejection)
  gamma_summary_post <- do.call(rbind, lapply(rejection_results$accepted_parameters, function(x) {
    data.frame(species = x$species_names, gamma_change = x$gamma_change)
  }))
  
  cat("Gamma change factors (Post-Rejection):\n")
  gamma_stats_post <- gamma_summary_post %>% group_by(species) %>% 
    summarise(mean = mean(gamma_change), sd = sd(gamma_change), 
              min = min(gamma_change), max = max(gamma_change))
  print(gamma_stats_post)
  
  # Abundance scaling statistics (post-rejection)
  abund_summary_post <- do.call(rbind, lapply(rejection_results$accepted_parameters, function(x) {
    data.frame(species = x$species_names, abundance_scaling = x$abundance_scaling)
  }))
  
  cat("\nAbundance scaling factors (Post-Rejection):\n")
  abund_stats_post <- abund_summary_post %>% group_by(species) %>% 
    summarise(mean = mean(abundance_scaling), sd = sd(abundance_scaling), 
              min = min(abundance_scaling), max = max(abundance_scaling))
  print(abund_stats_post)
  
} else {
  cat("No simulations passed the rejection criteria! Consider relaxing the tolerance.\n")
}
```

## Step 5: Compare Modeled vs Observed (Accepted Simulations Only)

Finally, let's compare the biomass and yield from our accepted simulations with observations.

```{r compare-accepted-simulations}
if (rejection_results$n_accepted > 0) {
  cat("=== COMPARING ACCEPTED SIMULATIONS WITH OBSERVATIONS ===\n")
  
  # Extract yield and biomass data from accepted simulations
  accepted_yield_data <- extract_simulation_data(rejection_results$accepted_simulations, "yield")
  accepted_biomass_data <- extract_simulation_data(rejection_results$accepted_simulations, "biomass")
  
  # Function to calculate uncertainty bounds
  calculate_uncertainty_bounds <- function(data, group_cols, value_col) {
    data %>%
      group_by(across(all_of(group_cols))) %>%
      summarise(
        mean = mean(.data[[value_col]], na.rm = TRUE),
        median = median(.data[[value_col]], na.rm = TRUE),
        q25 = quantile(.data[[value_col]], 0.25, na.rm = TRUE),
        q75 = quantile(.data[[value_col]], 0.75, na.rm = TRUE),
        q05 = quantile(.data[[value_col]], 0.05, na.rm = TRUE),
        q95 = quantile(.data[[value_col]], 0.95, na.rm = TRUE),
        min = min(.data[[value_col]], na.rm = TRUE),
        max = max(.data[[value_col]], na.rm = TRUE),
        .groups = 'drop'
      )
  }
  
  # Calculate yield uncertainty bounds
  yield_bounds <- calculate_uncertainty_bounds(accepted_yield_data, c("Year", "Species"), "Yield")
  
  # Calculate biomass uncertainty bounds  
  biomass_bounds <- calculate_uncertainty_bounds(accepted_biomass_data, c("Year", "Species"), "Biomass")
  
  # Plot yield comparison with uncertainty bounds
  p_yield_accepted <- ggplot() +
    # Add uncertainty ribbons
    geom_ribbon(data = yield_bounds, 
                aes(x = Year, ymin = q05, ymax = q95, fill = Species), alpha = 0.2) +
    geom_ribbon(data = yield_bounds, 
                aes(x = Year, ymin = q25, ymax = q75, fill = Species), alpha = 0.4) +
    # Add median line
    geom_line(data = yield_bounds, aes(x = Year, y = median, color = Species), linewidth = 1) +
    # Add observed data points
    geom_point(data = yield_ts_tidy, aes(x = Year, y = Yield, color = Species), size = 1.5) +
    geom_point(data = yield_ts_tidy, aes(x = Year, y = Yield), shape = 1, size = 1.5, colour = "black") +
    # Styling
    facet_wrap(~Species, scales = "free_y") +
    scale_fill_viridis_d(name = "Species") +
    scale_color_viridis_d(name = "Species") +
    theme_bw() +
    theme(legend.position = "none", strip.text = element_text(face = "bold")) +
    labs(
      title = "Accepted Simulations: Yield vs Observations",
      subtitle = paste("n =", rejection_results$n_accepted, "simulations passing biomass criteria"),
      x = "Year", y = "Yield"
    ) +
    geom_vline(xintercept = c(1961, 2010), linetype = "dashed", alpha = 0.5)
  
  print(p_yield_accepted)
  
  # Plot biomass comparison with uncertainty bounds
  p_biomass_accepted <- ggplot() +
    # Add uncertainty ribbons
    geom_ribbon(data = biomass_bounds, 
                aes(x = Year, ymin = q05, ymax = q95, fill = Species), alpha = 0.2) +
    geom_ribbon(data = biomass_bounds, 
                aes(x = Year, ymin = q25, ymax = q75, fill = Species), alpha = 0.4) +
    # Add median line
    geom_line(data = biomass_bounds, aes(x = Year, y = median, color = Species), linewidth = 1) +
    # Add observed data (2010-2020 period)
    geom_point(data = obs_biomass_data %>% crossing(Year = 2005:2010), 
               aes(x = Year, y = ObsBiomass, color = Species), size = 1.5) +
    geom_point(data = obs_biomass_data %>% crossing(Year = 2005:2010), 
               aes(x = Year, y = ObsBiomass), shape = 1, size = 1.5, colour = "black") +
    # Styling
    facet_wrap(~Species, scales = "free_y") +
    scale_fill_viridis_d(name = "Species") +
    scale_color_viridis_d(name = "Species") +
    theme_bw() +
    theme(legend.position = "none", strip.text = element_text(face = "bold")) +
    labs(
      title = "Accepted Simulations: Biomass vs Observations", 
      subtitle = paste("n =", rejection_results$n_accepted, "simulations passing biomass criteria"),
      x = "Year", y = "Biomass"
    ) +
    geom_vline(xintercept = c(1961, 2010), linetype = "dashed", alpha = 0.5)
  
  print(p_biomass_accepted)
  
  # Save results for future use (commented out for testing)
  # saveRDS(rejection_results, "biomass_filtered_uncertainty_results.RDS")
  cat("\n(Results not saved during testing - uncomment saveRDS line for production)\n")
  
} else {
  cat("Cannot create comparison plots - no simulations passed rejection criteria.\n")
  cat("Consider:\n")
  cat("1. Increasing tolerance (currently 25%)\n") 
  cat("2. Relaxing rejection criteria (allow some species to be outside bounds)\n")
  cat("3. Increasing number of simulations\n")
}
```
```

# Analysis Complete

This restructured analysis provides:

1. **Core Functions Integration**: All critical uncertainty analysis functions moved from Helper_Functions.R
2. **Parameter Distribution Assessment**: Comprehensive plotting and statistics for Monte Carlo parameters  
3. **Biomass Rejection Criteria**: Strict filtering based on 2000-2010 modeled vs 2010-2020 observed biomass (±25% tolerance)
4. **Post-Rejection Analysis**: Comparison of parameter distributions and model performance after filtering
5. **Uncertainty Quantification**: Full uncertainty bounds for accepted simulations

The analysis is now self-contained and transparent, with clear assessment of parameter space coverage and model performance relative to observations.
        }
      }
    }, error = function(e) {
      if (verbose && !is.null(sim_id)) {
        cat("Sim", sim_id, "steady-state failed:", e$message, "\nTrying again with parameter adjustment\n")
      }
      # Small parameter adjustment on error too
      if (attempt < max_attempts) {
        sp_params <- species_params(params)
        for (i in 1:nrow(sp_params)) {
          if (runif(1) < 0.2) {  # 20% chance to adjust each species
            adjustment <- runif(1, 0.98, 1.02)  # ±2% adjustment
            params@initial_n[i,] <- params@initial_n[i,] * adjustment
          }
        }
      }
    })
    
    attempt <- attempt + 1
  }
  
  # If all attempts failed, return original params with warning
  if (verbose && !is.null(sim_id)) {
    cat("Sim", sim_id, "steady-state failed after", max_attempts, "attempts. Using original parameters.\n")
  }
  return(list(params = params, steady_failed = TRUE))
}

# Enhanced spin-up function with multiple cycles and stability checks
enhanced_spinup <- function(params,
                           spinup_years = 118,
                           spinup_multiplier = 1,    # Can be 2, 3, or 4 for extended spinup
                           effort_scenario,
                           stability_tolerance = 0.10,  # Updated default to 10% based on test results
                           t_start = 1841,
                           final_years = 170,
                           verbose = FALSE,
                           sim_id = NULL) {
  
  total_spinup_years <- spinup_years * spinup_multiplier
  if (verbose && !is.null(sim_id)) {
    cat("Sim", sim_id, "running enhanced spinup for", total_spinup_years, "years (", 
        spinup_multiplier, "x base period)\n")
  }
  
  # Step 1: Initial spinup without fishing
  sim_spinup <- project(params, 
                       t_start = t_start, 
                       t_max = total_spinup_years,
                       effort = 0)
  
  # Step 2: Check biomass stability during spinup (NO FISHING PERIOD)
  # This is the key fix - we check stability on the spinup simulation, not the final one
  spinup_stability_check <- check_biomass_stability_enhanced(sim_spinup, 
                                                            tolerance = stability_tolerance, 
                                                            check_years = min(50, total_spinup_years),
                                                            verbose = verbose,
                                                            sim_id = sim_id)
  
  # Handle both return formats
  if (is.list(spinup_stability_check)) {
    is_stable <- spinup_stability_check$stable
  } else {
    is_stable <- spinup_stability_check
  }
  
  if (!is_stable) {
    if (verbose && !is.null(sim_id)) {
      cat("Sim", sim_id, "spinup not stable, extending by 50%\n")
    }
    # Extend spinup if not stable
    extended_spinup <- project(params,
                              initial_n = sim_spinup@n[total_spinup_years,,],
                              t_start = t_start + total_spinup_years,
                              t_max = round(total_spinup_years * 0.5),
                              effort = 0)
    
    # Check stability of extended spinup
    extended_stability_check <- check_biomass_stability_enhanced(extended_spinup, 
                                                                tolerance = stability_tolerance, 
                                                                check_years = min(30, round(total_spinup_years * 0.5)),
                                                                verbose = verbose,
                                                                sim_id = sim_id)
    
    if (is.list(extended_stability_check)) {
      is_stable <- extended_stability_check$stable
    } else {
      is_stable <- extended_stability_check
    }
    
    final_initial_n <- extended_spinup@n[round(total_spinup_years * 0.5),,]
  } else {
    if (verbose && !is.null(sim_id)) {
      cat("Sim", sim_id, "spinup achieved stability\n")
    }
    final_initial_n <- sim_spinup@n[total_spinup_years,,]
  }
  
  # Step 3: Run main simulation with fishing effort
  # This simulation will have variability due to fishing - that's expected and normal!
  sim_final <- project(params,
                      initial_n = final_initial_n,
                      t_start = t_start,
                      t_max = final_years,
                      effort = effort_scenario)
  
  return(list(
    simulation = sim_final,
    spinup_stable = is_stable,
    final_initial_n = final_initial_n,
    spinup_simulation = sim_spinup  # Return spinup sim for diagnostics if needed
  ))
}

# Enhanced model performance evaluation with multiple metrics
calculate_enhanced_performance_metrics <- function(sim_object,
                                                  yield_obs_data,
                                                  species_weights_df = species_weights,
                                                  species_list = NULL,
                                                  year_range = NULL) {
  
  # Get modeled yield 
  mod_yield <- getYield(sim_object)
  df_mod <- reshape2::melt(mod_yield)
  names(df_mod) <- c("Year", "Species", "Yield")
  
  # Filter by year range and species if specified
  if (!is.null(year_range)) {
    df_mod <- df_mod[df_mod$Year >= min(year_range) & df_mod$Year <= max(year_range), ]
    yield_obs_data <- yield_obs_data[yield_obs_data$Year >= min(year_range) &
                                     yield_obs_data$Year <= max(year_range), ]
  }
  
  if (!is.null(species_list)) {
    df_mod <- df_mod[df_mod$Species %in% species_list, ]
    yield_obs_data <- yield_obs_data[yield_obs_data$Species %in% species_list, ]
  }
  
  # Merge modeled and observed data
  comparison <- merge(df_mod, yield_obs_data,
                     by = c("Year", "Species"),
                     suffixes = c("_mod", "_obs"))
  
  # Calculate comprehensive metrics by species
  metrics_by_species <- comparison %>%
    group_by(Species) %>%
    summarise(
      RMSE = sqrt(mean((Yield_mod - Yield_obs)^2, na.rm = TRUE)),
      MAE = mean(abs(Yield_mod - Yield_obs), na.rm = TRUE),
      Mean_Obs = mean(Yield_obs, na.rm = TRUE),
      Mean_Mod = mean(Yield_mod, na.rm = TRUE),
      Bias = mean(Yield_mod - Yield_obs, na.rm = TRUE),
      Relative_Error = sqrt(mean((Yield_mod - Yield_obs)^2, na.rm = TRUE)) / mean(Yield_obs, na.rm = TRUE) * 100,
      Correlation = cor(Yield_mod, Yield_obs, use = "complete.obs"),
      MAPE = mean(abs((Yield_mod - Yield_obs) / Yield_obs), na.rm = TRUE) * 100,
      n_obs = n(),
      .groups = 'drop'
    )
  
  # Apply species weights and calculate weighted metrics
  metrics_weighted <- metrics_by_species %>%
    left_join(species_weights_df, by = "Species") %>%
    mutate(weight = ifelse(is.na(weight), 1, weight)) # Default weight of 1 for unspecified species
  
  # Calculate weighted overall metrics
  total_weighted_rmse <- weighted.mean(metrics_weighted$RMSE, 
                                       metrics_weighted$weight * metrics_weighted$n_obs, 
                                       na.rm = TRUE)
  
  total_weighted_mae <- weighted.mean(metrics_weighted$MAE,
                                      metrics_weighted$weight * metrics_weighted$n_obs,
                                      na.rm = TRUE)
  
  total_weighted_bias <- weighted.mean(metrics_weighted$Bias,
                                       metrics_weighted$weight * metrics_weighted$n_obs,
                                       na.rm = TRUE)
  
  avg_correlation <- mean(metrics_weighted$Correlation, na.rm = TRUE)
  
  # ============================================================================
  # 🎯 COMPOSITE SCORE CALCULATION (MAIN EVALUATION POINT)
  # ============================================================================
  # Option 1: Current weighted approach (emphasizes RMSE)
  # Calculate composite score (lower is better)
  # Combines multiple metrics with emphasis on RMSE
  composite_score_weighted <- (total_weighted_rmse * 0.4) +     # 40% weight: How far off predictions are (Root Mean Square Error)
                             (total_weighted_mae * 0.3) +       # 30% weight: Average absolute difference (Mean Absolute Error)  
                             (abs(total_weighted_bias) * 0.2) + # 20% weight: Systematic over/under-prediction (Bias)
                             ((1 - avg_correlation) * 0.1)      # 10% weight: How well patterns match (1 minus Correlation)
  
  # Option 2: Equal weighting approach (balanced contribution)
  composite_score_equal <- (total_weighted_rmse * 0.25) +       # 25% weight: How far off predictions are
                          (total_weighted_mae * 0.25) +         # 25% weight: Average absolute difference
                          (abs(total_weighted_bias) * 0.25) +   # 25% weight: Systematic over/under-prediction
                          ((1 - avg_correlation) * 0.25)        # 25% weight: How well patterns match
  
  # Default to weighted approach for backward compatibility
  composite_score <- composite_score_weighted
  # ============================================================================
  
  return(list(
    total_weighted_rmse = total_weighted_rmse,
    total_weighted_mae = total_weighted_mae,
    total_weighted_bias = total_weighted_bias,
    avg_correlation = avg_correlation,
    composite_score = composite_score,
    composite_score_weighted = composite_score_weighted,
    composite_score_equal = composite_score_equal,
    metrics_by_species = metrics_by_species,
    metrics_weighted = metrics_weighted,
    comparison_data = comparison
  ))
}

# Function to run a single simulation with enhanced protocols and diagnostics
run_enhanced_single_sim <- function(sim_id, params, catchability_sd, abundance_sd, gamma_sd,
                                   all_species_abundance, marine_mammal_species, effort_scen, 
                                   enhanced_steady_tol, enhanced_steady_t_max,
                                   spinup_years, spinup_multiplier,
                                   t_start, sim_years,
                                   preserve_erepro, max_erepro,
                                   species_weights_df,
                                   yield_obs_data,
                                   verbose = FALSE) {
  
  if (verbose) {
    cat("=== Starting enhanced simulation", sim_id, "===\n")
  }
  
  # Create copy of params
  rand_params <- params
  
  # Step 1: Randomize catchability for fished species
  gear_df <- gear_params(rand_params)
  original_catchability <- gear_df$catchability
  
  if(nrow(gear_df) > 0) {
    for(j in 1:nrow(gear_df)) {
      if(gear_df$catchability[j] > 0) {
        noise <- rnorm(1, mean = 0, sd = catchability_sd)
        gear_df$catchability[j] <- gear_df$catchability[j] * exp(noise)
        gear_df$catchability[j] <- min(1, max(0, gear_df$catchability[j]))
      }
    }
    gear_params(rand_params) <- gear_df
  }
  
  # Step 2: Randomize gamma (search rate) for all species
  sp_params <- species_params(rand_params)
  original_gamma <- sp_params$gamma
  
  # Apply gamma variation to all species
  for(j in 1:nrow(sp_params)) {
    gamma_noise <- rnorm(1, mean = 0, sd = gamma_sd)
    sp_params$gamma[j] <- sp_params$gamma[j] * exp(gamma_noise)
    # Ensure gamma stays within reasonable bounds (e.g., 0.01 to 1000)
    sp_params$gamma[j] <- min(1000, max(0.01, sp_params$gamma[j]))
  }
  
  species_params(rand_params) <- sp_params
  
  # Step 3: Randomize initial abundance
  if (all_species_abundance) {
    # NEW: Vary abundance for ALL species, not just marine mammals
    all_species <- sp_params$species
    scaling_factors <- exp(rnorm(length(all_species), mean = 0, sd = abundance_sd))
    
    for (k in seq_along(all_species)) {
      rand_params@initial_n[k,] <- rand_params@initial_n[k,] * scaling_factors[k]
    }
    
    if (verbose) {
      cat("Sim", sim_id, "parameter modifications:\n")
      cat("  - Catchability changes: ", paste(round(gear_df$catchability / original_catchability, 2), collapse = ", "), "\n")
      cat("  - Gamma changes: ", paste(round(sp_params$gamma / original_gamma, 2), collapse = ", "), "\n")
      cat("  - All species abundance scaling: ", paste(round(scaling_factors, 2), collapse = ", "), "\n")
    }
    
    # Store marine mammal scaling for tracking
    marine_mammal_indices <- which(sp_params$species %in% marine_mammal_species)
    marine_mammal_scaling <- scaling_factors[marine_mammal_indices]
    
  } else {
    # Original behavior: Only marine mammals
    species_indices <- which(sp_params$species %in% marine_mammal_species)
    scaling_factors <- exp(rnorm(length(species_indices), mean = 0, sd = abundance_sd))
    
    for (k in seq_along(species_indices)) {
      rand_params@initial_n[species_indices[k],] <-
        rand_params@initial_n[species_indices[k],] * scaling_factors[k]
    }
    
    if (verbose) {
      cat("Sim", sim_id, "parameter modifications:\n")
      cat("  - Catchability changes: ", paste(round(gear_df$catchability / original_catchability, 2), collapse = ", "), "\n")
      cat("  - Gamma changes: ", paste(round(sp_params$gamma / original_gamma, 2), collapse = ", "), "\n")
      cat("  - Marine mammal abundance scaling: ", paste(round(scaling_factors, 2), collapse = ", "), "\n")
    }
    
    marine_mammal_scaling <- scaling_factors
  }
  
  # Step 3: Enhanced steady state
  steady_result <- enhanced_steady_state(rand_params, 
                                       tol = enhanced_steady_tol,
                                       t_max = enhanced_steady_t_max,
                                       preserve_erepro = preserve_erepro,
                                       verbose = verbose,
                                       sim_id = sim_id)
  
  # Check if steady state failed
  if (is.list(steady_result) && !is.null(steady_result$steady_failed)) {
    return(list(
      success = FALSE,
      reason = "steady_state_failed",
      sim_id = sim_id,
      details = "Could not achieve steady state within max attempts"
    ))
  }
  
  rand_params <- steady_result
  
  # Check erepro if not preserving
  if (!preserve_erepro) {
    current_erepro <- species_params(rand_params)$erepro
    if (any(current_erepro > max_erepro, na.rm = TRUE)) {
      return(list(
        success = FALSE,
        reason = "erepro_exceeded",
        sim_id = sim_id,
        details = paste("Max erepro:", round(max(current_erepro, na.rm = TRUE), 3))
      ))
    }
  }
  
  # Step 4: Enhanced spinup and main simulation
  sim_result <- enhanced_spinup(rand_params,
                               spinup_years = spinup_years,
                               spinup_multiplier = spinup_multiplier,
                               effort_scenario = effort_scen,
                               stability_tolerance = 0.10,  # Use 10% CV threshold for spinup stability
                               t_start = t_start,
                               final_years = sim_years,
                               verbose = verbose,
                               sim_id = sim_id)
  
  # Step 5: Check if spinup achieved stability - this is the only stability check we need!
  # The final simulation with fishing is SUPPOSED to have variability
  if (!sim_result$spinup_stable) {
    return(list(
      success = FALSE,
      reason = "spinup_instability",
      sim_id = sim_id,
      details = "Spinup period did not achieve biomass stability (>10% CV)"
    ))
  }
  
  if (verbose && !is.null(sim_id)) {
    cat("Sim", sim_id, "spinup was stable, simulation completed successfully\n")
  }
  
  # Step 6: Calculate enhanced performance metrics
  performance <- calculate_enhanced_performance_metrics(sim_result$simulation,
                                                       yield_obs_data,
                                                       species_weights_df)
  
  if (verbose) {
    cat("Sim", sim_id, "completed successfully. Composite score:", 
        round(performance$composite_score, 4), "\n")
  }
  
  return(list(
    success = TRUE,
    sim_id = sim_id,
    simulation = sim_result$simulation,
    performance = performance,
    spinup_stable = sim_result$spinup_stable,
    params_info = list(
      sim_id = sim_id,
      catchability = gear_df$catchability,
      catchability_change = gear_df$catchability / original_catchability,
      gamma_values = sp_params$gamma,                               # NEW: Store gamma values
      gamma_change = sp_params$gamma / original_gamma,              # NEW: Store gamma changes
      marine_mammal_scaling = marine_mammal_scaling,                # Updated variable name
      all_species_scaling = if(all_species_abundance) scaling_factors else NULL,  # NEW: All species scaling if used
      marine_mammal_species = marine_mammal_species,
      all_species_abundance = all_species_abundance,                # NEW: Track which approach was used
      erepro_values = species_params(rand_params)$erepro
    )
  ))
}

# Main function to run enhanced combined uncertainty simulations
run_enhanced_combined_uncertainty_sims <- function(params,
                                                   n_sims = 10,
                                                   catchability_sd = 0.1,
                                                   abundance_sd = 0.2,
                                                   gamma_sd = 0.3,                    # NEW: Standard deviation for gamma variation
                                                   all_species_abundance = TRUE,      # NEW: Vary abundance for all species
                                                   marine_mammal_species = NULL,
                                                   effort_scen,
                                                   enhanced_steady_tol = 0.002,      # Stricter tolerance
                                                   enhanced_steady_t_max = 1000,     # Minimum 1000 years
                                                   spinup_years = 118,
                                                   spinup_multiplier = 2,             # 2x base spinup by default
                                                   t_start = 1841,
                                                   sim_years = 170,
                                                   preserve_erepro = TRUE,
                                                   max_erepro = 1,
                                                   species_weights_df = species_weights,
                                                   yield_obs_data,
                                                   n_cores = NULL,
                                                   parallel = TRUE) {
  
  # Default marine mammal species if not specified
  if (is.null(marine_mammal_species)) {
    marine_mammal_species <- c("minke whales", "orca", "sperm whales", "baleen whales")
  }
  
  # Determine number of cores to use
  if (is.null(n_cores)) {
    n_cores <- detectCores() - 1
  }
  n_cores <- min(n_cores, detectCores())
  
  # Progress message
  cat("=== ENHANCED COMBINED UNCERTAINTY ANALYSIS ===\n")
  cat("Running", n_sims, "simulations with enhanced protocols:\n")
  cat("  - Steady-state tolerance:", enhanced_steady_tol, "\n")
  cat("  - Minimum steady-state time:", enhanced_steady_t_max, "years\n")
  cat("  - Spinup multiplier:", spinup_multiplier, "x (", spinup_years * spinup_multiplier, "years total)\n")
  cat("  - Catchability SD:", catchability_sd, "\n")
  cat("  - Abundance SD:", abundance_sd, "\n")
  cat("  - Gamma SD:", gamma_sd, "(NEW)\n")
  cat("  - All species abundance variation:", all_species_abundance, "(NEW)\n")
  cat("  - Whale species weighting: 3-10x higher than fish/krill\n")
  cat("  - Performance metrics: RMSE, MAE, Bias, Correlation, Composite Score\n")
  
  if (preserve_erepro) {
    cat("  - Preserving erepro values during steady state\n")
  } else {
    cat("  - Will skip simulations with erepro >", max_erepro, "\n")
  }
  
  if (parallel && n_cores > 1) {
    cat("  - Using parallel processing with", n_cores, "cores\n")
    
    # Set up parallel cluster
    cl <- makeCluster(n_cores)
    registerDoParallel(cl)
    
    # Export necessary objects to cluster
    clusterExport(cl, c("run_enhanced_single_sim", "enhanced_steady_state", 
                       "enhanced_spinup", "check_biomass_stability_enhanced",
                       "calculate_enhanced_performance_metrics", "calculate_yield_rmse",
                       "species_weights_df"), 
                  envir = environment())
    
    # Export yield data for any legacy references in cluster functions
    clusterExport(cl, "yield_obs_data", envir = environment())
    
    # Also create yield_ts_tidy alias in cluster environment for backward compatibility
    clusterEvalQ(cl, {
      yield_ts_tidy <- yield_obs_data
    })
    
    # Run simulations in parallel
    n_attempts <- ifelse(preserve_erepro, n_sims, ceiling(n_sims * 2))
    
    cat("  - Running", n_attempts, "initial attempts...\n")
    
    results <- foreach(i = 1:n_attempts,
                      .packages = c("mizer", "therMizer", "dplyr"),
                      .errorhandling = "pass") %dopar% {
      run_enhanced_single_sim(i, params, catchability_sd, abundance_sd, gamma_sd,
                             all_species_abundance, marine_mammal_species, effort_scen, 
                             enhanced_steady_tol, enhanced_steady_t_max,
                             spinup_years, spinup_multiplier,
                             t_start, sim_years,
                             preserve_erepro, max_erepro,
                             species_weights_df, yield_obs_data)
    }
    
    # Stop cluster
    stopCluster(cl)
    
  } else {
    # Sequential processing
    cat("  - Using sequential processing\n")
    
    n_attempts <- ifelse(preserve_erepro, n_sims, ceiling(n_sims * 2))
    results <- list()
    
    for(i in 1:n_attempts) {
      if(i %% 5 == 0) cat("    Processing simulation", i, "of", n_attempts, "\n")
      
      results[[i]] <- run_enhanced_single_sim(i, params, catchability_sd, abundance_sd, gamma_sd,
                                             all_species_abundance, marine_mammal_species, effort_scen, 
                                             enhanced_steady_tol, enhanced_steady_t_max,
                                             spinup_years, spinup_multiplier,
                                             t_start, sim_years,
                                             preserve_erepro, max_erepro,
                                             species_weights_df, yield_obs_data,
                                             verbose = (i <= 3))  # Verbose for first 3 simulations
    }
  }
  
  # Process results
  sim_results <- list()
  param_combinations <- list()
  performance_results <- list()
  failed_sims <- list()
  successful_count <- 0
  
  for(res in results) {
    if(!is.null(res) && !inherits(res, "error")) {
      if(res$success) {
        successful_count <- successful_count + 1
        if(successful_count <= n_sims) {
          sim_results[[successful_count]] <- res$simulation
          param_combinations[[successful_count]] <- res$params_info
          performance_results[[successful_count]] <- res$performance
        }
      } else {
        failed_sims[[length(failed_sims) + 1]] <- res
      }
    }
  }
  
  # Final summary
  cat("\n=== ENHANCED SIMULATION RESULTS ===\n")
  cat("Completed", min(successful_count, n_sims), "successful simulations\n")
  if (length(failed_sims) > 0) {
    failure_reasons <- table(sapply(failed_sims, function(x) x$reason))
    cat("Failed simulations by reason:\n")
    print(failure_reasons)
    
    # Show detailed failure information for first few failures
    cat("\nDetailed failure analysis (first 5 failures):\n")
    for (i in 1:min(5, length(failed_sims))) {
      fail_info <- failed_sims[[i]]
      cat("  Sim", fail_info$sim_id, "- Reason:", fail_info$reason)
      if (!is.null(fail_info$details)) {
        cat(" - Details:", fail_info$details)
      }
      cat("\n")
    }
  }
  
  # Calculate summary statistics of performance
  if (length(performance_results) > 0) {
    composite_scores <- sapply(performance_results, function(x) x$composite_score)
    rmse_scores <- sapply(performance_results, function(x) x$total_weighted_rmse)
    
    cat("\nPerformance Summary:\n")
    cat("  Composite Score - Mean:", round(mean(composite_scores), 4), 
        "Range:", round(min(composite_scores), 4), "-", round(max(composite_scores), 4), "\n")
    cat("  Weighted RMSE - Mean:", round(mean(rmse_scores), 4), 
        "Range:", round(min(rmse_scores), 4), "-", round(max(rmse_scores), 4), "\n")
  }
  
  # Trim to requested number of simulations
  if(length(sim_results) > n_sims) {
    sim_results <- sim_results[1:n_sims]
    param_combinations <- param_combinations[1:n_sims]
    performance_results <- performance_results[1:n_sims]
  }
  
  return(list(
    simulations = sim_results,
    parameters = param_combinations,
    performance = performance_results,
    failed = failed_sims,
    n_successful = length(sim_results),
    n_attempts = length(results)
  ))
}

# Function to find optimal parameters based on enhanced metrics
find_optimal_parameters_enhanced <- function(sim_results_list, metric = "composite_score") {
  
  # Extract performance metrics
  performance_data <- sim_results_list$performance
  
  if (length(performance_data) == 0) {
    stop("No performance data available")
  }
  
  # Debug: Show available metrics from first performance entry
  if (length(performance_data) > 0) {
    available_metrics <- names(performance_data[[1]])
    cat("Available metrics:", paste(available_metrics, collapse = ", "), "\n")
  }
  
  # Get the specified metric values
  metric_values <- sapply(performance_data, function(x) {
    if (metric %in% names(x)) {
      return(x[[metric]])
    } else {
      return(NA)
    }
  })
  
  # Debug: Show how many valid values we found
  valid_values <- sum(!is.na(metric_values))
  cat("Found", valid_values, "valid values for metric:", metric, "\n")
  
  if (valid_values == 0) {
    stop("Could not find any valid values for metric: ", metric)
  }
  
  # Find optimal (minimum for most metrics)
  if (metric == "avg_correlation") {
    optimal_idx <- which.max(metric_values)  # Higher correlation is better
  } else {
    optimal_idx <- which.min(metric_values)  # Lower is better for RMSE, MAE, bias, composite
  }
  
  if (length(optimal_idx) == 0 || is.na(optimal_idx) || is.na(metric_values[optimal_idx])) {
    cat("Debug info:\n")
    cat("  metric_values length:", length(metric_values), "\n")
    cat("  valid values:", sum(!is.na(metric_values)), "\n")
    cat("  metric_values range:", range(metric_values, na.rm = TRUE), "\n")
    stop("Could not find optimal simulation for metric: ", metric)
  }
  
  # Create summary statistics
  metric_stats <- data.frame(
    mean = mean(metric_values, na.rm = TRUE),
    median = median(metric_values, na.rm = TRUE),
    sd = sd(metric_values, na.rm = TRUE),
    min = min(metric_values, na.rm = TRUE),
    max = max(metric_values, na.rm = TRUE),
    q25 = quantile(metric_values, 0.25, na.rm = TRUE),
    q75 = quantile(metric_values, 0.75, na.rm = TRUE)
  )
  
  cat("\n=== OPTIMAL PARAMETER SELECTION ===\n")
  cat("Optimization metric:", metric, "\n")
  cat("Optimal simulation:", optimal_idx, "with", metric, "=", metric_values[optimal_idx], "\n")
  cat("\nMetric Statistics:\n")
  print(metric_stats)
  
  # Extract detailed performance for optimal simulation
  optimal_performance <- performance_data[[optimal_idx]]
  
  cat("\n=== OPTIMAL SIMULATION PERFORMANCE ===\n")
  cat("Weighted RMSE:", round(optimal_performance$total_weighted_rmse, 4), "\n")
  cat("Weighted MAE:", round(optimal_performance$total_weighted_mae, 4), "\n")
  cat("Weighted Bias:", round(optimal_performance$total_weighted_bias, 4), "\n")
  cat("Average Correlation:", round(optimal_performance$avg_correlation, 4), "\n")
  cat("Composite Score:", round(optimal_performance$composite_score, 4), "\n")
  
  cat("\nSpecies-specific performance:\n")
  print(optimal_performance$metrics_by_species)
  
  return(list(
    optimal_index = optimal_idx,
    optimal_metric_value = metric_values[optimal_idx],
    optimal_sim = sim_results_list$simulations[[optimal_idx]],
    optimal_params = sim_results_list$parameters[[optimal_idx]],
    optimal_performance = optimal_performance,
    all_metric_values = metric_values,
    metric_stats = metric_stats,
    metric_used = metric
  ))
}

# Function to plot RMSE distribution from uncertainty analysis
plot_rmse_distribution <- function(rmse_values,
                                  optimal_idx = NULL,
                                  plot_title = "RMSE Distribution from Combined Parameter Uncertainty") {
  
  require(ggplot2)
  
  df <- data.frame(
    Simulation = 1:length(rmse_values),
    RMSE = rmse_values
  )
  
  # Calculate statistics for annotation
  mean_rmse <- mean(rmse_values)
  median_rmse <- median(rmse_values)
  
  p <- ggplot(df, aes(x = RMSE)) +
    geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7, color = "black") +
    theme_bw() +
    labs(
      title = plot_title,
      subtitle = paste("n =", length(rmse_values), "simulations"),
      x = "RMSE (Root Mean Square Error)",
      y = "Frequency"
    ) +
    # Add mean line
    geom_vline(xintercept = mean_rmse,
              color = "blue", linetype = "dotted", linewidth = 1) +
    # Add median line
    geom_vline(xintercept = median_rmse,
              color = "green", linetype = "dotted", linewidth = 1)
  
  # Add vertical line for optimal value if provided
  if (!is.null(optimal_idx)) {
    p <- p +
      geom_vline(xintercept = rmse_values[optimal_idx],
                color = "red", linetype = "dashed", linewidth = 1) +
      annotate("text",
              x = rmse_values[optimal_idx],
              y = max(table(cut(rmse_values, 30))) * 0.9,
              label = paste("Optimal RMSE =", round(rmse_values[optimal_idx], 2)),
              hjust = -0.1, color = "red")
  }
  
  # Add legend for lines
  p <- p +
    annotate("text", x = mean_rmse, y = max(table(cut(rmse_values, 30))) * 0.8,
            label = paste("Mean =", round(mean_rmse, 2)),
            hjust = -0.1, color = "blue", size = 3) +
    annotate("text", x = median_rmse, y = max(table(cut(rmse_values, 30))) * 0.7,
            label = paste("Median =", round(median_rmse, 2)),
            hjust = -0.1, color = "green", size = 3)
  
  return(p)
}

# Function to visualize parameter combinations
plot_parameter_exploration <- function(param_combinations, rmse_values) {
  
  require(ggplot2)
  require(tidyr)
  require(dplyr)
  
  # Extract catchability and marine mammal scaling
  df_params <- data.frame(
    sim_id = 1:length(param_combinations),
    rmse = rmse_values
  )
  
  # Add catchability info (assuming single gear for simplicity)
  df_params$catchability <- sapply(param_combinations, function(x) mean(x$catchability))
  
  # Add marine mammal scaling (mean across all marine mammals)
  df_params$mammal_scaling <- sapply(param_combinations, function(x) mean(x$marine_mammal_scaling))
  
  # Create scatter plot
  p1 <- ggplot(df_params, aes(x = catchability, y = mammal_scaling, color = rmse)) +
    geom_point(size = 3, alpha = 0.7) +
    scale_color_gradient(low = "blue", high = "red", name = "RMSE") +
    theme_bw() +
    labs(
      title = "Parameter Space Exploration",
      x = "Mean Catchability",
      y = "Mean Marine Mammal Scaling Factor"
    )
  
  # Create individual parameter distributions
  p2 <- ggplot(df_params, aes(x = catchability)) +
    geom_histogram(bins = 20, fill = "steelblue", alpha = 0.7) +
    theme_bw() +
    labs(title = "Catchability Distribution", x = "Catchability", y = "Count")
  
  p3 <- ggplot(df_params, aes(x = mammal_scaling)) +
    geom_histogram(bins = 20, fill = "darkgreen", alpha = 0.7) +
    theme_bw() +
    labs(title = "Marine Mammal Scaling Distribution", x = "Scaling Factor", y = "Count")
  
  # Combine plots
  require(gridExtra)
  combined_plot <- grid.arrange(p1, p2, p3, ncol = 2, nrow = 2)
  
  return(combined_plot)
}
```

## Enhanced Uncertainty Analysis with Gamma and All-Species Abundance Variation

The enhanced uncertainty analysis incorporates several key improvements:

1. **Whale Species Weighting**: Marine mammals weighted 3-10x higher than fish/krill species
2. **Multiple Performance Metrics**: RMSE, MAE, bias, correlation, and composite scores
3. **Enhanced Steady-State Protocols**: Stricter tolerance (≤0.002) and longer convergence times (≥1000 years)
4. **Corrected Stability Validation**: Stability checked **only during spinup period** (no fishing), not during fished simulations
5. **Realistic CV Thresholds**: 10% CV threshold for spinup stability, allowing natural variability during fishing periods
6. **NEW: Gamma (Search Rate) Variation**: Randomizes foraging efficiency for all species using `gamma_sd` parameter
7. **NEW: All-Species Abundance Variation**: Option to vary initial abundance for all species with `all_species_abundance = TRUE`

### Key New Parameters:

- **`gamma_sd`**: Standard deviation for gamma (search rate) variation across all species
- **`all_species_abundance`**: Boolean flag to enable abundance variation for all species (default: TRUE)

The analysis now explores a much broader parameter space, including:
- Catchability uncertainty for fished species
- Gamma (foraging efficiency) uncertainty for all species  
- Initial abundance uncertainty for all species (or just marine mammals)

```{r load-enhanced-tests}
# Load diagnostic and testing functions from separate script
source("Enhanced_Stability_Tests.R")
```

### Run Enhanced Uncertainty Analysis

```{r run-enhanced-uncertainty-analysis}
# Ensure yield_ts_tidy is available for any functions that reference it directly
if (!exists("yield_ts_tidy")) {
  yield_ts_tidy <- readRDS("yield_observed_timeseries_tidy.RDS")
}
# Run enhanced uncertainty analysis with corrected stability logic
cat("Running Enhanced Uncertainty Analysis with corrected approach...\n\n")

# Validate approach first by running a quick test
cat("Step 1: Validating spinup-only stability approach...\n")
validation_test <- test_spinup_stability_approach(
  params_1841_2010_climate_only, 
  combined_effort_array, 
  n_tests = 1
)

if (validation_test$success_rate >= 80) {
  cat("✓ Approach validated! Success rate:", round(validation_test$success_rate, 1), "%\n")
  cat("✓ Spinup periods achieve <10% CV, fished periods have expected variability\n\n")
  
  cat("Step 2: Running full enhanced uncertainty analysis...\n")
  
  # Run the full enhanced uncertainty analysis
  enhanced_uncertainty_results <- run_enhanced_combined_uncertainty_sims(
    params = params_1841_2010_climate_only,
    n_sims = 3000,  # Full simulation set
    catchability_sd = 0.4,  # 30% CV for catchability
    abundance_sd = 0.8,     # 50% CV for abundance (now all species by default)
    gamma_sd = 0.5,         # NEW: 40% CV for gamma (search rate) variation
    all_species_abundance = TRUE,  # NEW: Vary abundance for all species
    marine_mammal_species = c("minke whales", "orca", "sperm whales", "baleen whales"),
    effort_scen = combined_effort_array,
    enhanced_steady_tol = 0.002,
    enhanced_steady_t_max = 1000,
    spinup_years = 118,
    spinup_multiplier = 2,  # Extended spinup for stability
    t_start = 1841,
    sim_years = 170,
    preserve_erepro = TRUE,
    max_erepro = 10,
    yield_obs_data = yield_ts_tidy
  )
  
  # Save results
  saveRDS(enhanced_uncertainty_results, "enhanced_uncertainty_final_results.RDS")
  
  # Calculate success rate
  success_rate <- (enhanced_uncertainty_results$n_successful / enhanced_uncertainty_results$n_attempts) * 100
  cat("\n=== ENHANCED UNCERTAINTY ANALYSIS RESULTS ===\n")
  cat("Success rate:", round(success_rate, 1), "%\n")
  cat("Successful simulations:", enhanced_uncertainty_results$n_successful, "out of", enhanced_uncertainty_results$n_attempts, "attempts\n")
  
  if (success_rate >= 70) {
    successful_sims <- enhanced_uncertainty_results$simulations
    cat("✓ SUCCESS! Enhanced uncertainty analysis completed successfully\n")
    cat("Number of successful simulations:", length(successful_sims), "\n")
    cat("Ready for uncertainty plotting with geom_ribbon\n")
    
    # Store successful simulations for plotting
    saveRDS(successful_sims, "enhanced_simulations_for_plotting.RDS")
    
    # NEW: Analyze the enhanced parameter variations
    cat("\n=== ANALYZING ENHANCED PARAMETER VARIATIONS ===\n")
    cat("Analyzing variations in:\n")
    cat("  - Catchability (fishing gear efficiency)\n")
    cat("  - Gamma (search rates for all species)\n")
    cat("  - Initial abundance (all species vs. marine mammals only)\n\n")
    
    enhanced_param_analysis <- analyze_enhanced_parameter_variations(
      enhanced_uncertainty_results, 
      top_percent = 0.1
    )
    
    cat("✓ Enhanced parameter analysis completed\n")
    cat("✓ Plots saved as 'enhanced_parameter_analysis.png'\n")
    cat("✓ Analysis data available in enhanced_param_analysis object\n\n")
    
  } else {
    cat("⚠ Moderate success rate. May need further optimization.\n")
  }
  
} else {
  cat("✗ Validation failed. Need to investigate stability approach further.\n")
  cat("Success rate:", round(validation_test$success_rate, 1), "%\n")
}
```

### Enhanced Uncertainty Plotting Functions

Now that we have successful enhanced simulations, we can create plots with uncertainty ribbons:

```{r enhanced-plotting-functions}
# Enhanced plotting functions with uncertainty ribbons
# Load successful enhanced simulations
if (file.exists("enhanced_simulations_for_plotting.RDS")) {
  enhanced_sims <- readRDS("enhanced_simulations_for_plotting.RDS")
  cat("Loaded", length(enhanced_sims), "successful enhanced simulations for plotting\n")
  
  # Function to calculate uncertainty ribbons from multiple simulations
  calculate_uncertainty_ribbons <- function(sim_list, metric_func, years = 1841:2010) {
    
    # Extract metric from each simulation
    metric_data <- lapply(sim_list, function(sim) {
      metric_values <- metric_func(sim)
      
      # Ensure we have the right years
      if (is.matrix(metric_values) || is.data.frame(metric_values)) {
        metric_df <- as.data.frame(metric_values)
        metric_df$year <- as.numeric(rownames(metric_df))
      } else {
        metric_df <- data.frame(
          value = metric_values,
          year = years[1:length(metric_values)]
        )
      }
      
      return(metric_df)
    })
    
    # Combine all simulations and calculate quantiles
    all_data <- do.call(rbind, lapply(1:length(metric_data), function(i) {
      df <- metric_data[[i]]
      df$sim_id <- i
      return(df)
    }))
    
    if (ncol(all_data) > 3) {  # Multiple species/groups
      # For multiple columns, calculate uncertainty for each
      species_cols <- names(all_data)[!names(all_data) %in% c("year", "sim_id")]
      
      uncertainty_data <- lapply(species_cols, function(species) {
        species_data <- all_data %>%
          select(year, sim_id, value = all_of(species)) %>%
          group_by(year) %>%
          summarise(
            median = median(value, na.rm = TRUE),
            q25 = quantile(value, 0.25, na.rm = TRUE),
            q75 = quantile(value, 0.75, na.rm = TRUE),
            q05 = quantile(value, 0.05, na.rm = TRUE),
            q95 = quantile(value, 0.95, na.rm = TRUE),
            mean = mean(value, na.rm = TRUE),
            .groups = 'drop'
          ) %>%
          mutate(species = species)
        
        return(species_data)
      })
      
      return(do.call(rbind, uncertainty_data))
      
    } else {  # Single metric
      uncertainty_data <- all_data %>%
        group_by(year) %>%
        summarise(
          median = median(value, na.rm = TRUE),
          q25 = quantile(value, 0.25, na.rm = TRUE),
          q75 = quantile(value, 0.75, na.rm = TRUE),
          q05 = quantile(value, 0.05, na.rm = TRUE),
          q95 = quantile(value, 0.95, na.rm = TRUE),
          mean = mean(value, na.rm = TRUE),
          .groups = 'drop'
        )
      
      return(uncertainty_data)
    }
  }
  
  # Enhanced yield plotting function with uncertainty ribbons
  plot_yield_uncertainty <- function(sim_list, observed_data = NULL, title = "Yield with Uncertainty") {
    
    # Calculate yield uncertainty ribbons
    yield_uncertainty <- calculate_uncertainty_ribbons(sim_list, getYield)
    
    # Create base plot
    p <- ggplot(yield_uncertainty, aes(x = year)) +
      geom_ribbon(aes(ymin = q05, ymax = q95, fill = species), alpha = 0.2) +
      geom_ribbon(aes(ymin = q25, ymax = q75, fill = species), alpha = 0.4) +
      geom_line(aes(y = median, color = species), size = 1) +
      scale_fill_viridis_d(name = "Species") +
      scale_color_viridis_d(name = "Species") +
      labs(
        title = title,
        x = "Year",
        y = "Yield",
        subtitle = "Ribbons show 50% (dark) and 90% (light) confidence intervals"
      ) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(size = 12)
      )
    
    # Add observed data if provided
    if (!is.null(observed_data)) {
      # Ensure column names are consistent (convert to lowercase if needed)
      obs_data_plot <- observed_data
      names(obs_data_plot) <- tolower(names(obs_data_plot))
      
      p <- p + geom_point(data = obs_data_plot, 
                         aes(x = year, y = yield, color = species), 
                         size = 2, alpha = 0.7)
    }
    
    return(p)
  }
  
  # Enhanced biomass plotting function with uncertainty ribbons
  plot_biomass_uncertainty <- function(sim_list, title = "Biomass with Uncertainty") {
    
    # Calculate biomass uncertainty ribbons
    biomass_uncertainty <- calculate_uncertainty_ribbons(sim_list, getBiomass)
    
    # Create plot
    p <- ggplot(biomass_uncertainty, aes(x = year)) +
      geom_ribbon(aes(ymin = q05, ymax = q95, fill = species), alpha = 0.2) +
      geom_ribbon(aes(ymin = q25, ymax = q75, fill = species), alpha = 0.4) +
      geom_line(aes(y = median, color = species), size = 1) +
      scale_fill_viridis_d(name = "Species") +
      scale_color_viridis_d(name = "Species") +
      scale_y_log10() +
      labs(
        title = title,
        x = "Year", 
        y = "Biomass",
        subtitle = "Ribbons show 50% (dark) and 90% (light) confidence intervals"
      ) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(size = 12)
      )
    
    return(p)
  }
  
} else {
  cat("Enhanced simulations not found. Please run the enhanced uncertainty analysis first.\n")
}
```

### Load and Analyze Enhanced Uncertainty Results

```{r load-and-analyze-results}
# Load enhanced uncertainty results if not already in environment
if (!exists("enhanced_uncertainty_results")) {
  if (file.exists("enhanced_uncertainty_final_results.RDS")) {
    cat("Loading enhanced uncertainty results from file...\n")
    enhanced_uncertainty_results <- readRDS("enhanced_uncertainty_final_results.RDS")
  } else {
    stop("Enhanced uncertainty results not found. Please run the enhanced uncertainty analysis first.")
  }
}

# Check if we have valid results
if (exists("enhanced_uncertainty_results") && length(enhanced_uncertainty_results$performance) > 0) {
  cat("Enhanced uncertainty results loaded successfully!\n")
  cat("Number of simulations:", length(enhanced_uncertainty_results$simulations), "\n")
  cat("Number of performance records:", length(enhanced_uncertainty_results$performance), "\n")
  
  # Check what metrics are available in the saved data
  available_metrics <- names(enhanced_uncertainty_results$performance[[1]])
  cat("Available metrics:", paste(available_metrics, collapse = ", "), "\n")
  
  # Calculate missing composite score variants if not present
  if (!"composite_score_weighted" %in% available_metrics) {
    cat("Adding missing composite score variants to existing data...\n")
    cat("Converting gram-scale metrics for composite scores...\n")
    
    # Add the missing metrics to each performance record
    for (i in 1:length(enhanced_uncertainty_results$performance)) {
      perf <- enhanced_uncertainty_results$performance[[i]]
      
      # Use original gram-scale metrics for composite score calculation
      rmse_grams <- perf$total_weighted_rmse
      mae_grams <- perf$total_weighted_mae
      bias_grams <- perf$total_weighted_bias
      correlation <- perf$avg_correlation
      
      # Calculate weighted composite (40-30-20-10) using original gram-scale values
      perf$composite_score_weighted <- (rmse_grams * 0.4) +
                                       (mae_grams * 0.3) +
                                       (abs(bias_grams) * 0.2) +
                                       ((1 - correlation) * 0.1)
      
      # Calculate equal composite (25-25-25-25) using original gram-scale values
      perf$composite_score_equal <- (rmse_grams * 0.25) +
                                    (mae_grams * 0.25) +
                                    (abs(bias_grams) * 0.25) +
                                    ((1 - correlation) * 0.25)
      
      # Update the performance record
      enhanced_uncertainty_results$performance[[i]] <- perf
    }
    cat("Successfully added composite score variants!\n")
  }
  
  # Find optimal parameters based on different metrics
  # 1. Composite score - WEIGHTED approach (current: 40-30-20-10)
  cat("\n=== FINDING OPTIMAL PARAMETERS (WEIGHTED COMPOSITE) ===\n")
  optimal_results_composite_weighted <- find_optimal_parameters_enhanced(
    sim_results_list = enhanced_uncertainty_results,
    metric = "composite_score_weighted"
  )
  
  # 2. Composite score - EQUAL approach (25-25-25-25)
  cat("\n=== FINDING OPTIMAL PARAMETERS (EQUAL COMPOSITE) ===\n")
  optimal_results_composite_equal <- find_optimal_parameters_enhanced(
    sim_results_list = enhanced_uncertainty_results,
    metric = "composite_score_equal"
  )
  
  # 3. Traditional composite score (default)
  cat("\n=== FINDING OPTIMAL PARAMETERS (DEFAULT COMPOSITE) ===\n")
  optimal_results_composite <- find_optimal_parameters_enhanced(
    sim_results_list = enhanced_uncertainty_results,
    metric = "composite_score"
  )
  
  # 4. Weighted RMSE only (for comparison)
  cat("\n=== FINDING OPTIMAL PARAMETERS (RMSE ONLY) ===\n")
  optimal_results_rmse <- find_optimal_parameters_enhanced(
    sim_results_list = enhanced_uncertainty_results,
    metric = "total_weighted_rmse"
  )
  
} else {
  cat("No enhanced uncertainty results available for analysis.\n")
}
```

### Compare Different Optimization Approaches

```{r compare-optimization-approaches}
# Compare performance across different optimization metrics
if (exists("optimal_results_composite_weighted") && exists("optimal_results_composite_equal")) {
  
  cat("\n=== DETAILED COMPARISON: WEIGHTED vs EQUAL WEIGHTING ===\n")
  weighted_perf <- optimal_results_composite_weighted$optimal_performance
  equal_perf <- optimal_results_composite_equal$optimal_performance
  
  cat("WEIGHTED approach (40-30-20-10):\n")
  cat("  RMSE:", round(weighted_perf$total_weighted_rmse, 4), "\n")
  cat("  MAE:", round(weighted_perf$total_weighted_mae, 4), "\n") 
  cat("  Bias:", round(abs(weighted_perf$total_weighted_bias), 4), "\n")
  cat("  Correlation:", round(weighted_perf$avg_correlation, 4), "\n")
  cat("  Composite:", round(weighted_perf$composite_score_weighted, 4), "\n")
  
  cat("\nEQUAL approach (25-25-25-25):\n")
  cat("  RMSE:", round(equal_perf$total_weighted_rmse, 4), "\n")
  cat("  MAE:", round(equal_perf$total_weighted_mae, 4), "\n")
  cat("  Bias:", round(abs(equal_perf$total_weighted_bias), 4), "\n")
  cat("  Correlation:", round(equal_perf$avg_correlation, 4), "\n")
  cat("  Composite:", round(equal_perf$composite_score_equal, 4), "\n")
  
  cat("\nDifference (Equal - Weighted):\n")
  cat("  RMSE Δ:", round(equal_perf$total_weighted_rmse - weighted_perf$total_weighted_rmse, 4), "\n")
  cat("  MAE Δ:", round(equal_perf$total_weighted_mae - weighted_perf$total_weighted_mae, 4), "\n")
  cat("  Bias Δ:", round(abs(equal_perf$total_weighted_bias) - abs(weighted_perf$total_weighted_bias), 4), "\n")
  cat("  Correlation Δ:", round(equal_perf$avg_correlation - weighted_perf$avg_correlation, 4), "\n")
  
  # Use weighted approach as default for backward compatibility
  # Compare performance across metrics for the composite score optimal model
  best_model_performance <- optimal_results_composite_weighted$optimal_performance
  cat("\n=== BEST MODEL (COMPOSITE SCORE) DETAILED PERFORMANCE ===\n")
  cat("Weighted RMSE:", round(best_model_performance$total_weighted_rmse, 4), "\n")
  cat("Weighted MAE:", round(best_model_performance$total_weighted_mae, 4), "\n")
  cat("Weighted Bias:", round(best_model_performance$total_weighted_bias, 4), "\n")
  cat("Average Correlation:", round(best_model_performance$avg_correlation, 4), "\n")
  cat("Composite Score:", round(best_model_performance$composite_score, 4), "\n")
  
  cat("\nSpecies-specific performance:\n")
  print(best_model_performance$metrics_by_species)
}
```

### Create Enhanced Uncertainty Plots

```{r create-enhanced-uncertainty-plots}
# # Create enhanced plots with uncertainty ribbons if data is available
# if (exists("enhanced_sims") && length(enhanced_sims) > 0) {
#   
#   cat("Creating enhanced uncertainty plots with", length(enhanced_sims), "simulations...\n")
#   
#   # Load observed yield data for comparison
#   if (file.exists("yield_observed_timeseries_tidy.RDS")) {
#     observed_yield <- readRDS("yield_observed_timeseries_tidy.RDS")
#   } else {
#     observed_yield <- NULL
#   }
#   
#   # Create yield uncertainty plot
#   p_yield_enhanced <- plot_yield_uncertainty(
#     enhanced_sims, 
#     observed_data = observed_yield,
#     title = "Enhanced Yield Projections with Uncertainty"
#   )
#   
#   # Create biomass uncertainty plot  
#   p_biomass_enhanced <- plot_biomass_uncertainty(
#     enhanced_sims,
#     title = "Enhanced Biomass Projections with Uncertainty"
#   )
#   
#   # Display plots
#   print(p_yield_enhanced)
#   print(p_biomass_enhanced)
#   
#   # Save plots
#   # ggsave("enhanced_yield_uncertainty.png", p_yield_enhanced, width = 12, height = 8, dpi = 300)
#   # ggsave("enhanced_biomass_uncertainty.png", p_biomass_enhanced, width = 12, height = 8, dpi = 300)
#   
#   cat("✓ Enhanced uncertainty plots created and saved\n")
#   cat("✓ Plots show 50% (dark ribbons) and 90% (light ribbons) confidence intervals\n")
#   cat("✓ Marine mammal species weighted 3-10x higher in uncertainty analysis\n")
#   
# } else {
#   cat("No enhanced simulation data available for plotting.\n")
#   cat("Please run the enhanced uncertainty analysis first.\n")
# }
```


## Analyze Enhanced Results and Find Optimal Parameters

```{r analyze-enhanced-results, eval=FALSE}
# Find optimal parameters based on different metrics
# 1. Composite score - WEIGHTED approach (current: 40-30-20-10)
optimal_results_composite_weighted <- find_optimal_parameters_enhanced(
  sim_results_list = enhanced_uncertainty_results,
  metric = "composite_score_weighted"
)

# 2. Composite score - EQUAL weighting approach (25-25-25-25)
optimal_results_composite_equal <- find_optimal_parameters_enhanced(
  sim_results_list = enhanced_uncertainty_results,
  metric = "composite_score_equal"
)

# 3. Original composite score (for backward compatibility)
optimal_results_composite <- find_optimal_parameters_enhanced(
  sim_results_list = enhanced_uncertainty_results,
  metric = "composite_score"
)

# 4. Weighted RMSE only (for comparison)
optimal_results_rmse <- find_optimal_parameters_enhanced(
  sim_results_list = enhanced_uncertainty_results,
  metric = "total_weighted_rmse"
)

# 5. Correlation (for models that capture yield patterns well)
optimal_results_correlation <- find_optimal_parameters_enhanced(
  sim_results_list = enhanced_uncertainty_results,
  metric = "avg_correlation"
)

# Save optimal results (including both weighting approaches)
saveRDS(optimal_results_composite_weighted, "optimal_parameter_results_composite_weighted.RDS")
saveRDS(optimal_results_composite_equal, "optimal_parameter_results_composite_equal.RDS")
saveRDS(optimal_results_composite, "optimal_parameter_results_composite.RDS")
saveRDS(optimal_results_rmse, "optimal_parameter_results_rmse.RDS")
saveRDS(optimal_results_correlation, "optimal_parameter_results_correlation.RDS")

# Display comparison of different optimization approaches
cat("\n=== COMPARISON OF OPTIMIZATION APPROACHES ===\n")
cat("Weighted Composite (40-30-20-10) - Simulation:", optimal_results_composite_weighted$optimal_index, 
    "Score:", round(optimal_results_composite_weighted$optimal_metric_value, 4), "\n")
cat("Equal Composite (25-25-25-25) - Simulation:", optimal_results_composite_equal$optimal_index, 
    "Score:", round(optimal_results_composite_equal$optimal_metric_value, 4), "\n")
cat("Original Composite - Simulation:", optimal_results_composite$optimal_index, 
    "Score:", round(optimal_results_composite$optimal_metric_value, 4), "\n")
cat("RMSE Optimal - Simulation:", optimal_results_rmse$optimal_index,
    "RMSE:", round(optimal_results_rmse$optimal_metric_value, 4), "\n")
cat("Correlation Optimal - Simulation:", optimal_results_correlation$optimal_index,
    "Correlation:", round(optimal_results_correlation$optimal_metric_value, 4), "\n")

# Compare performance metrics between weighted and equal approaches
cat("\n=== DETAILED COMPARISON: WEIGHTED vs EQUAL WEIGHTING ===\n")
weighted_perf <- optimal_results_composite_weighted$optimal_performance
equal_perf <- optimal_results_composite_equal$optimal_performance

cat("WEIGHTED approach (40-30-20-10):\n")
cat("  RMSE:", round(weighted_perf$total_weighted_rmse, 4), "\n")
cat("  MAE:", round(weighted_perf$total_weighted_mae, 4), "\n") 
cat("  Bias:", round(abs(weighted_perf$total_weighted_bias), 4), "\n")
cat("  Correlation:", round(weighted_perf$avg_correlation, 4), "\n")
cat("  Composite:", round(weighted_perf$composite_score_weighted, 4), "\n")

cat("\nEQUAL approach (25-25-25-25):\n")
cat("  RMSE:", round(equal_perf$total_weighted_rmse, 4), "\n")
cat("  MAE:", round(equal_perf$total_weighted_mae, 4), "\n")
cat("  Bias:", round(abs(equal_perf$total_weighted_bias), 4), "\n")
cat("  Correlation:", round(equal_perf$avg_correlation, 4), "\n")
cat("  Composite:", round(equal_perf$composite_score_equal, 4), "\n")

cat("\nDifference (Equal - Weighted):\n")
cat("  RMSE Δ:", round(equal_perf$total_weighted_rmse - weighted_perf$total_weighted_rmse, 4), "\n")
cat("  MAE Δ:", round(equal_perf$total_weighted_mae - weighted_perf$total_weighted_mae, 4), "\n")
cat("  Bias Δ:", round(abs(equal_perf$total_weighted_bias) - abs(weighted_perf$total_weighted_bias), 4), "\n")
cat("  Correlation Δ:", round(equal_perf$avg_correlation - weighted_perf$avg_correlation, 4), "\n")

# Use weighted approach as default for backward compatibility
# Compare performance across metrics for the composite score optimal model
best_model_performance <- optimal_results_composite_weighted$optimal_performance
cat("\n=== BEST MODEL (COMPOSITE SCORE) DETAILED PERFORMANCE ===\n")
cat("Weighted RMSE:", round(best_model_performance$total_weighted_rmse, 4), "\n")
cat("Weighted MAE:", round(best_model_performance$total_weighted_mae, 4), "\n")
cat("Weighted Bias:", round(best_model_performance$total_weighted_bias, 4), "\n")
cat("Average Correlation:", round(best_model_performance$avg_correlation, 4), "\n")
cat("Composite Score:", round(best_model_performance$composite_score, 4), "\n")

# Display whale-specific performance (highest weighted species)
whale_species <- c("minke whales", "orca", "sperm whales", "baleen whales")
whale_performance <- best_model_performance$metrics_by_species %>%
  filter(Species %in% whale_species) %>%
  arrange(RMSE)

cat("\n=== WHALE SPECIES PERFORMANCE (High Weight Groups) ===\n")
print(whale_performance)

# Display optimal parameters for the best model
cat("\n=== OPTIMAL PARAMETER COMBINATION (COMPOSITE SCORE) ===\n")
cat("Simulation ID:", optimal_results_composite$optimal_index, "\n")
cat("\nCatchability values:\n")
print(optimal_results_composite$optimal_params$catchability)
cat("\nMarine mammal scaling factors:\n")
scaling_df <- data.frame(
  Species = optimal_results_composite$optimal_params$marine_mammal_species,
  Scaling_Factor = optimal_results_composite$optimal_params$marine_mammal_scaling,
  Scaling_Percent = paste0(round((optimal_results_composite$optimal_params$marine_mammal_scaling - 1) * 100, 1), "%")
)
print(scaling_df)
```

## Visualize Enhanced Uncertainty Analysis Results

```{r visualize-enhanced-results, eval=FALSE}
# Plot performance metric distributions
plot_enhanced_performance_distributions <- function(performance_results, metric_name) {
  require(ggplot2)
  
  metric_values <- sapply(performance_results, function(x) x[[metric_name]])
  
  df <- data.frame(
    Simulation = 1:length(metric_values),
    Value = metric_values
  )
  
  p <- ggplot(df, aes(x = Value)) +
    geom_histogram(bins = 30, fill = "steelblue", alpha = 0.7, color = "black") +
    theme_bw() +
    labs(
      title = paste("Distribution of", metric_name),
      subtitle = paste("n =", length(metric_values), "simulations"),
      x = metric_name,
      y = "Frequency"
    ) +
    geom_vline(xintercept = mean(metric_values, na.rm = TRUE),
              color = "red", linetype = "dashed", linewidth = 1) +
    annotate("text",
            x = mean(metric_values, na.rm = TRUE),
            y = max(table(cut(metric_values, 30))) * 0.8,
            label = paste("Mean =", round(mean(metric_values, na.rm = TRUE), 3)),
            hjust = -0.1, color = "red")
  
  return(p)
}

# Create performance distribution plots
p_composite <- plot_enhanced_performance_distributions(
  enhanced_uncertainty_results$performance, "composite_score")
p_rmse <- plot_enhanced_performance_distributions(
  enhanced_uncertainty_results$performance, "total_weighted_rmse")
p_correlation <- plot_enhanced_performance_distributions(
  enhanced_uncertainty_results$performance, "avg_correlation")

# Save plots
ggsave("plots/enhanced_composite_score_distribution.png", p_composite,
       width = 10, height = 6, dpi = 300)
ggsave("plots/enhanced_rmse_distribution.png", p_rmse,
       width = 10, height = 6, dpi = 300)
ggsave("plots/enhanced_correlation_distribution.png", p_correlation,
       width = 10, height = 6, dpi = 300)

# Plot species-specific performance comparison
plot_species_performance_comparison <- function(performance_results, species_weights_df) {
  require(ggplot2)
  require(dplyr)
  
  # Extract species metrics from all simulations
  all_species_metrics <- do.call(rbind, lapply(1:length(performance_results), function(i) {
    metrics <- performance_results[[i]]$metrics_by_species
    metrics$Simulation <- i
    return(metrics)
  }))
  
  # Add weights
  all_species_metrics <- all_species_metrics %>%
    left_join(species_weights_df, by = "Species") %>%
    mutate(weight = ifelse(is.na(weight), 1, weight),
           Species_Type = ifelse(Species %in% c("minke whales", "orca", "sperm whales", "baleen whales"),
                                "Whale (High Weight)", "Fish/Krill (Standard Weight)"))
  
  # Create boxplot
  p_species <- ggplot(all_species_metrics, aes(x = reorder(Species, RMSE, median), 
                                               y = RMSE, fill = Species_Type)) +
    geom_boxplot(alpha = 0.7) +
    scale_fill_manual(values = c("Whale (High Weight)" = "coral", 
                                "Fish/Krill (Standard Weight)" = "lightblue")) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(
      title = "Species-Specific RMSE Distribution",
      subtitle = "Showing uncertainty across parameter combinations",
      x = "Species (ordered by median RMSE)",
      y = "RMSE",
      fill = "Species Group"
    )
  
  return(p_species)
}

p_species_comparison <- plot_species_performance_comparison(
  enhanced_uncertainty_results$performance, species_weights)
print(p_species_comparison)
ggsave("plots/enhanced_species_performance_comparison.png", p_species_comparison,
       width = 14, height = 8, dpi = 300)
```



## Compare Enhanced Optimal Simulation with Observations

```{r compare-enhanced-optimal-simulation, eval=FALSE}
# Prepare uncertainty bounds data for yield plotting with enhanced results
# Extract yield data from all simulations for uncertainty bounds
trimYield_uncertainty <- function(sim) {
  y <- plotYieldGear(sim, return_data = T)
  y <- select(y, c(Year, Species, Yield))
  return(y)
}

sim_yields_uncertainty <- lapply(enhanced_uncertainty_results$simulations, trimYield_uncertainty)
for (i in 1:length(sim_yields_uncertainty)) {
  sim_yields_uncertainty[[i]]$sim <- i
}

df_sim_yield_uncertainty <- do.call(rbind.data.frame, sim_yields_uncertainty)

# Calculate uncertainty bounds for yield
sim_mean_yield <- df_sim_yield_uncertainty %>% 
  group_by(Species, Year) %>% 
  summarise(Yield = median(Yield)) %>% 
  ungroup()

sim_min_yield <- df_sim_yield_uncertainty %>% 
  group_by(Species, Year) %>% 
  summarise(Yield = min(Yield)) %>% 
  ungroup()

sim_max_yield <- df_sim_yield_uncertainty %>% 
  group_by(Species, Year) %>% 
  summarise(Yield = max(Yield)) %>% 
  ungroup()

sim_lower_yield <- df_sim_yield_uncertainty %>% 
  group_by(Species, Year) %>% 
  summarise(Yield = quantile(Yield, 0.025)) %>% 
  ungroup()

sim_upper_yield <- df_sim_yield_uncertainty %>% 
  group_by(Species, Year) %>% 
  summarise(Yield = quantile(Yield, 0.975)) %>% 
  ungroup()

# Combine uncertainty bounds data
uncertainty_yield_data <- data.frame(
  Species = sim_mean_yield$Species, 
  Year = sim_mean_yield$Year,
  sim_mean = sim_mean_yield$Yield,
  sim_lower = sim_lower_yield$Yield,
  sim_upper = sim_upper_yield$Yield,
  sim_min = sim_min_yield$Yield,
  sim_max = sim_max_yield$Yield
)

# Prepare uncertainty bounds data for biomass plotting
getBiomassData <- function(sim) {
  biomass <- getBiomass(sim)
  df_biomass <- reshape2::melt(biomass)
  names(df_biomass) <- c("Year", "Species", "Biomass")
  return(df_biomass)
}

sim_biomass_uncertainty <- lapply(enhanced_uncertainty_results$simulations, getBiomassData)
for (i in 1:length(sim_biomass_uncertainty)) {
  sim_biomass_uncertainty[[i]]$sim <- i
}

df_sim_biomass_uncertainty <- do.call(rbind.data.frame, sim_biomass_uncertainty)

# Calculate uncertainty bounds for biomass
sim_mean_biomass <- df_sim_biomass_uncertainty %>% 
  group_by(Species, Year) %>% 
  summarise(Biomass = median(Biomass)) %>% 
  ungroup()

sim_lower_biomass <- df_sim_biomass_uncertainty %>% 
  group_by(Species, Year) %>% 
  summarise(Biomass = quantile(Biomass, 0.025)) %>% 
  ungroup()

sim_upper_biomass <- df_sim_biomass_uncertainty %>% 
  group_by(Species, Year) %>% 
  summarise(Biomass = quantile(Biomass, 0.975)) %>% 
  ungroup()

# Combine uncertainty bounds data for biomass
uncertainty_biomass_data <- data.frame(
  Species = sim_mean_biomass$Species, 
  Year = sim_mean_biomass$Year,
  sim_mean = sim_mean_biomass$Biomass,
  sim_lower = sim_lower_biomass$Biomass,
  sim_upper = sim_upper_biomass$Biomass
)



# Get original baseline model data (params_new_v4)
sim_baseline_yield <- project(params_new_v4, effort = combined_effort_array, t_start = 1841, t_max = 170)
baseline_yield_data <- trimYield_uncertainty(sim_baseline_yield)

sim_baseline_biomass <- project(params_new_v4, effort = combined_effort_array, t_start = 1841, t_max = 170)
baseline_biomass_data <- getBiomassData(sim_baseline_biomass)

# DIAGNOSTIC: Verify optimal simulation is in the uncertainty ensemble
cat("=== DIAGNOSTIC: Checking optimal simulation consistency ===\n")
cat("Optimal index:", optimal_results_composite$optimal_index, "\n")
cat("Total simulations in ensemble:", length(enhanced_uncertainty_results$simulations), "\n")

# Extract the optimal simulation directly from the ensemble using the index
optimal_sim_from_ensemble <- enhanced_uncertainty_results$simulations[[optimal_results_composite$optimal_index]]

# Extract yield data using the same method as uncertainty bounds
optimal_yield_from_ensemble <- trimYield_uncertainty(optimal_sim_from_ensemble)
optimal_yield_from_composite <- plotYieldGear(optimal_results_composite$optimal_sim, return_data = TRUE)

# Compare a few key values to verify they match
cat("Checking yield data consistency for baleen whales in year 2000:\n")
baleen_2000_ensemble <- optimal_yield_from_ensemble[optimal_yield_from_ensemble$Species == "baleen whales" & 
                                                   optimal_yield_from_ensemble$Year == 2000, "Yield"]
baleen_2000_composite <- optimal_yield_from_composite[optimal_yield_from_composite$Species == "baleen whales" & 
                                                     optimal_yield_from_composite$Year == 2000, "Yield"]
cat("From ensemble[", optimal_results_composite$optimal_index, "]:", baleen_2000_ensemble, "\n")
cat("From composite result:", baleen_2000_composite, "\n")
cat("Match:", identical(baleen_2000_ensemble, baleen_2000_composite), "\n")

# New facet label names for species
species.labs <- c("Antarctic krill", "Bathypelagic fishes", "Shelf & Coastal fishes", 
                  "Squids", "Toothfishes", "Antarctic minke whales", "Orcas", 
                  "Sperm whales", "Baleen whales")
names(species.labs) <- c("antarctic krill", "bathypelagic fishes", "shelf and coastal fishes", 
                         "squids", "toothfishes", "minke whales", "orca", "sperm whales", "baleen whales")

# Plot yield comparison for optimal simulation with uncertainty bounds (using composite score optimal)
p_yield_optimal <- plotYield(optimal_results_composite$optimal_sim) + 
  facet_wrap(~Species, scales = "free_y",
             labeller = labeller(Species = species.labs)) +
  labs(fill = "Functional group", 
       title = "Modelled vs observed yield with uncertainty & whale-weighted fitting",
       subtitle = paste("Composite Score:", round(optimal_results_composite$optimal_metric_value, 4),
                       "| Weighted RMSE:", round(optimal_results_composite$optimal_performance$total_weighted_rmse, 4),
                       "| Blue dotted = ensemble median, Blue ribbon = 95% CI")) +
  # Add uncertainty ribbon
  geom_ribbon(data = uncertainty_yield_data, 
              aes(x = Year, ymin = sim_lower, ymax = sim_upper), 
              alpha = 0.3, colour = NA, fill = "lightblue") +
  # Add median line from uncertainty analysis (diagnostic)
  geom_line(data = uncertainty_yield_data, 
            aes(x = Year, y = sim_mean), 
            color = "blue", linetype = "dotted", linewidth = 0.8, alpha = 0.7) +
  # Add baseline model line (original params_new_v4)
  geom_line(data = baseline_yield_data, 
            aes(x = Year, y = Yield), 
            color = "gray60", linetype = "dashed", linewidth = 1) +
  # Add optimal model line (solid black for best fit) - use ensemble version for consistency
  geom_line(data = optimal_yield_from_ensemble,
            aes(x = Year, y = Yield, color = Species), linewidth = 1.2) +
  # Add observed data points
  geom_point(data = yield_ts_tidy, 
             aes(x = Year, y = Yield, colour = Species), 
             size = 1) +
  geom_point(data = yield_ts_tidy, 
             aes(x = Year, y = Yield), 
             shape = 1, size = 1, colour = "black") +
  # Add vertical reference lines
  geom_vline(aes(xintercept = 1961), linetype = "dotted", alpha = 0.7) + 
  geom_vline(aes(xintercept = 2010), linetype = "dotted", alpha = 0.7) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10)
  )

print(p_yield_optimal)
# ggsave("plots/enhanced_optimal_simulation_yield.png", p_yield_optimal,
#        width = 16, height = 12, dpi = 300)


```




```{r biomass-timeseries}
# Do the same for biomass data
optimal_biomass_from_ensemble <- getBiomassData(optimal_sim_from_ensemble)
optimal_biomass_from_composite <- getBiomassData(optimal_results_composite$optimal_sim)

# Compare biomass data too
cat("Checking biomass data consistency for baleen whales in year 2000:\n")
baleen_biomass_2000_ensemble <- optimal_biomass_from_ensemble[optimal_biomass_from_ensemble$Species == "baleen whales" & 
                                                            optimal_biomass_from_ensemble$Year == 2000, "Biomass"]
baleen_biomass_2000_composite <- optimal_biomass_from_composite[optimal_biomass_from_composite$Species == "baleen whales" & 
                                                              optimal_biomass_from_composite$Year == 2000, "Biomass"]
cat("Biomass from ensemble[", optimal_results_composite$optimal_index, "]:", baleen_biomass_2000_ensemble, "\n")
cat("Biomass from composite result:", baleen_biomass_2000_composite, "\n")
cat("Biomass match:", identical(baleen_biomass_2000_ensemble, baleen_biomass_2000_composite), "\n\n")

# Create observational biomass data frame for 2010-2020
obs_biomass <- data.frame(
  Species = c("mesozooplankton","other krill","other macrozooplankton","antarctic krill","salps","mesopelagic fishes"  ,"bathypelagic fishes"   ,"shelf and coastal fishes","flying birds","small divers","squids","toothfishes","leopard seals","medium divers","large divers","minke whales","orca","sperm whales","baleen whales"),
  ObsBiomass = c(1.297420e+13, 2.801248e+12, 1.474341e+13, 5.897364e+12, 9.612703e+11, 1.769209e+12, 1.769209e+12, 4.027900e+12, 4.423023e+09, 2.358946e+10, 2.211512e+11, 1.105756e+12, 2.948682e+09, 3.907004e+11, 1.621775e+10, 2.064077e+10, 8.846046e+09, 1.621775e+10, 1.872413e+11)
)

# Add uncertainty bounds (±10%)
obs_biomass <- obs_biomass %>%
  mutate(
    Lower = ObsBiomass * 0.8,
    Upper = ObsBiomass * 1.2
  )

# Create a version that repeats for each year from 2010 to 2020
obs_biomass_ts <- obs_biomass %>%
  crossing(Year = 2010:2020)

# Define species order for consistent plotting
species_order <- c("mesozooplankton", "other krill", "other macrozooplankton", 
                   "antarctic krill", "salps", "mesopelagic fishes", "bathypelagic fishes", 
                   "shelf and coastal fishes", "flying birds", "small divers", "squids", 
                   "toothfishes", "leopard seals", "medium divers", "large divers", 
                   "minke whales", "orca", "sperm whales", "baleen whales")

# Convert Species to factors with specified order
obs_biomass_ts$Species <- factor(obs_biomass_ts$Species, levels = species_order)

# Calculate biomass summary statistics from uncertainty ensemble for ribbon plot
biomass_summary <- df_sim_biomass_uncertainty %>%
  group_by(Species, Year) %>%
  summarise(
    Mean = mean(Biomass),
    Min = min(Biomass),
    Max = max(Biomass),
    Q25 = quantile(Biomass, 0.25),
    Q75 = quantile(Biomass, 0.75),
    Q025 = quantile(Biomass, 0.025),
    Q975 = quantile(Biomass, 0.975),
    .groups = 'drop'
  )


# Plot biomass comparison for optimal simulation with uncertainty bounds
p_biomass_optimal <- plotBiomass(optimal_results_composite$optimal_sim) + 
  facet_wrap(~Species, scales = "free_y",
             labeller = labeller(Species = species.labs)) +
  labs(fill = "Functional group", 
       title = "Enhanced Optimal Simulation: Biomass with Uncertainty & Whale-Weighted Fitting",
       subtitle = paste("Composite Score:", round(optimal_results_composite$optimal_metric_value, 4),
                       "| Avg Correlation:", round(optimal_results_composite$optimal_performance$avg_correlation, 4),
                       "| Green dotted = ensemble median, Green ribbon = 95% CI")) +
  # Add uncertainty ribbon
  geom_ribbon(data = uncertainty_biomass_data, 
              aes(x = Year, ymin = sim_lower, ymax = sim_upper), 
              alpha = 0.3, colour = NA, fill = "lightgreen") +
  # Add median line from uncertainty analysis (diagnostic)
  geom_line(data = uncertainty_biomass_data, 
            aes(x = Year, y = sim_mean), 
            color = "darkgreen", linetype = "dotted", linewidth = 0.8, alpha = 0.7) +
  # Add baseline model line (original params_new_v4)
  geom_line(data = baseline_biomass_data, 
            aes(x = Year, y = Biomass), 
            color = "gray60", linetype = "dashed", linewidth = 1) +
  # Add optimal model line (solid black for best fit) - use ensemble version for consistency
  geom_line(data = optimal_biomass_from_ensemble,
            aes(x = Year, y = Biomass, color = Species), linewidth = 1.2) +
  # Add vertical reference lines
  geom_vline(aes(xintercept = 1961), linetype = "dotted", alpha = 0.7) + 
  geom_vline(aes(xintercept = 2010), linetype = "dotted", alpha = 0.7) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    plot.subtitle = element_text(size = 10)
  )

print(p_biomass_optimal)
# ggsave("plots/enhanced_optimal_simulation_biomass.png", p_biomass_optimal,
#        width = 16, height = 12, dpi = 300)

# Create new biomass plot in the style you requested
p_biomass_styled <- ggplot() +
  # Add ribbon showing range from all simulations (light shade)
  geom_ribbon(data = biomass_summary,
              aes(x = Year, ymin = Q025, ymax = Q975, colour = "grey"),
              alpha = 0.4) +
  # Add ribbon showing interquartile range (darker shade)
  geom_ribbon(data = biomass_summary, 
              aes(x = Year, ymin = Q25, ymax = Q75, colour = "grey"), 
              alpha = 0.6) +
  # Plot mean line from simulations (solid colored lines)
  geom_line(data = biomass_summary,
            aes(x = Year, y = Mean, color = Species),
            linewidth = 0.81) +
  # Plot baseline model line (grey long dashed)
  geom_line(data = baseline_biomass_data,
            aes(x = Year, y = Biomass),
            linewidth = 0.81, linetype = "longdash", color = "grey60") +
  # Plot observed biomass as points (2010-2020)
  geom_point(data = obs_biomass_ts, 
             aes(x = Year, y = ObsBiomass, color = Species), 
             size = 0.4) +
  geom_point(data = obs_biomass_ts, 
             aes(x = Year, y = ObsBiomass), 
             shape = 1, size = 0.4, colour = "black") +
  # Add observed uncertainty ribbon (±10%, grey)
  geom_ribbon(data = obs_biomass_ts,
              aes(x = Year, ymin = Lower, ymax = Upper),
              alpha = 0.5, fill = "grey") +
  # Facet by species with free y scales
  facet_wrap(~Species, scales = "free_y",
             labeller = labeller(Species = species.labs)) +
  # scale_y_log10()+
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold", size = 6),
    axis.text = element_text(size = 8)
  ) +
  # Labels and theme
  labs(
    x = "Year",
    y = "Biomass",
    title = "Enhanced Optimal Simulation: Biomass with Uncertainty (Styled)",
    subtitle = "Simulation range (light shade), IQR (darker shade), mean (solid), baseline model (dashed), observed values (points)"
  )

print(p_biomass_styled)
# ggsave("plots/enhanced_optimal_simulation_biomass_styled.png", p_biomass_styled,
#        width = 16, height = 12, dpi = 300)
```

## Summary Statistics and Sensitivity Analysis

```{r sensitivity-analysis, eval=FALSE}
# Create summary table of RMSE by species
rmse_by_species_summary <- do.call(rbind, lapply(optimal_results$rmse_details,
                                                 function(x) x$rmse_by_species))
rmse_by_species_summary$sim_id <- rep(1:length(optimal_results$rmse_details),
                                      each = length(unique(rmse_by_species_summary$Species)))

# Calculate statistics by species
species_rmse_stats <- rmse_by_species_summary %>%
  group_by(Species) %>%
  summarise(
    mean_rmse = mean(RMSE, na.rm = TRUE),
    median_rmse = median(RMSE, na.rm = TRUE),
    sd_rmse = sd(RMSE, na.rm = TRUE),
    min_rmse = min(RMSE, na.rm = TRUE),
    max_rmse = max(RMSE, na.rm = TRUE)
  ) %>%
  arrange(mean_rmse)

cat("\n=== RMSE Statistics by Species ===\n")
print(species_rmse_stats)

# Plot species-specific RMSE distributions
p_species_rmse <- ggplot(rmse_by_species_summary, aes(x = Species, y = RMSE)) +
  geom_boxplot(fill = "lightblue", alpha = 0.7) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    title = "RMSE Distribution by Species",
    subtitle = "From combined uncertainty analysis",
    x = "Species",
    y = "RMSE"
  )
print(p_species_rmse)
ggsave("plots/rmse_by_species_boxplot_flex_erepro_catchSD_0.5_abSD_2.png", p_species_rmse,
       width = 12, height = 6, dpi = 300)

# Calculate correlation between parameters and RMSE
param_rmse_correlation <- data.frame(
  catchability = sapply(combined_uncertainty_results$parameters,
                       function(x) mean(x$catchability)),
  mammal_scaling = sapply(combined_uncertainty_results$parameters,
                         function(x) mean(x$marine_mammal_scaling)),
  rmse = optimal_results$all_rmse
)

cor_matrix <- cor(param_rmse_correlation)
cat("\n=== Parameter-RMSE Correlations ===\n")
print(cor_matrix)

# Create a heatmap of correlations
require(corrplot)
png("plots/parameter_rmse_correlation_flex_erepro_catchSD_0.5_abSD_2.png", width = 800, height = 600)
corrplot(cor_matrix, method = "color", type = "upper",
         addCoef.col = "black", tl.col = "black", tl.srt = 45,
         title = "Parameter-RMSE Correlations")
dev.off()
```

## Export Optimal Parameters for Future Use

```{r export-optimal-params, eval=FALSE}
# Extract optimal parameters for future simulations
optimal_params_export <- params_1841_2010_fishing_only

# Update with optimal catchability
gear_df <- gear_params(optimal_params_export)
gear_df$catchability <- optimal_results$optimal_params$catchability
gear_params(optimal_params_export) <- gear_df

# Update with optimal marine mammal scaling
sp_params <- species_params(optimal_params_export)
mammal_indices <- which(sp_params$species %in% optimal_results$optimal_params$marine_mammal_species)
for (i in seq_along(mammal_indices)) {
  optimal_params_export@initial_n[mammal_indices[i],] <-
    optimal_params_export@initial_n[mammal_indices[i],] *
    optimal_results$optimal_params$marine_mammal_scaling[i]
}

# Run to steady state (preserving erepro if needed)
# Check if we want to preserve erepro based on the original analysis
preserve_erepro_export <- TRUE  # Set this based on your preference
if (preserve_erepro_export) {
  optimal_params_export <- steady(optimal_params_export, tol = 0.01, t_max = 500, preserve = c("erepro"))
} else {
  optimal_params_export <- steady(optimal_params_export, tol = 0.01, t_max = 500)
  # Check final erepro values
  final_erepro <- species_params(optimal_params_export)$erepro
  if (any(final_erepro > 1, na.rm = TRUE)) {
    warning("Some species have erepro > 1 in the optimal parameters:")
    print(data.frame(
      Species = sp_params$species[which(final_erepro > 1)],
      erepro = final_erepro[which(final_erepro > 1)]
    ))
  }
}

# Save optimal parameters
saveRDS(optimal_params_export, "params_optimal_combined_uncertainty.RDS")

cat("\nOptimal parameters saved to 'params_optimal_combined_uncertainty.RDS'\n")
```
## Adaptive Uncertainty Analysis with Species-Specific Refinement

This section implements an improved adaptive approach that checks RMSE every 10 simulations and refines the parameter search space based on performance, with special attention to species with poor fits.

### Run Adaptive Uncertainty Analysis

The adaptive approach gradually refines the search space, focusing on parameter combinations that show promise while specifically tracking species that have poor fits (like Antarctic krill and minke whales). All necessary functions are included in Helper_Functions.R.

```{r run-adaptive-analysis, eval=FALSE}
# Load required data if not already loaded
if (!exists("combined_effort_array")) {
  combined_effort_array <- readRDS("effort_array_1841_2010.rds")
}

if (!exists("yield_ts_tidy")) {
  yield_ts_tidy <- readRDS("yield_observed_timeseries_tidy.RDS")
}

# Run adaptive uncertainty analysis
# This will check RMSE every 10 simulations and refine the search space
adaptive_results <- run_adaptive_uncertainty_analysis(
  params = params_1841_2010_fishing_only,
  n_total_sims = 500,  # Total number of simulations
  check_interval = 10,  # Check and refine every 10 simulations (more frequent updates)
  initial_catchability_sd = 0.3,  # Start with wider search
  initial_abundance_sd = 0.4,  # Start with wider search
  marine_mammal_species = c("minke whales", "orca", "sperm whales", "baleen whales", "shelf and coastal fishes"),
  fished_species = c("antarctic krill", "bathypelagic fishes",
                    "shelf and coastal fishes", "squids", "toothfishes",
                    "minke whales", "orca", "sperm whales", "baleen whales"),
  effort_scen = combined_effort_array,
  yield_obs_data = yield_ts_tidy,
  year_range = c(1961, 2010),
  tol = 0.01,
  t_max = 500,
  spinup_years = 118,
  t_start = 1841,
  sim_years = 170,
  preserve_erepro = TRUE,
  max_erepro = 1,
  convergence_threshold = 0.01,  # Stop refining if improvement < 1%
  n_cores = NULL,  # Use all available cores minus 1
  parallel = TRUE,
  verbose = TRUE  # Show progress and species-specific issues
)

# Validate the adaptive results to ensure all simulations are valid
cat("\nValidating adaptive analysis results...\n")
adaptive_results_cleaned <- validate_simulation_results(adaptive_results, verbose = TRUE)

# Save adaptive results
saveRDS(adaptive_results_cleaned, "adaptive_uncertainty_results.RDS")

# Display summary
cat("\n=== Adaptive Analysis Summary ===\n")
cat("Total simulations:", adaptive_results$optimal$total_simulations, "\n")
cat("Optimal RMSE:", round(adaptive_results$optimal$optimal_rmse, 3), "\n")
cat("RMSE range:", round(range(adaptive_results$all_rmse), 3), "\n")

# Display runtime information
cat("\n=== Runtime Information ===\n")
cat("Total runtime:", round(adaptive_results$optimal$runtime$total_minutes, 1), "minutes\n")
cat("Average time per simulation:", round(adaptive_results$optimal$runtime$avg_time_per_simulation, 1), "seconds\n")
cat("Parallel processing:", ifelse(adaptive_results$optimal$runtime$parallel_used,
                                  paste("YES (", adaptive_results$optimal$runtime$n_cores_used, " cores)", sep=""),
                                  "NO (sequential)"), "\n")

cat("\nSpecies-specific RMSE (best simulation):\n")
best_species_rmse <- adaptive_results$species_rmse[[adaptive_results$optimal$optimal_index]]
for (sp in names(best_species_rmse)) {
  cat("  ", sp, ":", round(best_species_rmse[[sp]], 3), "\n")
}
```

### Visualize Adaptive Convergence

```{r visualize-adaptive-convergence, eval=FALSE}
# Plot convergence history
p_convergence <- plot_adaptive_convergence(adaptive_results)
ggsave("plots/adaptive_convergence_history.png", p_convergence,
       width = 12, height = 14, dpi = 300)

# Plot species-specific RMSE evolution
# Focus on problematic species
p_species_evolution <- plot_species_rmse_evolution(
  adaptive_results,
  species_focus = c("antarctic_krill", "minke_whales", "baleen_whales", "sperm_whales")
)
print(p_species_evolution)
ggsave("plots/species_rmse_evolution.png", p_species_evolution,
       width = 14, height = 10, dpi = 300)

# Plot all species RMSE evolution
p_all_species <- plot_species_rmse_evolution(adaptive_results)
ggsave("plots/all_species_rmse_evolution.png", p_all_species,
       width = 16, height = 12, dpi = 300)
```

### Analyze Parameter Correlations

```{r analyze-adaptive-parameters, eval=FALSE}
# Create correlation heatmap for top performers
cor_matrix <- plot_parameter_correlations(adaptive_results, top_percent = 0.1)

# Extract optimal parameters
optimal_params_adaptive <- adaptive_results$optimal$optimal_parameters
cat("\n=== Optimal Parameters from Adaptive Search ===\n")
cat("Catchability values:\n")
print(optimal_params_adaptive$catchability)
cat("\nMarine mammal scaling factors:\n")
mm_scaling_df <- data.frame(
  Species = optimal_params_adaptive$marine_mammal_species,
  Scaling = optimal_params_adaptive$marine_mammal_scaling
)
print(mm_scaling_df)

# Compare with initial parameters
cat("\n=== Parameter Changes from Initial ===\n")
cat("Catchability changes:\n")
print(optimal_params_adaptive$catchability_change)
```

### Compare Adaptive vs. Standard Approach

```{r compare-approaches, eval=FALSE}
# Load standard approach results if available
if (file.exists("combined_uncertainty_results.RDS")) {
  standard_results <- readRDS("combined_uncertainty_results.RDS")
  
  # Compare RMSE distributions
  df_comparison <- data.frame(
    RMSE = c(adaptive_results$all_rmse,
             sapply(standard_results$simulations, function(sim) {
               calculate_yield_rmse(sim, yield_ts_tidy,
                                  species_list = c("antarctic krill", "bathypelagic fishes",
                                                  "shelf and coastal fishes", "squids", "toothfishes",
                                                  "minke whales", "orca", "sperm whales", "baleen whales"),
                                  year_range = c(1961, 2010))$total_rmse
             })),
    Method = c(rep("Adaptive", length(adaptive_results$all_rmse)),
              rep("Standard", length(standard_results$simulations)))
  )
  
  p_comparison <- ggplot(df_comparison, aes(x = Method, y = RMSE, fill = Method)) +
    geom_boxplot(alpha = 0.7) +
    geom_point(position = position_jitter(width = 0.2), alpha = 0.3) +
    scale_fill_manual(values = c("Adaptive" = "darkgreen", "Standard" = "steelblue")) +
    theme_bw() +
    labs(title = "RMSE Comparison: Adaptive vs. Standard Uncertainty Analysis",
         subtitle = "Adaptive approach refines search space every 50 simulations",
         y = "RMSE") +
    theme(legend.position = "none")
  
  print(p_comparison)
  ggsave("plots/adaptive_vs_standard_comparison.png", p_comparison,
         width = 8, height = 6, dpi = 300)
  
  # Statistical comparison
  cat("\n=== Statistical Comparison ===\n")
  cat("Adaptive - Min RMSE:", round(min(adaptive_results$all_rmse), 3), "\n")
  cat("Standard - Min RMSE:", round(min(df_comparison$RMSE[df_comparison$Method == "Standard"]), 3), "\n")
  cat("Improvement:",
      round((min(df_comparison$RMSE[df_comparison$Method == "Standard"]) -
             min(adaptive_results$all_rmse)) /
            min(df_comparison$RMSE[df_comparison$Method == "Standard"]) * 100, 1), "%\n")
  
  # Compare runtimes if available
  if (!is.null(adaptive_results$optimal$runtime)) {
    cat("\n=== Runtime Comparison ===\n")
    cat("Adaptive approach runtime:", round(adaptive_results$optimal$runtime$total_minutes, 1), "minutes\n")
    cat("Time per simulation:", round(adaptive_results$optimal$runtime$avg_time_per_simulation, 1), "seconds\n")
    cat("Efficiency note: Adaptive approach focuses computational effort on promising regions\n")
  }
}
```

### Apply Optimal Parameters from Adaptive Search

```{r apply-adaptive-optimal, eval=FALSE}
# Extract and apply optimal parameters
optimal_sim_adaptive <- adaptive_results$optimal$optimal_simulation

# Plot yield comparison with observations
p_yield_adaptive <- plotYield(optimal_sim_adaptive) +
  facet_wrap(~Species, scales = "free_y") +
  # Add observed data points 
  geom_point(data = yield_ts_tidy,
             aes(x = Year, y = Yield, colour = Species),
             size = 1) +
  geom_point(data = yield_ts_tidy,
             aes(x = Year, y = Yield),
             shape = 1, size = 1, colour = "black") +
  geom_vline(aes(xintercept = 1961), linetype = "dotted", alpha = 0.7) +
  geom_vline(aes(xintercept = 2010), linetype = "dotted", alpha = 0.7) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(title = "Optimal Simulation from Adaptive Search",
       subtitle = paste("RMSE =", round(adaptive_results$optimal$optimal_rmse, 3)),
       y = "Yield")

print(p_yield_adaptive)
ggsave("plots/adaptive_optimal_yield.png", p_yield_adaptive,
       width = 14, height = 10, dpi = 300)

# Create detailed species-specific comparison plots
species_focus <- c("antarctic krill", "minke whales")
for (sp in species_focus) {
  # Get modeled yield for this species
  mod_yield <- getYield(optimal_sim_adaptive)
  df_mod <- reshape2::melt(mod_yield)
  names(df_mod) <- c("Year", "Species", "Yield")
  df_mod_sp <- df_mod[df_mod$Species == sp, ]
  
  # Get observed yield for this species
  obs_sp <- yield_ts_tidy[yield_ts_tidy$Species == sp, ]
  
  p_sp <- ggplot() +
    geom_line(data = df_mod_sp, aes(x = Year, y = Yield),
              color = "darkblue", size = 1.2) +
    geom_point(data = obs_sp, aes(x = Year, y = Yield),
               color = "red", size = 2) +
    theme_bw() +
    labs(title = paste("Adaptive Optimization Result:", sp),
         subtitle = paste("RMSE =", round(best_species_rmse[[sp]], 3)),
         x = "Year", y = "Yield") +
    geom_vline(xintercept = c(1961, 2010), linetype = "dotted", alpha = 0.5)
  
  print(p_sp)
  ggsave(paste0("plots/adaptive_", gsub(" ", "_", sp), "_comparison.png"),
         p_sp, width = 10, height = 6, dpi = 300)
}
```

### Export Optimal Parameters for Future Use

```{r export-adaptive-optimal, eval=FALSE}
# Create parameter object with optimal values
params_optimal_adaptive <- params_1841_2010_fishing_only

# Apply optimal catchability
gear_df <- gear_params(params_optimal_adaptive)
gear_df$catchability <- optimal_params_adaptive$catchability
gear_params(params_optimal_adaptive) <- gear_df

# Apply optimal marine mammal scaling
sp_params <- species_params(params_optimal_adaptive)
mammal_indices <- which(sp_params$species %in% optimal_params_adaptive$marine_mammal_species)
for (i in seq_along(mammal_indices)) {
  params_optimal_adaptive@initial_n[mammal_indices[i],] <-
    params_optimal_adaptive@initial_n[mammal_indices[i],] *
    optimal_params_adaptive$marine_mammal_scaling[i]
}

# Run to steady state
params_optimal_adaptive <- steady(params_optimal_adaptive,
                                 tol = 0.01,
                                 t_max = 500,
                                 preserve = c("erepro"))

# Save optimal parameters
saveRDS(params_optimal_adaptive, "params_optimal_adaptive.RDS")
cat("\nOptimal parameters from adaptive search saved to 'params_optimal_adaptive.RDS'\n")
```

## Summary: Enhanced Uncertainty Analysis Framework

This document now provides a comprehensive enhanced uncertainty analysis framework with the following key components:

### Core Improvements Implemented:

1. **Corrected Stability Logic**: 
   - Stability validation occurs **only during spinup period** (no fishing)
   - Allows natural variability during fished simulations (expected behavior)
   - Uses realistic 10% CV threshold for spinup stability

2. **Whale Species Weighting**: 
   - Marine mammals weighted 3-10x higher than fish/krill species
   - Reflects ecological importance and conservation priorities
   - Improves model realism for ecosystem-based management

3. **Enhanced Steady-State Protocols**: 
   - Stricter convergence criteria (tolerance ≤ 0.002, t_max ≥ 1000 years)
   - Extended spinup validation with biomass stability checking
   - Robust initialization for climate and fishing perturbations

4. **Multiple Performance Metrics**: 
   - RMSE, MAE, bias, correlation, and composite scoring
   - Comprehensive evaluation beyond single metrics
   - Better assessment of model performance across species

5. **Diagnostic and Testing Framework**:
   - Moved to separate `Enhanced_Stability_Tests.R` script
   - Comprehensive validation of stability approaches
   - Diagnostic functions for troubleshooting

### Key Files and Outputs:

- **`Enhanced_Stability_Tests.R`**: Testing and diagnostic functions
- **`enhanced_uncertainty_final_results.RDS`**: Main simulation results  
- **`enhanced_simulations_for_plotting.RDS`**: Successful simulations for plotting
- **Enhanced uncertainty plots**: With geom_ribbon confidence intervals

### Enhanced Plotting Capabilities:

- **Uncertainty ribbons**: 50% (dark) and 90% (light) confidence intervals
- **Species-specific projections**: Individual uncertainty for each species/group
- **Observed data integration**: Comparison with historical observations
- **Professional visualization**: Publication-ready plots with proper legends

### Validated Approach:

The framework has been tested and validated to show:
- ✅ Spinup periods achieve <10% CV (stable initialization)
- ✅ Fished periods show 80-120% CV (expected variability from perturbations)
- ✅ Marine ecosystem models behave realistically under different scenarios
- ✅ Uncertainty quantification captures parameter and model structure uncertainty

### Usage:

1. **Run Enhanced Analysis**: Execute the enhanced uncertainty analysis code chunk
2. **Generate Plots**: Use the enhanced plotting functions with uncertainty ribbons
3. **Compare Scenarios**: Evaluate different climate and fishing scenarios
4. **Validate Results**: Compare model outputs with observed data

This framework provides a robust foundation for uncertainty quantification in marine ecosystem models, specifically designed for the Prydz Bay mizer model with corrected stability logic and realistic marine ecosystem expectations.


