---
title: "R Notebook"
output: html_notebook
---

Set up the environment:
```{r}
# Load libraries
library(remotes)
remotes::install_github("sizespectrum/therMizer")
library(therMizer)
library(mizerExperimental)
```

Load steady therMizer parameters:
```{r}
# Generate parameters
params <- readRDS("therMizer_params_v04.RDS")
```


```{r}
sim <- project(params)
```



```{r}
pre_yield_years <- 1930:1960

yield_timeseries <- read.csv("catch_timeseries.csv")

rownames(yield_timeseries) <- yield_timeseries$Year

yield_timeseries <- yield_timeseries[,-1]

species_names <- species_params(params)$species

colnames(yield_timeseries) <- species_names

yield_array <- as(yield_timeseries, "matrix")

yield_reference_period <- yield_array[0:31,]

yield_ref_per_mean <- colMeans(yield_reference_period)
```




Set up time steps
```{r}
time_steps <- 1961:2010
```



Set up the resource:
### input npp data

```{r}
isimip_plankton <- read.table("GFDL_resource_spectra_annual.dat") # log10 abundance

isimip_plankton <- as(isimip_plankton, "matrix")
sizes <- params@w_full
rownames(isimip_plankton) <- time_steps
colnames(isimip_plankton) <- sizes # signif(sizes, 6)

dim(isimip_plankton)
```



```{r}
n_pp_array <- array(NA, dim = c(length(time_steps), length(sizes)), dimnames = list(time = time_steps, w = sizes)) # setup an empty array for the corrected plankton

# Fill array
n_pp_array[1,] <- isimip_plankton[1,] 
 for (t in seq(1,length(time_steps) - 1,1)) {
   n_pp_array[t + 1,] <- isimip_plankton[t + 1,]
 }

head(n_pp_array[1,]) #make it comparible
head(n_pp_array[50,])
head(log10(initialNResource(params)))
```


### temperature data

#input temperature data 
```{r}
# tos <- readRDS("tos_annual.rds")
# t500m <- readRDS("t500m_annual.rds")
# t1000m <- readRDS("t1000m_annual.rds")
# t1500m <- readRDS("t1500m_annual.rds")
# tob <- readRDS("tob_annual.rds")

tos <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/tos_annual_WAO_corrected.rds")
t500m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t500m_annual_WAO_corrected.rds")
t1000m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t1000m_annual_WAO_corrected.rds")
t1500m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t1500m_annual_WAO_corrected.rds")
tob <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/tob_annual_WAO_corrected.rds")

ocean_temp <- abind::abind(tos,t500m,t1000m, t1500m, tob, along = 2)

realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")

colnames(ocean_temp) <- realm_names
rownames(ocean_temp) <- time_steps
```


#vertical_migration array set up----
```{r}
realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")
colnames(ocean_temp) <- realm_names
species_names <- as.character(params@species_params$species)
sizes <- params@w

vertical_migration_array <- array(0, dim = (c(length(realm_names),
                                              length(species_names), length(sizes))),
                                  dimnames = list(realm = realm_names, sp = species_names,
                                                  w = signif(sizes, 3)))
```


```{r}
zone01 <- which(realm_names =="tos")
zone02 <- which(realm_names == "t500m")
zone03 <- which(realm_names =="t1000m")
zone04 <- which(realm_names == "t1500m")
zone05 <- which(realm_names == "tob")
```



```{r}
#mesozooplankton
MesZoo <- which(species_names=="mesozooplankton")
vertical_migration_array[zone01, MesZoo, ] <- 0.25
vertical_migration_array[zone02, MesZoo, ] <- 0.25
vertical_migration_array[zone03, MesZoo, ] <- 0.25
vertical_migration_array[zone04, MesZoo, ] <- 0.25
vertical_migration_array[zone05, MesZoo, ] <- 0
#other krill
Oth_krill <- which(species_names=="other krill")
vertical_migration_array[zone01, Oth_krill, ] <- 0.25
vertical_migration_array[zone02, Oth_krill, ] <- 0.25
vertical_migration_array[zone03, Oth_krill, ] <- 0.25
vertical_migration_array[zone04, Oth_krill, ] <- 0.25
vertical_migration_array[zone05, Oth_krill, ] <- 0
#other macrozooplankton
Oth_MacroZoo <- which(species_names=="other macrozooplankton")
vertical_migration_array[zone01, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone02, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone03, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone04, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone05, Oth_MacroZoo, ] <- 0
#antarctic krill
Ant_krill <- which(species_names=="antarctic krill")
vertical_migration_array[zone01, Ant_krill, ] <- 0.25
vertical_migration_array[zone02, Ant_krill, ] <- 0.25
vertical_migration_array[zone03, Ant_krill, ] <- 0.25
vertical_migration_array[zone04, Ant_krill, ] <- 0.25
vertical_migration_array[zone05, Ant_krill, ] <- 0
#salps
salps <- which(species_names=="salps")
vertical_migration_array[zone01, salps, ] <- 0.25
vertical_migration_array[zone02, salps, ] <- 0.25
vertical_migration_array[zone03, salps, ] <- 0.25
vertical_migration_array[zone04, salps, ] <- 0.25
vertical_migration_array[zone05, salps, ] <- 0
#mesopelagic fishes
MesoFish <- which(species_names=="mesopelagic fishes")
vertical_migration_array[zone01, MesoFish, ] <- 0.2
vertical_migration_array[zone02, MesoFish, ] <- 0.4
vertical_migration_array[zone03, MesoFish, ] <- 0.4
vertical_migration_array[zone04, MesoFish, ] <- 0
vertical_migration_array[zone05, MesoFish, ] <- 0
#bathypelagic fishes
BathyFish <- which(species_names=="bathypelagic fishes")
vertical_migration_array[zone01, BathyFish, ] <- 0
vertical_migration_array[zone02, BathyFish, ] <- 0.15
vertical_migration_array[zone03, BathyFish, ] <- 0.15
vertical_migration_array[zone04, BathyFish, ] <- 0.35
vertical_migration_array[zone05, BathyFish, ] <- 0.35
#shelf and coastal fishes
SCFish <- which(species_names=="shelf and coastal fishes")
vertical_migration_array[zone01, SCFish, ] <- 0.2
vertical_migration_array[zone02, SCFish, ] <- 0.2
vertical_migration_array[zone03, SCFish, ] <- 0.2
vertical_migration_array[zone04, SCFish, ] <- 0.2
vertical_migration_array[zone05, SCFish, ] <- 0.2
#flying birds
Birds <- which(species_names=="flying birds")
vertical_migration_array[zone01, Birds, ] <- 1
vertical_migration_array[zone02, Birds, ] <- 0
vertical_migration_array[zone03, Birds, ] <- 0
vertical_migration_array[zone04, Birds, ] <- 0
vertical_migration_array[zone05, Birds, ] <- 0
#small divers
Sm_divers <- which(species_names=="small divers")
vertical_migration_array[zone01, Sm_divers, ] <- 1
vertical_migration_array[zone02, Sm_divers, ] <- 0
vertical_migration_array[zone03, Sm_divers, ] <- 0
vertical_migration_array[zone04, Sm_divers, ] <- 0
vertical_migration_array[zone05, Sm_divers, ] <- 0
#squids
squids <- which(species_names=="squids")
vertical_migration_array[zone01, squids, ] <- 0.2
vertical_migration_array[zone02, squids, ] <- 0.2
vertical_migration_array[zone03, squids, ] <- 0.2
vertical_migration_array[zone04, squids, ] <- 0.2
vertical_migration_array[zone05, squids, ] <- 0.2
#toothfishes
toothfishes <- which(species_names=="toothfishes")
vertical_migration_array[zone01, toothfishes, ] <- 0.2
vertical_migration_array[zone02, toothfishes, ] <- 0.2
vertical_migration_array[zone03, toothfishes, ] <- 0.2
vertical_migration_array[zone04, toothfishes, ] <- 0.2
vertical_migration_array[zone05, toothfishes, ] <- 0.2
#leopard seals
seals <- which(species_names=="leopard seals")
vertical_migration_array[zone01, seals, ] <- 0.5
vertical_migration_array[zone02, seals, ] <- 0.5
vertical_migration_array[zone03, seals, ] <- 0
vertical_migration_array[zone04, seals, ] <- 0
vertical_migration_array[zone05, seals, ] <- 0
#medium divers
Md_divers <- which(species_names=="medium divers")
vertical_migration_array[zone01, Md_divers, ] <- 0.5
vertical_migration_array[zone02, Md_divers, ] <- 0.5
vertical_migration_array[zone03, Md_divers, ] <- 0
vertical_migration_array[zone04, Md_divers, ] <- 0
vertical_migration_array[zone05, Md_divers, ] <- 0
#large diver
Lrg_divers <- which(species_names=="large divers")
vertical_migration_array[zone01, Lrg_divers, ] <- 0.5
vertical_migration_array[zone02, Lrg_divers, ] <- 0.5
vertical_migration_array[zone03, Lrg_divers, ] <- 0
vertical_migration_array[zone04, Lrg_divers, ] <- 0
vertical_migration_array[zone05, Lrg_divers, ] <- 0
#minke whales
minke <- which(species_names=="minke whales")
vertical_migration_array[zone01, minke, ] <- 0.4
vertical_migration_array[zone02, minke, ] <- 0.4
vertical_migration_array[zone03, minke, ] <- 0.1
vertical_migration_array[zone04, minke, ] <- 0.1
vertical_migration_array[zone05, minke, ] <- 0
#orca
orca <- which(species_names=="orca")
vertical_migration_array[zone01, orca, ] <- 0.4
vertical_migration_array[zone02, orca, ] <- 0.4
vertical_migration_array[zone03, orca, ] <- 0.1
vertical_migration_array[zone04, orca, ] <- 0.1
vertical_migration_array[zone05, orca, ] <- 0
#sperm whales
sperm <- which(species_names=="sperm whales")
vertical_migration_array[zone01, sperm, ] <- 0.3
vertical_migration_array[zone02, sperm, ] <- 0.3
vertical_migration_array[zone03, sperm, ] <- 0.2
vertical_migration_array[zone04, sperm, ] <- 0.2
vertical_migration_array[zone05, sperm, ] <- 0
#baleen whales
Baleen <- which(species_names=="baleen whales")
vertical_migration_array[zone01, Baleen, ] <- 0.4
vertical_migration_array[zone02, Baleen, ] <- 0.4
vertical_migration_array[zone03, Baleen, ] <- 0.1
vertical_migration_array[zone04, Baleen, ] <- 0.1
vertical_migration_array[zone05, Baleen, ] <- 0
```


```{r}
exposure_array <- array(0, dim = (c(length(realm_names), length(species_names))), 
                  dimnames = list(realm = realm_names, sp = species_names)) # realm x species

for (r in seq(1,length(realm_names),1)) {
    for (s in seq(1,length(species_names),1)) {
        if (any(vertical_migration_array[r,s,] > 0)) {
            exposure_array[r,s] = 1
        }
    }
}
```



```{r}
isimip_effort <- read.csv("effort_array.csv")

years <- isimip_effort[,1]

isimip_effort <- isimip_effort[,-1]
rownames(isimip_effort) <- years

# colnames(isimip_effort) <- rownames(gear_params(params))
# colnames(isimip_effort) <-  gear_params(params)$gear # rep("kinfe_edge_gear",19)
 # colnames(isimip_effort) <- rep("knife_edge_gear",19)
 
for (i in 1:ncol(isimip_effort)) {
  colnames(isimip_effort)[i] <- "knife_edge_gear"
}

head(isimip_effort)
glimpse(isimip_effort)
dim(isimip_effort)

effort_1961_2010 <- isimip_effort[-c(1:31,82:90),]

isimip_effort_array <- as(effort_1961_2010, "matrix")

gear_name = colnames(isimip_effort)


effort_array <- array(NA, c(length(time_steps), length(gear_name)), dimnames = list(time = time_steps, gear = gear_name))
effort_array[1,] = isimip_effort_array[1,]
for (t in seq(1,length(time_steps) - 1, 1)) {
  effort_array[t + 1,] <- isimip_effort_array[t + 1,]
}

effort_array[is.na(effort_array)] <- 0
```


```{r}
gear_params(params)
```



Add temperature and resource information to the parameters:
```{r}
# params <- upgradeTherParams(params, 
#                             ocean_temp_array = ocean_temp,
#                             n_pp_array = n_pp_array, 
#                             vertical_migration_array = vertical_migration_array,
#                             exposure_array = exposure_array,
#                             aerobic_effect = TRUE, metabolism_effect = TRUE)

# params_AS_off <- upgradeTherParams(params, 
#                             ocean_temp_array = ocean_temp,
#                             n_pp_array = n_pp_array, 
#                             vertical_migration_array = vertical_migration_array,
#                             exposure_array = exposure_array,
#                             aerobic_effect = FALSE, metabolism_effect = TRUE)
# 
# params_MET_off <- upgradeTherParams(params, 
#                             ocean_temp_array = ocean_temp,
#                             n_pp_array = n_pp_array, 
#                             vertical_migration_array = vertical_migration_array,
#                             exposure_array = exposure_array,
#                             aerobic_effect = TRUE, metabolism_effect = FALSE)
# 
# params_BOTH_off <- upgradeTherParams(params, 
#                             ocean_temp_array = ocean_temp,
#                             n_pp_array = n_pp_array, 
#                             vertical_migration_array = vertical_migration_array,
#                             exposure_array = exposure_array,
#                             aerobic_effect = FALSE, metabolism_effect = FALSE)

# Nice test to see that things are working correctly
plotTherPerformance(params)
# plotTherPerformance(params_AS_off)

```


```{r}
params_v2 <- params

# gear_params(params_v2)$catchability <- ifelse(gear_params(params_v2)$catchability > 0, 0.5, gear_params(params_v2)$catchability)

gear_params(params_v2)$catchability <- 0.5

```


```{r}
dim(effort_array)
head(effort_array)
effort_array[1,]

gear_params(params_v2)$gear <- gear_params(params_v2)$species

gear_name = gear_params(params_v2)$gear
colnames(effort_array) <- gear_name
head(effort_array)
# 
# gear_params(params)$gear <- gear_name
# gear_params(params)$gear
# head(effort_array)
```





```{r}
# params_v2@initial_effort <- effort_array[1,]

ref_period_effort <- colMeans(effort_array[c(40:50),])

params_v2@initial_effort <- ref_period_effort

params_v2 <- steady(params_v2)

sim_ref <- mizer::project(params_v2, t_max = 200, effort = ref_period_effort)

plotYieldObservedVsModel(params_v2)
plotYieldObservedVsModel(sim_ref, ratio = T)
gear_params(sim_ref@params)$catchability
plotYield(sim_ref)
```

```{r}
#Test ratio
ratio<- plotYieldObservedVsModel(params_v2, return_data = TRUE)$ratio
gear_params(params_v2)$catchability[c(8,12,16,19)]<-gear_params(params_v2)$catchability[c(8,12,16,19)]/ratio

gear_params(params_v2)$catchability <- ifelse(gear_params(params_v2)$catchability == 0.5, 0, gear_params(params_v2)$catchability)

gear_params(params_v2)$catchability[11] <- 0.1

params_v3 <- steady(params_v2)

plotYieldObservedVsModel(params_v3, ratio=T)

# ref_period_effort
```


```{r}
IWC_effort_full <- read.csv("FishMIP_fishing_data/IWC_effort.csv")

glimpse(IWC_effort_full)
```

```{r}
IWC_effort_full_tidy <- IWC_effort_full %>% 
  select(Year, Species, effort_standard) %>% 
  filter(Year <= 1960) %>% 
  group_by(Species) %>% 
  summarise(mean_effort = mean(effort_standard))

IWC_effort_full_tidy
```


Try to use optim to find a more appropriate steady-state prior to initial year of the model run (i.e., 1961), using the yield rather than the biomasses that were used to find the steady-state for the 2010-2019 period


```{r}
# params_v2@initial_effort <- effort_array[1,]
params_yield <- params_v3

species_params(params_yield)$yield_observed <- yield_ref_per_mean

pre_ref_period_effort <- ref_period_effort
pre_ref_period_effort[19] <- 0.3488271
pre_ref_period_effort[18] <- 0.3190192

params_yield@initial_effort <- pre_ref_period_effort


params_yield_v2 <-  params_yield %>% 
  steady() %>% calibrateYield() %>% steady() %>% 
  calibrateYield() %>% steady() %>% 
  calibrateYield() %>% steady() %>% 
  calibrateYield() %>% steady() %>% 
  calibrateYield() %>% steady() %>% 
  calibrateYield() %>% steady() %>% 
  calibrateYield() %>% steady() %>% 
  calibrateYield() %>% steady()

sim_ref <- mizer::project(params_yield_v2, t_max = 200, effort = pre_ref_period_effort)

plotYieldObservedVsModel(params_yield)
plotYieldObservedVsModel(sim_ref, ratio = T)
gear_params(sim_ref@params)$catchability
plotYield(sim_ref)
```

```{r}
params_yield_v2 <- tuneParams(params_yield_v2)
```



```{r}
yield_ref_per_mean
```


```{r}
# vary_df <- data.frame("name" = c("erepro"),
#                       "length" = c(19),
#                       "lower" = c(0.000001), 
#                       "upper" = c(0.999999),
#                       "slot" = c("species_params"),
#                       "unit" = c("linear"))
# 
# vary <- c(params_v3@species_params$erepro)
```


```{r}
# tic()
# optim_results <- mizerHowTo::fastOptim(params = params_v3,
#                  vary = vary,
#                  vary_df = vary_df,
#                  errorFun = mizerHowTo::getErrorCustom,
#                  data_type = "yield_observed",
#                  time_series = effort_array)
# toc()
# 
# saveRDS(optim_results, "optimPar_yield_TS.RDS")

```


```{r}
paramsT <- upgradeTherParams(params_yield_v2, 
                            ocean_temp_array = ocean_temp,
                            n_pp_array = n_pp_array, 
                            vertical_migration_array = vertical_migration_array,
                            exposure_array = exposure_array,
                            aerobic_effect = TRUE, metabolism_effect = TRUE)

```


```{r}
paramsT <- upgradeTherParams(params_v3, 
                            ocean_temp_array = ocean_temp,
                            n_pp_array = n_pp_array, 
                            vertical_migration_array = vertical_migration_array,
                            exposure_array = exposure_array,
                            aerobic_effect = TRUE, metabolism_effect = TRUE)

```


Project to steady-state, with and without fishing:
```{r}
# Project to steady
# params_steady_F <- projectToSteady(paramsT, t_max = 50, effort = effort_array[1,]) 
paramsT@initial_effort <- effort_array[1,] # skip if using 1930-1960 mean yield
params_steady_F <- steady(paramsT, t_max = 200, preserve = c("reproduction_level")) 
```


```{r}
new_repro_level <- getReproductionLevel(params_steady_F)
new_repro_level[17] <- 0.999999

new_erepro_level <- species_params(params_steady_F)$erepro
new_erepro_level[18] <- 0.999999

```

## steady with yield
```{r}
params_steady_F_v2 <- setBevertonHolt(params_steady_F, reproduction_level = new_repro_level)
params_steady_F_v2 <- setBevertonHolt(params_steady_F_v2, erepro = new_erepro_level)

params_steady_F_v2 <- steady(params_steady_F_v2, t_max = 200, preserve = c("erepro"))

getReproductionLevel(params_steady_F_v2)
species_params(params_steady_F_v2)$erepro

params_steady_F_v2 <- params_steady_F_v2 %>% matchYields() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchYields() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchYields() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchYields() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchYields() %>% steady(t_max = 200, preserve = c("erepro"))

plotBiomassObservedVsModel(params_steady_F_v2)
params_steady_F_v2@initial_effort
```

## steady with biomass

```{r}
params_steady_F_v2 <- setBevertonHolt(params_steady_F, reproduction_level = new_repro_level)
params_steady_F_v2 <- setBevertonHolt(params_steady_F_v2, erepro = new_erepro_level)

params_steady_F_v2 <- steady(params_steady_F_v2, t_max = 200, preserve = c("erepro"))

getReproductionLevel(params_steady_F_v2)
species_params(params_steady_F_v2)$erepro

params_steady_F_v2 <- params_steady_F_v2 %>% matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro"))

plotBiomassObservedVsModel(params_steady_F_v2)
params_steady_F_v2@initial_effort
```

# Using steady-state from reference period (2010-2019)
```{r}
# params_steady_noF <- projectToSteady(params_steady_F_v2, t_max = 200, effort = 0) 
# sim_steady_noF <- projectToSteady(paramsT, t_max = 50, effort = 0, return_sim = TRUE) 
# 
# # Include a validation run with F = 0.2 (the estimated F for most of the species modeled, per stock assessments)
# params_steady_validate <- projectToSteady(paramsT, t_max = 50, effort = 0.2) 
# sim_steady_validate <- projectToSteady(paramsT, t_max = 50, effort = 0.2, return_sim = TRUE) # replace this with the calibration matchBiomass() steps from Ezekiel's
# 
# # Test to see how things look
# plot(sim_steady_F)
# plot(sim_steady_noF)
# plot(sim_steady_validate)
```
# Using steady-state based on mean yield from 1930-1960

```{r}
# getReproductionLevel(params_steady_F_v2)
# params_steady_F_v2@species_params$erepro

params_steady_noF <- params_steady_F_v2
params_steady_noF@initial_effort <- 0

params_steady_noF <- steady(params_steady_noF, t_max = 100) 

# sim_steady_validate <- projectToSteady(paramsT, t_max = 50, effort = 0.2, return_sim = TRUE) # replace this with the calibration matchBiomass() steps from Ezekiel's

# Test to see how things look
plot(params_steady_noF)
```


Initial effort from the reference period used to create steady-state so that the effort 

Effort from 1961 to steady state and then use


Run simulations:
```{r}
# # And do a simulation using the steady parameters
# projection_steady_F <- mizer::project(params_steady_F_v2, t_max = 50, effort = effort_array)
# projection_steady_noF <- mizer::project(params_steady_noF, t_max = 50,  effort = 0)
# 
# # projection_steady_validate <- mizer::project(params_steady_validate, t_max = 50,  effort = 0.2)
# 
# # Test to see how things look
# plot(projection_steady_F)
# plot(projection_steady_noF)
# # plot(projection_steady_validate)
# 
# plotYield(projection_steady_F)
# plotlyBiomass(projection_steady_F)
# plotlyBiomass(projection_steady_noF)
# 
# plotlyBiomassRelative(projection_steady_F)
# plotlyBiomassRelative(projection_steady_noF)

```
CLimate data for pre-1961

Take the outputs from controlclim 


Run simulations:
```{r}
# And do a simulation using the steady parameters
projection_steady_F <- mizer::project(params_steady_F_v2, t_max = 50, effort = effort_array)
projection_steady_noF <- mizer::project(params_steady_noF, t_max = 50,  effort = 0)

# projection_steady_validate <- mizer::project(params_steady_validate, t_max = 50,  effort = 0.2)

# Test to see how things look
plot(projection_steady_F)
plot(projection_steady_noF)
# plot(projection_steady_validate)

plotYield(projection_steady_F)
plotlyBiomass(projection_steady_F)
plotlyBiomass(projection_steady_noF)

plotlyBiomassRelative(projection_steady_F)
plotlyBiomassRelative(projection_steady_noF)

```













