---
title: "Wrangling FishMIP Fishing Effort"
author: "Kieran Murphy"
date: "Sys.Date()"
output: rmdformats::readthedown
---

Taking fishing effort and catch from FishMIP output located at:

http://portal.sf.utas.edu.au/thredds/catalog/gem/fishmip/catalog.html

Follow InputData/ -> fishing/ -> histsoc/ then use the regional files filter to the Prydz Bay regional model domain

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Packages


```{r install-packages, include = FALSE}
# source('./R/install_required_packages.R') #run this line to install the required packages
```

```{r load-packages, warning=FALSE, message=FALSE}
# library(palmerpenguins) #penguins data
library(tidyverse) #wrangling
# library(tidymodels) #modeling and machine learning
# library(vip) #variable importance
# library(fastshap) #SHapley Additive exPlanations
library(janitor) #cleaning variable names
library(lubridate) #manipulating dates
library(tictoc) #timing computations
# library(ggcorrplot) #correlation plots
```

## The FishMIP data

Load in and inspect the master dataframes for historical catch and effort

```{r FishMIP_hist_data, echo=TRUE}
df_FishMIP_catch <- read.csv("FishMIP_fishing_data/calibration_catch_histsoc_1850_2004_regional_models.csv")
df_FishMIP_effort <- read.csv("FishMIP_fishing_data/effort_histsoc_1841_2010_regional_models.csv")
```

And let's take a glimpse at the two datasets:

```{r glimpse-FishMIP_data}
dplyr::glimpse(df_FishMIP_catch)
dplyr::glimpse(df_FishMIP_effort)
```
Check how many regions are included and identify the region of interest, Prydz Bay
```{r unique-region}
unique(df_FishMIP_catch$region)
```
Great, there it is. It's called `Prydz.Bay`


Filter dataframe so now I have the Prydz Bay data only

```{r filter-catch}
df_Prydz_catch <- df_FishMIP_catch %>% 
  dplyr::filter(region == "Prydz.Bay") %>% 
  dplyr::group_by(Year,Sector,SAUP, FGroup) %>% 
  dplyr::mutate(total_catch = sum(Reported,IUU,Discards))

head(df_Prydz_catch)
```

```{r}
# saveRDS(df_Prydz_catch, "FishMIP_fishing_data/calibration_catch_histsoc_1850_2004_Prydz_Bay.csv")
```


```{r filter-effort}
df_Prydz_effort <- df_FishMIP_effort %>% 
  dplyr::filter(region == "Prydz.Bay")

head(df_Prydz_effort)
dplyr::glimpse(df_Prydz_effort)
summary(df_Prydz_effort)
```

```{r}
# saveRDS(df_Prydz_effort, "FishMIP_fishing_data/effort_histsoc_1841_2010_Prydz_Bay.csv")
```



Time series of catch by functional group
```{r}
df_Prydz_catch %>% 
  ggplot(aes(x = Year, y = total_catch, colour = FGroup)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(~FGroup)
```

Time series of effort by functional group
```{r}
df_Prydz_effort %>% 
  ggplot(aes(x = Year, y = NomActive, colour = FGroup)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(~FGroup)
```

Aggregate `FGroup` so that they align with functional groups in the Prydz Bay size spectrum model

```{r unique_FGroup}
unique(df_Prydz_catch$FGroup)
```


```{r}
params <- readRDS("params_latest_xx.RDS")
params@species_params$species
params@species_params$w_max
```


"demersal<30cm" = "shelf and coastal fishes"

"rays<90cm" = NA - not represented in the current functional groups

"benthopelagic30-90cm" = "shelf and coastal fishes"

"benthopelagic>=90cm" = "toothfishes"

"krill" = "antarctic krill"

"pelagic30-90cm" = "shelf and coastal fishes"

"bathydemersal>=90cm" = "toothfishes"

"pelagic<30cm" = "shelf and coastal fishes": When I extracted data at the Marine Ecosystem (ME) level on sea around us for East Antarctic Dronning Maud Land, East Antarctic Enderby Land, and East Antarctic Wilkes Land (which describe the model region better than using the RFMO selection of CCAMLR), all the reported catch fit into the "shelf and coastal fishes" group. At the CCAMLR level, there are significant lanternfish catches, i.e., "mesopelagic fishes"

"lobsterscrab" = NA - not represented in the current functional groups

"cephalopods" = "squids"

"demersal30-90cm" = "shelf and coastal fishes"

"bathypelagic<30cm" = "bathypelagic fishes"

"bathydemersal30-90cm" = "shelf and coastal fishes"

```{r tidy-penguins}
df_Prydz_catch_prepared <- df_Prydz_catch %>%
  # janitor::clean_names() %>%
  # dplyr::select(c(species, island, date_egg, culmen_length_mm:delta_13_c_o_oo)) %>%
  dplyr::mutate(species = case_when(FGroup == "demersal<30cm" ~ "shelf and coastal fishes",
                             FGroup == "rays<90cm" ~ "rays<90cm",
                             FGroup == "benthopelagic30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "benthopelagic>=90cm" ~ "toothfishes",
                             FGroup == "krill" ~ "antarctic krill",
                             FGroup == "pelagic30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "bathydemersal>=90cm" ~ "toothfishes",
                             FGroup == "pelagic<30cm" ~ "shelf and coastal fishes",
                             FGroup == "lobsterscrab" ~ "lobsterscrab",
                             FGroup == "cephalopods" ~ "squids",
                             FGroup == "demersal30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "bathypelagic<30cm" ~ "bathypelagic fishes",
                             FGroup == "bathydemersal30-90cm" ~ "shelf and coastal fishes"
                             )) %>% 
  filter(!FGroup=="rays<90cm" & !FGroup=="lobsterscrab") %>% 
  group_by(Year, species) %>% 
  summarise(catch = sum(total_catch))
  # dplyr::mutate(species = factor(species),
  #               island = factor(island),
  #               sex = factor(sex),
  #               year = lubridate::year(date_egg)) %>%
  # dplyr::rename(bill_length_mm = culmen_length_mm,
  #               bill_depth_mm = culmen_depth_mm,
  #               delta_15_n = delta_15_n_o_oo,
  #               delta_13_c = delta_13_c_o_oo) %>%
  # tidyr::drop_na()

glimpse(df_Prydz_catch_prepared)
head(df_Prydz_catch_prepared)
summary(df_Prydz_catch_prepared)

# saveRDS(df_Prydz_catch_prepared, "FishMIP_fishing_data/catch_Prydz_Bay_1950_2004_resolution_v1.RDS")
```

Time series of catch by functional group
```{r}
df_Prydz_catch_prepared %>% 
  ggplot(aes(x = Year, y = catch, colour = species)) +
  geom_point() +
  scale_y_log10() +
  facet_wrap(~species)
```

The other option is to remove effort by functional group and just use effort  gear types

Select the main gear types and filter the region effort based on these
"Trawl_Bottom" = ice fish (shelf & coastal fish + toothfish) each gear can have more than one species (see gearParams())
Midwater/pelagic trawl = krill
Longline = toothfish

```{r unique_Gear}
unique(df_Prydz_effort$Gear)
```
The gears don't seem to match up well with what should actually be catching what, i.e., "Lines_Handlines_and_poles", "Lines_Longlines", "Pots_and_Traps" etc catching krill 


```{r tidy-penguins}
df_Prydz_effort_prepared <- df_Prydz_effort %>%
  # janitor::clean_names() %>%
  # dplyr::select(c(species, island, date_egg, culmen_length_mm:delta_13_c_o_oo)) %>%
  dplyr::mutate(species = case_when(FGroup == "demersal<30cm" ~ "shelf and coastal fishes",
                             FGroup == "rays<90cm" ~ "rays<90cm",
                             FGroup == "benthopelagic30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "benthopelagic>=90cm" ~ "toothfishes",
                             FGroup == "krill" ~ "antarctic krill",
                             FGroup == "pelagic30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "bathydemersal>=90cm" ~ "toothfishes",
                             FGroup == "pelagic<30cm" ~ "shelf and coastal fishes",
                             FGroup == "lobsterscrab" ~ "lobsterscrab",
                             FGroup == "cephalopods" ~ "squids",
                             FGroup == "demersal30-90cm" ~ "shelf and coastal fishes",
                             FGroup == "bathypelagic<30cm" ~ "bathypelagic fishes",
                             FGroup == "bathydemersal30-90cm" ~ "shelf and coastal fishes"
                             )) %>% 
  filter(!FGroup=="rays<90cm" & !FGroup=="lobsterscrab") %>% 
  group_by(Year, species) %>% 
  summarise(effort = sum(NomActive)) %>% 
  filter(!effort < 1e-15) %>%
  group_by(species) %>% 
  mutate(max_effort = max(effort)) %>%
  ungroup() %>% 
  group_by(species,Year) %>%
  mutate(effort_standard = effort/max_effort) %>% 
  ungroup() %>% 
  # dplyr::mutate(species = factor(species),
  #               island = factor(island),
  #               sex = factor(sex),
  #               year = lubridate::year(date_egg)) %>%
  # dplyr::rename(bill_length_mm = culmen_length_mm,
  #               bill_depth_mm = culmen_depth_mm,
  #               delta_15_n = delta_15_n_o_oo,
  #               delta_13_c = delta_13_c_o_oo) %>%
  tidyr::drop_na()

glimpse(df_Prydz_effort_prepared)
head(df_Prydz_effort_prepared)
summary(df_Prydz_effort_prepared)

# saveRDS(df_Prydz_effort_prepared, "FishMIP_fishing_data/effort_Prydz_Bay_1950_2010_resolution_v1.RDS")
```

Time series of effort by functional group
```{r}
df_Prydz_effort_prepared %>% 
  ggplot(aes(x = Year, y = effort, colour = species)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(~species)
```


Time series of effort by functional group
```{r}
df_Prydz_effort_prepared %>% 
  ggplot(aes(x = Year, y = effort_standard, colour = species)) +
  geom_point() +
  # scale_y_log10() +
  facet_wrap(~species)
```
