---
title: "R Notebook"
output: html_notebook
---

Set up the environment:
```{r}
# Load libraries
library(remotes)
remotes::install_github("sizespectrum/therMizer")
library(therMizer)
library(mizerExperimental)
library(tidyverse)
```

Load steady therMizer parameters:
```{r}
# Generate parameters
params <- readRDS("therMizer_params_v04.RDS")
```



```{r}
yield_timeseries <- read.csv("catch_timeseries.csv")

rownames(yield_timeseries) <- yield_timeseries$Year

yield_timeseries <- yield_timeseries[,-1]

species_names <- species_params(params)$species

colnames(yield_timeseries) <- species_names

yield_array <- as(yield_timeseries, "matrix")
```







Set up time steps
```{r}
time_steps <- 1961:2010

# time_steps <- seq(as.Date("1961-01-01"),as.Date("2010-12-01"),by="months")
```



Set up the resource:
### input npp data

```{r}
isimip_plankton <- read.table("GFDL_resource_spectra_annual.dat") # log10 abundance
# isimip_plankton <- read.table("FishMIP_Plankton_Forcing/GFDL_resource_spectra_w_full_142.dat") # log10 abundance

isimip_plankton <- as(isimip_plankton, "matrix")
sizes <- params@w_full
rownames(isimip_plankton) <- time_steps
colnames(isimip_plankton) <- sizes # signif(sizes, 6)

dim(isimip_plankton)
```



```{r}
n_pp_array <- array(NA, dim = c(length(time_steps), length(sizes)), dimnames = list(time = time_steps, w = sizes)) # setup an empty array for the corrected plankton

# Fill array
n_pp_array[1,] <- isimip_plankton[1,] 
 for (t in seq(1,length(time_steps) - 1,1)) {
   n_pp_array[t + 1,] <- isimip_plankton[t + 1,]
 }

head(n_pp_array[1,]) #make it comparible
head(n_pp_array[50,])
head(log10(initialNResource(params)))
```


### temperature data

#input temperature data 
```{r}
# tos <- readRDS("FishMIP_Temperature_Forcing/tos.rds")
# t500m <- readRDS("FishMIP_Temperature_Forcing/t500m.rds")
# t1000m <- readRDS("FishMIP_Temperature_Forcing/t1000m.rds")
# t1500m <- readRDS("FishMIP_Temperature_Forcing/t1500m.rds")
# tob <- readRDS("FishMIP_Temperature_Forcing/tob.rds")

tos <- readRDS("FishMIP_Temperature_Forcing/tos_annual.rds")
t500m <- readRDS("FishMIP_Temperature_Forcing/t500m_annual.rds")
t1000m <- readRDS("FishMIP_Temperature_Forcing/t1000m_annual.rds")
t1500m <- readRDS("FishMIP_Temperature_Forcing/t1500m_annual.rds")
tob <- readRDS("FishMIP_Temperature_Forcing/tob_annual.rds")

# tos <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/tos_annual_WAO_corrected.rds")
# t500m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t500m_annual_WAO_corrected.rds")
# t1000m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t1000m_annual_WAO_corrected.rds")
# t1500m <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/t1500m_annual_WAO_corrected.rds")
# tob <- readRDS("FishMIP_Temperature_Forcing/WAO_corrected/tob_annual_WAO_corrected.rds")

ocean_temp <- abind::abind(tos,t500m,t1000m, t1500m, tob, along = 2)

realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")

colnames(ocean_temp) <- realm_names
rownames(ocean_temp) <- time_steps
```


#vertical_migration array set up----
```{r}
realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")
colnames(ocean_temp) <- realm_names
species_names <- as.character(params@species_params$species)
sizes <- params@w

vertical_migration_array <- array(0, dim = (c(length(realm_names),
                                              length(species_names), length(sizes))),
                                  dimnames = list(realm = realm_names, sp = species_names,
                                                  w = signif(sizes, 3)))
```


```{r}
zone01 <- which(realm_names =="tos")
zone02 <- which(realm_names == "t500m")
zone03 <- which(realm_names =="t1000m")
zone04 <- which(realm_names == "t1500m")
zone05 <- which(realm_names == "tob")
```



```{r}
#mesozooplankton
MesZoo <- which(species_names=="mesozooplankton")
vertical_migration_array[zone01, MesZoo, ] <- 0.25
vertical_migration_array[zone02, MesZoo, ] <- 0.25
vertical_migration_array[zone03, MesZoo, ] <- 0.25
vertical_migration_array[zone04, MesZoo, ] <- 0.25
vertical_migration_array[zone05, MesZoo, ] <- 0
#other krill
Oth_krill <- which(species_names=="other krill")
vertical_migration_array[zone01, Oth_krill, ] <- 0.25
vertical_migration_array[zone02, Oth_krill, ] <- 0.25
vertical_migration_array[zone03, Oth_krill, ] <- 0.25
vertical_migration_array[zone04, Oth_krill, ] <- 0.25
vertical_migration_array[zone05, Oth_krill, ] <- 0
#other macrozooplankton
Oth_MacroZoo <- which(species_names=="other macrozooplankton")
vertical_migration_array[zone01, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone02, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone03, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone04, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone05, Oth_MacroZoo, ] <- 0
#antarctic krill
Ant_krill <- which(species_names=="antarctic krill")
vertical_migration_array[zone01, Ant_krill, ] <- 0.25
vertical_migration_array[zone02, Ant_krill, ] <- 0.25
vertical_migration_array[zone03, Ant_krill, ] <- 0.25
vertical_migration_array[zone04, Ant_krill, ] <- 0.25
vertical_migration_array[zone05, Ant_krill, ] <- 0
#salps
salps <- which(species_names=="salps")
vertical_migration_array[zone01, salps, ] <- 0.25
vertical_migration_array[zone02, salps, ] <- 0.25
vertical_migration_array[zone03, salps, ] <- 0.25
vertical_migration_array[zone04, salps, ] <- 0.25
vertical_migration_array[zone05, salps, ] <- 0
#mesopelagic fishes
MesoFish <- which(species_names=="mesopelagic fishes")
vertical_migration_array[zone01, MesoFish, ] <- 0.2
vertical_migration_array[zone02, MesoFish, ] <- 0.4
vertical_migration_array[zone03, MesoFish, ] <- 0.4
vertical_migration_array[zone04, MesoFish, ] <- 0
vertical_migration_array[zone05, MesoFish, ] <- 0
#bathypelagic fishes
BathyFish <- which(species_names=="bathypelagic fishes")
vertical_migration_array[zone01, BathyFish, ] <- 0
vertical_migration_array[zone02, BathyFish, ] <- 0.15
vertical_migration_array[zone03, BathyFish, ] <- 0.15
vertical_migration_array[zone04, BathyFish, ] <- 0.35
vertical_migration_array[zone05, BathyFish, ] <- 0.35
#shelf and coastal fishes
SCFish <- which(species_names=="shelf and coastal fishes")
vertical_migration_array[zone01, SCFish, ] <- 0.2
vertical_migration_array[zone02, SCFish, ] <- 0.2
vertical_migration_array[zone03, SCFish, ] <- 0.2
vertical_migration_array[zone04, SCFish, ] <- 0.2
vertical_migration_array[zone05, SCFish, ] <- 0.2
#flying birds
Birds <- which(species_names=="flying birds")
vertical_migration_array[zone01, Birds, ] <- 1
vertical_migration_array[zone02, Birds, ] <- 0
vertical_migration_array[zone03, Birds, ] <- 0
vertical_migration_array[zone04, Birds, ] <- 0
vertical_migration_array[zone05, Birds, ] <- 0
#small divers
Sm_divers <- which(species_names=="small divers")
vertical_migration_array[zone01, Sm_divers, ] <- 1
vertical_migration_array[zone02, Sm_divers, ] <- 0
vertical_migration_array[zone03, Sm_divers, ] <- 0
vertical_migration_array[zone04, Sm_divers, ] <- 0
vertical_migration_array[zone05, Sm_divers, ] <- 0
#squids
squids <- which(species_names=="squids")
vertical_migration_array[zone01, squids, ] <- 0.2
vertical_migration_array[zone02, squids, ] <- 0.2
vertical_migration_array[zone03, squids, ] <- 0.2
vertical_migration_array[zone04, squids, ] <- 0.2
vertical_migration_array[zone05, squids, ] <- 0.2
#toothfishes
toothfishes <- which(species_names=="toothfishes")
vertical_migration_array[zone01, toothfishes, ] <- 0.2
vertical_migration_array[zone02, toothfishes, ] <- 0.2
vertical_migration_array[zone03, toothfishes, ] <- 0.2
vertical_migration_array[zone04, toothfishes, ] <- 0.2
vertical_migration_array[zone05, toothfishes, ] <- 0.2
#leopard seals
seals <- which(species_names=="leopard seals")
vertical_migration_array[zone01, seals, ] <- 0.5
vertical_migration_array[zone02, seals, ] <- 0.5
vertical_migration_array[zone03, seals, ] <- 0
vertical_migration_array[zone04, seals, ] <- 0
vertical_migration_array[zone05, seals, ] <- 0
#medium divers
Md_divers <- which(species_names=="medium divers")
vertical_migration_array[zone01, Md_divers, ] <- 0.5
vertical_migration_array[zone02, Md_divers, ] <- 0.5
vertical_migration_array[zone03, Md_divers, ] <- 0
vertical_migration_array[zone04, Md_divers, ] <- 0
vertical_migration_array[zone05, Md_divers, ] <- 0
#large diver
Lrg_divers <- which(species_names=="large divers")
vertical_migration_array[zone01, Lrg_divers, ] <- 0.5
vertical_migration_array[zone02, Lrg_divers, ] <- 0.5
vertical_migration_array[zone03, Lrg_divers, ] <- 0
vertical_migration_array[zone04, Lrg_divers, ] <- 0
vertical_migration_array[zone05, Lrg_divers, ] <- 0
#minke whales
minke <- which(species_names=="minke whales")
vertical_migration_array[zone01, minke, ] <- 0.4
vertical_migration_array[zone02, minke, ] <- 0.4
vertical_migration_array[zone03, minke, ] <- 0.1
vertical_migration_array[zone04, minke, ] <- 0.1
vertical_migration_array[zone05, minke, ] <- 0
#orca
orca <- which(species_names=="orca")
vertical_migration_array[zone01, orca, ] <- 0.4
vertical_migration_array[zone02, orca, ] <- 0.4
vertical_migration_array[zone03, orca, ] <- 0.1
vertical_migration_array[zone04, orca, ] <- 0.1
vertical_migration_array[zone05, orca, ] <- 0
#sperm whales
sperm <- which(species_names=="sperm whales")
vertical_migration_array[zone01, sperm, ] <- 0.3
vertical_migration_array[zone02, sperm, ] <- 0.3
vertical_migration_array[zone03, sperm, ] <- 0.2
vertical_migration_array[zone04, sperm, ] <- 0.2
vertical_migration_array[zone05, sperm, ] <- 0
#baleen whales
Baleen <- which(species_names=="baleen whales")
vertical_migration_array[zone01, Baleen, ] <- 0.4
vertical_migration_array[zone02, Baleen, ] <- 0.4
vertical_migration_array[zone03, Baleen, ] <- 0.1
vertical_migration_array[zone04, Baleen, ] <- 0.1
vertical_migration_array[zone05, Baleen, ] <- 0
```


```{r}
exposure_array <- array(0, dim = (c(length(realm_names), length(species_names))), 
                  dimnames = list(realm = realm_names, sp = species_names)) # realm x species

for (r in seq(1,length(realm_names),1)) {
    for (s in seq(1,length(species_names),1)) {
        if (any(vertical_migration_array[r,s,] > 0)) {
            exposure_array[r,s] = 1
        }
    }
}
```



```{r}
time_steps_effort <- 1961:2010
isimip_effort <- read.csv("effort_array.csv")

years <- isimip_effort[,1]

isimip_effort <- isimip_effort[,-1]
rownames(isimip_effort) <- years

# colnames(isimip_effort) <- rownames(gear_params(params))
# colnames(isimip_effort) <-  gear_params(params)$gear # rep("kinfe_edge_gear",19)
 # colnames(isimip_effort) <- rep("knife_edge_gear",19)
 
for (i in 1:ncol(isimip_effort)) {
  colnames(isimip_effort)[i] <- "knife_edge_gear"
}

head(isimip_effort)
glimpse(isimip_effort)
dim(isimip_effort)

effort_1961_2010 <- isimip_effort[-c(1:31,82:90),]

isimip_effort_array <- as(effort_1961_2010, "matrix")

gear_name = colnames(isimip_effort)


effort_array <- array(NA, c(length(time_steps_effort), length(gear_name)), dimnames = list(time = time_steps_effort, gear = gear_name))
effort_array[1,] = isimip_effort_array[1,]
for (t in seq(1,length(time_steps_effort) - 1, 1)) {
  effort_array[t + 1,] <- isimip_effort_array[t + 1,]
}

effort_array[is.na(effort_array)] <- 0
```


```{r}
gear_params(params)
```



Add temperature and resource information to the parameters:
```{r}
# params <- upgradeTherParams(params, 
#                             ocean_temp_array = ocean_temp,
#                             n_pp_array = n_pp_array, 
#                             vertical_migration_array = vertical_migration_array,
#                             exposure_array = exposure_array,
#                             aerobic_effect = TRUE, metabolism_effect = TRUE)

# params_AS_off <- upgradeTherParams(params, 
#                             ocean_temp_array = ocean_temp,
#                             n_pp_array = n_pp_array, 
#                             vertical_migration_array = vertical_migration_array,
#                             exposure_array = exposure_array,
#                             aerobic_effect = FALSE, metabolism_effect = TRUE)
# 
# params_MET_off <- upgradeTherParams(params, 
#                             ocean_temp_array = ocean_temp,
#                             n_pp_array = n_pp_array, 
#                             vertical_migration_array = vertical_migration_array,
#                             exposure_array = exposure_array,
#                             aerobic_effect = TRUE, metabolism_effect = FALSE)
# 
# params_BOTH_off <- upgradeTherParams(params, 
#                             ocean_temp_array = ocean_temp,
#                             n_pp_array = n_pp_array, 
#                             vertical_migration_array = vertical_migration_array,
#                             exposure_array = exposure_array,
#                             aerobic_effect = FALSE, metabolism_effect = FALSE)

# Nice test to see that things are working correctly
plotTherPerformance(params)
# plotTherPerformance(params_AS_off)

```


```{r}
params_v2 <- params

# gear_params(params_v2)$catchability <- ifelse(gear_params(params_v2)$catchability > 0, 0.5, gear_params(params_v2)$catchability)

gear_params(params_v2)$catchability <- 0.5

```


```{r}
dim(effort_array)
head(effort_array)
effort_array[1,]

gear_params(params_v2)$gear <- gear_params(params_v2)$species
# gear_params(params_steady_F)$gear <- gear_params(params_steady_F)$species

# gear_name = gear_params(params_steady_F)$gear
gear_name = gear_params(params_v2)$gear
colnames(effort_array) <- gear_name
head(effort_array)
# 
# gear_params(params)$gear <- gear_name
# gear_params(params)$gear
# head(effort_array)

# saveRDS(effort_array, "effort_array_therMizer.RDS")
```





```{r}
# params_v2@initial_effort <- effort_array[1,]

ref_period_effort <- colMeans(effort_array[c(40:50),])
model_period_effort <- colMeans(effort_array)

initial_effort_updated <- ref_period_effort
initial_effort_updated[c(4,7,17,18)] <- model_period_effort[c(4,7,17,18)]

params_v2@initial_effort <- initial_effort_updated
# params_v2@initial_effort[c(4,7,11,18)] <- model_period_effort[c(4,7,11,18)]

params_v2 <- steady(params_v2, t_max = 1000)

sim_ref <- mizer::project(params_v2, t_max = 200, effort = initial_effort_updated)

plotYieldObservedVsModel(params_v2)
plotYieldObservedVsModel(sim_ref, ratio = T)
gear_params(sim_ref@params)$catchability
plotYield(sim_ref)
```

```{r}
#Test ratio
ratio<- plotYieldObservedVsModel(params_v2, return_data = TRUE)$ratio
gear_params(params_v2)$catchability[c(7,8,12,16,19)]<-gear_params(params_v2)$catchability[c(7,8,12,16,19)]/ratio

gear_params(params_v2)$catchability <- ifelse(gear_params(params_v2)$catchability == 0.5, 0, gear_params(params_v2)$catchability)

# gear_params(params_v2)$catchability[c(8)] <- gear_params(params_v2)$catchability[c(8)] * 1.5 # shelf and coastal fishes
gear_params(params_v2)$catchability[c(8)] <- 1.267937e-05 * 3 # shelf and coastal fishes

# gear_params(params_v2)$catchability[c(16)] <- gear_params(params_v2)$catchability[c(16)] * 1.5 # minke whales
gear_params(params_v2)$catchability[c(16)] <- 0.354628 * 2 # minke whales

gear_params(params_v2)$catchability[c(11)] <- 0.000005 # squids

gear_params(params_v2)$catchability[c(4)] <- 0.005 # ant krill

gear_params(params_v2)$catchability[c(17)] <- 0.1 # orca

gear_params(params_v2)$catchability[c(18)] <- 1 # sperm whales

params_v3 <- steady(params_v2)

plotYieldObservedVsModel(params_v3, ratio=T)

# ref_period_effort
```


```{r}
ratio_v2<- plotYieldObservedVsModel(params_v3, return_data = TRUE)$ratio

# gear_params(params_v3)$catchability[c(7,8,12,16,19)]<-gear_params(params_v3)$catchability[c(7,8,12,16,19)]/ratio_v2


gear_params(params_v3)$species[12]
gear_params(params_v3)$catchability[12]

ratio_v2[3]

gear_params(params_v3)$catchability[12]<-gear_params(params_v3)$catchability[c(12)]/ratio_v2[3]

params_v4 <- steady(params_v3)

plotYieldObservedVsModel(params_v4, ratio=T)


ratio_v3 <- plotYieldObservedVsModel(params_v4, return_data = TRUE)$ratio

gear_params(params_v4)$catchability[c(7,8,12,16,19)]<-gear_params(params_v4)$catchability[c(7,8,12,16,19)]/ratio_v3

params_v5 <- steady(params_v4)

plotYieldObservedVsModel(params_v5, ratio=T)

```

```{r}
getReproductionLevel(params_v5)
params_v5@species_params$erepro
```


```{r}
new_repro_level <- getReproductionLevel(params_v5)
# new_repro_level[c(17)] <- 0.999999

new_erepro_level <- species_params(params_v5)$erepro
new_erepro_level[c(17,18)] <- 0.999999
# new_erepro_level[c(16,17,18)] <- 0.999999


```

## steady with yield
```{r}
params_v6 <- setBevertonHolt(params_v5, reproduction_level = new_repro_level)
params_v7 <- setBevertonHolt(params_v6, erepro = new_erepro_level)

params_v8 <- steady(params_v7, t_max = 200, preserve = c("erepro"))

getReproductionLevel(params_v8)
species_params(params_v8)$erepro

params_v8 <- params_v8 %>% matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro"))

plotBiomassObservedVsModel(params_v8)
params_v8@initial_effort
```

```{r}
# params_v9 <- tuneParams(params_v8)
```



```{r}
IWC_effort_full <- read.csv("FishMIP_fishing_data/IWC_effort.csv")

glimpse(IWC_effort_full)
```

```{r}
IWC_effort_full_tidy <- IWC_effort_full %>% 
  select(Year, Species, effort_standard) %>% 
  filter(Year <= 1960) %>% 
  group_by(Species) %>% 
  summarise(mean_effort = mean(effort_standard))

IWC_effort_full_tidy
```



```{r}
paramsT <- upgradeTherParams(params_v8, 
                            ocean_temp_array = ocean_temp,
                            n_pp_array = n_pp_array, 
                            vertical_migration_array = vertical_migration_array,
                            exposure_array = exposure_array,
                            aerobic_effect = TRUE, metabolism_effect = TRUE)

```

## steady with biomass

```{r}
params_steady_F <- steady(paramsT, t_max = 200, preserve = c("erepro"))

getReproductionLevel(params_steady_F)
species_params(params_steady_F)$erepro

params_steady_F <- params_steady_F %>% matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro")) %>% 
  matchBiomasses() %>% steady(t_max = 200, preserve = c("erepro"))

getReproductionLevel(params_steady_F)
species_params(params_steady_F)$erepro

plotBiomassObservedVsModel(params_steady_F)
params_steady_F@initial_effort
```

# Using steady-state from reference period (2010-2019)
```{r}
params_steady_F_init1961 <- projectToSteady(params_steady_F, t_max = 200, effort = effort_array[1,])
params_steady_noF <- projectToSteady(params_steady_F, t_max = 200, effort = 0)

# # Include a validation run with F = 0.2 (the estimated F for most of the species modeled, per stock assessments)
# params_steady_validate <- projectToSteady(paramsT, t_max = 50, effort = 0.2) 
# sim_steady_validate <- projectToSteady(paramsT, t_max = 50, effort = 0.2, return_sim = TRUE) # replace this with the calibration matchBiomass() steps from Ezekiel's
# 
# # Test to see how things look
# plot(sim_steady_F)
# plot(sim_steady_noF)
# plot(sim_steady_validate)
```


Initial effort from the reference period used to create steady-state so that the effort 

Effort from 1961 to steady state and then use


Run simulations:
```{r}
params_steady_F <- projection_steady_F@params
params_steady_noF <- projection_steady_noF@params

# And do a simulation using the steady parameters
projection_steady_F <- mizer::project(params_steady_F, t_max = 50, effort = effort_array)
projection_steady_noF <- mizer::project(params_steady_noF, t_start = 1961, t_max = 49,  effort = 0)
# projection_steady_F_init1961 <- mizer::project(params_steady_F_init1961, t_max = 50, effort = effort_array)

# projection_steady_validate <- mizer::project(params_steady_validate, t_max = 50,  effort = 0.2)

# Test to see how things look
plot(projection_steady_F)
plot(projection_steady_noF)
# plot(projection_steady_F_init1961)
# plot(projection_steady_validate)

plotYield(projection_steady_F)
# plotYield(projection_steady_F_init1961)

plotlyBiomass(projection_steady_F)
plotlyBiomass(projection_steady_noF)
# plotlyBiomass(projection_steady_F_init1961)

plotlyBiomassRelative(projection_steady_F)
plotlyBiomassRelative(projection_steady_noF)
# plotlyBiomassRelative(projection_steady_F_init1961)



# saveRDS(projection_steady_F, "ISIMIP3a_1961_2010_fishing.rds")
# saveRDS(projection_steady_noF, "ISIMIP3a_1961_2010_no_fishing.rds")

```
Climate data for pre-1961

Take the outputs from controlclim 


Yield from current reference period (2010-2019)
Ratio of time-averaged yield in the ref period compared to the time-averaged yield in the decade prior to start of forcings (1950s) and then correct biomass according to the ratio.

Assumes a linear relationship between the ratio of yield:biomass, not neccessarily linear between yield and biomass.

Fmsy * unfinished biomass = msy (Can we use this to get the biomass from 1950s)?

To run a future projection would want to run historical scenario that ends in 2014 and then leads into the ssp2.6/8.5 GFDL scenarios

Results for presentaiton: 
- fished/unfished biomass. 
- growth plots
- barplots - comparing fished and unfished
- Variety of realised PPMRs
- Ecosystem indicators for mean of first 5 years to last 5 years - with and without fishing 
Time-average biomasses start and finish of time period and with/without fishing
mizer::getCommunitySlope()
mizer::getProportionOfLargeFish() # large animals or select specific species
mizer::getMeanMaxWeight()
mizer::getMeanWeight()

```{r}
projection_steady_F <- readRDS("ISIMIP3a_1961_2010_fishing.rds")
projection_steady_noF <- readRDS("ISIMIP3a_1961_2010_no_fishing.rds")
```


```{r}
p_B_Fishing <- plotBiomass(projection_steady_F) +
  theme_classic() +
  theme(axis.title =  element_text(size=12),
        axis.text  =  element_text(size=10))

p_B_noFishing <- plotBiomass(projection_steady_noF) +
  theme_classic() +
  theme(axis.title =  element_text(size=12),
        axis.text  =  element_text(size=10))

p_B_rel_Fishing <- plotBiomassRelative(projection_steady_F) +
  theme_classic() +
  theme(axis.title =  element_text(size=12),
        axis.text  =  element_text(size=10))

p_B_rel_noFishing <- plotBiomassRelative(projection_steady_noF) +
  theme_classic() +
  theme(axis.title =  element_text(size=12),
        axis.text  =  element_text(size=10))

p_B_Fishing
p_B_noFishing
p_B_rel_Fishing
p_B_rel_noFishing

ggsave("plots/Biomass_Fishing_1961_2010.tiff", plot = p_B_Fishing, units="cm", width=20, height=16, dpi =300)
ggsave("plots/Biomass_No_Fishing_1961_2010.tiff", plot = p_B_noFishing, units="cm", width=20, height=16, dpi =300)
ggsave("plots/Biomass_Relative_Fishing_1961_2010.tiff", plot = p_B_rel_Fishing, units="cm", width=20, height=16, dpi =300)
ggsave("plots/Biomass_Relative_No_Fishing_1961_2010.tiff", plot = p_B_rel_noFishing, units="cm", width=20, height=16, dpi =300)

```

```{r}
plotDietX(projection_steady_F@params)
```


species.labs <- c("Antarctic krill", "Bathypelagic fishes", "Shelf & Coastal fishes", "Squids", "Toothfishes", "Antarctic minke whales", "Orcas", "Sperm whales", "Baleen whales")
names(species.labs) <- c("antarctic krill", "bathypelagic fishes", "shelf and coastal fishes", "squids", "toothfishes", "minke whales", "orca", "sperm whales", "baleen whales")

```{r}
effort_ts <- effort_array %>% 
  melt()

head(effort_ts)

names(effort_ts) <- c("Year", "Species", "Norm_Effort")

effort_ts_tidy <- effort_ts %>% 
  filter(Species == "shelf and coastal fishes" |
           Species == "antarctic krill" |
           Species == "squids" |
           Species == "bathypelagic fishes" |
           Species == "toothfishes" |
           Species == "minke whales" |
           Species == "baleen whales"|
           Species == "orca"|
           Species == "sperm whales") %>%
  filter(!Norm_Effort == 0)

glimpse(effort_ts_tidy)
head(effort_ts_tidy)
```


```{r}
yield_ts <- yield_array %>% 
  melt()

head(yield_ts)

names(yield_ts) <- c("Year", "Species", "Yield")

yield_ts_tidy <- yield_ts %>% 
  # filter(Species == "shelf and coastal fishes" |
  #          Species == "squids" |
  #          Species == "toothfishes" |
  #          Species == "minke whales" |
  #          Species == "baleen whales"|
  #          Species == "sperm whales") %>% 
  filter(!Yield == 0)

glimpse(yield_ts_tidy)
head(yield_ts_tidy)
```


Plot observed yield
```{r}
# New facet label names for species
species.labs <- c("Antarctic krill", "Bathypelagic fishes", "Shelf & Coastal fishes", "Squids", "Toothfishes", "Antarctic minke whales", "Orcas", "Sperm whales", "Baleen whales")
names(species.labs) <- c("antarctic krill", "bathypelagic fishes", "shelf and coastal fishes", "squids", "toothfishes", "minke whales", "orca", "sperm whales", "baleen whales")

# df_mod_yield <- plotYield(projection_steady_F, return_data = T)

p_yield <- 
  plotYield(projection_steady_F) +
  facet_wrap(~Species,scales="free_y",
             labeller = labeller(Species = species.labs)) +
  labs(fill = "Functional group") +
  # facet_wrap(~Species,scales="free_y") +
  geom_point(data = yield_ts_tidy, aes(x = Year, y = Yield, colour = Species), size=1) +
  geom_point(data = yield_ts_tidy, aes(x = Year, y = Yield), shape = 1,size = 1,colour = "black") +
  theme(legend.title = element_text( size=3), legend.text=element_text(size=3)) + 
  geom_vline(aes(xintercept = 1961), linetype = "dashed") + 
  geom_vline(aes(xintercept = 2010), linetype = "dashed") +
  theme_bw() +
  theme(
    legend.position = "none"
  )
  # theme_classic()

p_yield_const_scale <-
  
  plotYield(projection_steady_F) + 
  facet_wrap(~Species,
             labeller = labeller(Species = species.labs)) +
  labs(fill = "Functional group") +
  # facet_wrap(~Species,scales="free_y") +
  geom_point(data = yield_ts_tidy, aes(x = Year, y = Yield, colour = Species), size=1) +
  geom_point(data = yield_ts_tidy, aes(x = Year, y = Yield), shape = 1,size = 1,colour = "black") +
  theme(legend.title = element_text( size=3), legend.text=element_text(size=3)) + 
  geom_vline(aes(xintercept = 1961), linetype = "dashed") + 
  geom_vline(aes(xintercept = 2010), linetype = "dashed") +
  # scale_x_continuous(breaks = c(1930,1945,1961,1985,2010)) +
  # scale_y_log10() +
  theme_bw() +
  theme(
    legend.position = "none"
  )

ggsave("plots/Compare_Modelled_Observed_Yields_low_res.tiff", plot = p_yield, units="cm", width=20, height=16, dpi =150)
# ggsave("plots/Compare_Modelled_Observed_Yields_same_scale.tiff", plot = p_yield_const_scale, units="cm", width=20, height=16, dpi =300)
```


```{r}
# New facet label names for species
species.labs <- c("Antarctic krill", "Bathypelagic fishes", "Shelf & Coastal fishes", "Squids", "Toothfishes", "Antarctic minke whales", "Orcas", "Sperm whales", "Baleen whales")
names(species.labs) <- c("antarctic krill", "bathypelagic fishes", "shelf and coastal fishes", "squids", "toothfishes", "minke whales", "orca", "sperm whales", "baleen whales")

# df_mod_yield <- plotYield(projection_steady_F, return_data = T)

p_effort <-  effort_ts_tidy %>% 
  ggplot() +
  
  # facet_wrap(~Species,scales="free_y") +
  geom_point(data = effort_ts_tidy, aes(x = Year, y = Norm_Effort, colour = Species), size=1) +
  geom_point(data = effort_ts_tidy, aes(x = Year, y = Norm_Effort), shape = 1,size = 1,colour = "black") +
    geom_line(data = effort_ts_tidy, aes(x = Year, y = Norm_Effort, colour = Species), size=1) +
    facet_wrap(~Species,scales="free_y",
             labeller = labeller(Species = species.labs)) +
  labs(fill = "Functional group") +
  theme(legend.title = element_text( size=3), legend.text=element_text(size=3)) + 
  geom_vline(aes(xintercept = 1961), linetype = "dashed") + 
  geom_vline(aes(xintercept = 2010), linetype = "dashed") +
  theme_classic() +
  theme(
    legend.position = "none"
  )

ggsave("plots/Effort_Forcing_low_res.tiff", plot = p_effort, units="cm", width=20, height=16, dpi =150)
```


```{r}
gear_params(projection_steady_F@params)$catchability
```


Results for presentaiton: 
- fished/unfished biomass. 
- growth plots
- barplots - comparing fished and unfished
- Variety of realised PPMRs
- Ecosystem indicators for mean of first 5 years to last 5 years - with and without fishing 
Time-average biomasses start and finish of time period and with/without fishing

```{r}

mizer::getCommunitySlope()
mizer::getProportionOfLargeFish() # large animals or select specific species
mizer::getMeanMaxWeight()
mizer::getMeanWeight()
```


## Mean weights
```{r}
years_comparison <- c("1961","2010")
# getMeanWeight(projection_steady_F, species = c("antarctic krill", "bathypelagic fishes", "shelf and coastal fishes", "squids", "toothfishes", "minke whales", "orca", "sperm whales", "baleen whales"))
getMeanMaxWeight(projection_steady_F, species = c("baleen whales"))[years_comparison,]

# years_comparison_noF <- c("0","50")
# getMeanWeight(projection_steady_noF, species = c("antarctic krill", "bathypelagic fishes", "shelf and coastal fishes", "squids", "toothfishes", "minke whales", "orca", "sperm whales", "baleen whales"))
getMeanMaxWeight(projection_steady_noF, species = c("baleen whales"))[years_comparison,]

getMeanMaxWeight(projection_steady_F, species = c("minke whales"))[years_comparison,]
getMeanMaxWeight(projection_steady_noF, species = c("minke whales"))[years_comparison,]

getMeanMaxWeight(projection_steady_F, species = c("sperm whales"))[years_comparison,]
getMeanMaxWeight(projection_steady_noF, species = c("sperm whales"))[years_comparison,]

```

```{r}
# getProportionOfLargeFish(projection_steady_F, species = c("antarctic krill"), min_w = 0.1, threshold_w = 2)
# getProportionOfLargeFish(projection_steady_noF, species = c("antarctic krill"), min_w = 0.1, threshold_w = 2)

getProportionOfLargeFish(projection_steady_F, species = c("shelf and coastal fishes"), min_w = 10, threshold_w = 1000)
getProportionOfLargeFish(projection_steady_noF, species = c("shelf and coastal fishes"), min_w = 10, threshold_w = 1000)
```

```{r}
getCommunitySlope(projection_steady_F)
getCommunitySlope(projection_steady_noF)

```
-1.017799  25.79908
-1.009425	 25.83120	









