---
title: "Prydz Bay mizer model calibration"
output: html_notebook
---

Set up the environment:
```{r}
# library(remotes)
# remotes::install_github("sizespectrum/therMizer")
# remotes::install_github("sizespectrum/mizerExperimental")
# library(therMizer)
library(mizerExperimental)
library(tidyverse)
```

Load latest mizer param object, which is a therMizer param object

```{r}
params <- readRDS("params_04_06_2024_v4.rds")
```

Inspect some key parameters
```{r}
species_params(params) |> select(erepro, R_max)
```

Let's run a simulation and inspect the dynamics
```{r}
sim <- project(params, t_max = 500)
plot(sim)
plotlyBiomass(sim)
plotlyBiomassRelative(sim)
```

There are still a lot of species with unstable biomass and what looks like some going extinct (orcas)

See if we can achieve a steady-state for the biomass

First we'll manually adjust erepro for any species that is above 1

```{r}
species_params(params)$erepro

species_params(params)$erepro <- ifelse(species_params(params)$erepro >= 1, 0.99, species_params(params)$erepro)

species_params(params)$erepro
```


```{r}
params <-
  steady(params, tol = 0.01, t_max = 200) %>% calibrateBiomass() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% 
  steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>%
  steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% 
  steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% 
  steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% 
  steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% 
  steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% 
  
  steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.005, t_max = 500) %>% matchBiomasses() %>%
  
  steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>%
  steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>% steady(tol = 0.0025, t_max = 500) %>% matchBiomasses() %>%
  
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses()
```



Let's run a simulation and inspect the dynamics
```{r}
sim <- project(params, t_max = 500)
plot(sim)
plotlyBiomass(sim)
plotlyBiomassRelative(sim)
```

```{r}
species_params(params)$erepro

species_params(params)$erepro <- ifelse(species_params(params)$erepro >= 1, 0.99, species_params(params)$erepro)

species_params(params)$erepro
```

```{r}
params <-
  steady(params, tol = 0.01, t_max = 200) %>% calibrateBiomass() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% 
  
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses()
```



Let's run a simulation and inspect the dynamics
```{r}
sim <- project(params, t_max = 1000)
plot(sim)
plotlyBiomass(sim)
plotlyBiomassRelative(sim)
```




```{r}
species_params(params)$erepro

# species_params(params)$erepro <- ifelse(species_params(params)$erepro >= 1, 0.99, species_params(params)$erepro)

# species_params(params)$erepro
```

Inspect ppmr and feeding kernel parameters
```{r}
species_params(params) |> select(pred_kernel_type, beta, ppmr_min, ppmr_max)
```

```{r}
params@species_params$pred_kernel_type[params@species_params$species == "minke whales"]  <- "box"
params@species_params$ppmr_min[params@species_params$species == "minke whales"]  <- 1e5
params@species_params$ppmr_max[params@species_params$species == "minke whales"] <-5e7
```


Inspect ppmr and feeding kernel parameters
```{r}
species_params(params) |> select(pred_kernel_type, beta, ppmr_min, ppmr_max)
```


```{r}
params <-
  steady(params, tol = 0.01, t_max = 200) %>% calibrateBiomass() %>% steady(tol = 0.01, t_max = 200) %>% matchBiomasses() %>% 
  
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>% steady(tol = 0.001, t_max = 1000) %>% matchBiomasses() %>%
  
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.0005, t_max = 5000) %>% matchBiomasses() %>%
  
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>% steady(tol = 0.00025, t_max = 5000) %>% matchBiomasses() %>%
  
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.0001, t_max = 10000) %>% matchBiomasses()
```


```{r}
params <-
  steady(params, tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses()
```



```{r}
params@species_params$ppmr_min[params@species_params$species == "baleen whales"]  <- 1e5
params@species_params$ppmr_max[params@species_params$species == "baleen whales"] <-5e6

params@species_params$ppmr_min[params@species_params$species == "minke whales"]  <- 1e5
params@species_params$ppmr_max[params@species_params$species == "minke whales"] <- 5e6
```



```{r}
params <-
  steady(params, tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.00001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>%
  steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses() %>% steady(tol = 0.000001, t_max = 10000) %>% matchBiomasses()
```


Let's run a simulation and inspect the dynamics
```{r}
sim <- project(params, t_max = 1000)
plot(sim)
plotlyBiomass(sim)
plotlyBiomassRelative(sim)
plotlyFeedingLevel(sim, include_critical = T)
```



```{r}
params <- readRDS("params_07_06_2024.rds")

params <- tuneParams(params)

# saveRDS(params, "params_07_06_2024.rds")
```


```{r}
sim <- project(params, t_max = 200)
plot(sim)
plotBiomassRelative(sim)
plotlyBiomass(sim)

params@species_params$erepro
```

```{r}

```


```{r}
getYieldVsF(params, species = "mesozooplankton")
```


```{r}
getYieldVsF(params, species = "baleen whales")
```


```{r}
# function to create yield curves:
makeSpeciesFmsyPlot<-function(speciesnum,params){
  #plotYieldVsF(params_pre, species = speciesnum,F_range = seq(0, 1, 0.05))
  # NOTE: not sure why above plot function does not work - wrote it what's sinside function below instead
  curve <- getYieldVsF(params, species = speciesnum, F_range = seq(0, 1, 0.1))
  pl <- ggplot(curve, aes(x = F, y = yield)) + geom_line() + 
        xlab("Fishing mortality (1/yr)") + ylab("Yield") + ggtitle(params@species_params$species[speciesnum])
  return(pl)
}

#check plots for each species - they should all look dome-shaped with a reasonable Fmsy

# check plot for EACH species MANUALLY

makeSpeciesFmsyPlot(speciesnum = 2,params)
makeSpeciesFmsyPlot(speciesnum = 19,params)

# Or loop through and then scroll though all  plots
specs<-unique(species_params(params)$species)
plots <- list()
for (i in 1:length(specs)) plots[[i]]<-makeSpeciesFmsyPlot(speciesnum = i,params)
# ## you can then manually go through each plot, e.g. change the number in [[]] below (or save them)
# plots[[4]]
```







