---
title: "Trying out therMizer on the North Sea model"
author: "Samik Datta"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I'm using Hiromi Beran's .Rmd document `Testing therMizer.Rmd` as my insipiration, which itself uses Phoebe Woodworth-Jefcoats' introduction to therMizer (<https://github.com/sizespectrum/therMizer>).

## Load packages

```{r}
# remotes::install_github("sizespectrum/therMizer") # if needed
library(therMizer)
library(mizer)
library(tidyverse)
```

## therMizer model parameters required

### Species and temperature parameters

-   The usual mizer species parameters

-   `temp_min` - lower bound of a species' thermal tolerance limit (source from `rfishbase` and the literature e.g. tagging studies)

-   `temp_max` - upper bound of a species' thermal tolerance limit (source from `rfishbase` and the literature e.g. tagging studies)

-   `realm` - depth strata which species inhabit and for which temperatures will be provided. Names need to match the realm names provided in the `ocean_temp` array you're using (examples include *epipelagic*, *mesopelagic*, *surface*, *bottom, diel vertical migration day, diel vertical migration night*)

-   `vertical migration` - simulates the proportion of time that a given size of a given species spends in a given realm (realm being depth stratum)

-   Has the dimensions `realm` x `sp` (species) x `w` (size)

-   Values can range from 0 to 1, and must sum to 1 across all realms for each species and size

-   `exposure` - links `vertical_migration` to `ocean_temp`

-   Has the dimensions of `realm` x `sp` (species)

-   The values are 1 for the realms to which a species is exposed and 0 elsewhere

-   In theory, you could set all values to 1, and as long as `vertical_migration` is constructed correctly, get the same results (because when multiple by `exposure` the result would still be 0 for realms in which species spend no time).

-   Realm names and order have to match those used in `vertical_migration` and `ocean_temp`

-   `ocean_temp` - an array with temperature(s) in degrees Celsius for each `realm`. Can be a vector, if temperature is constant over time, or an array for dynamic temperatures. If you're using time-varying temperatures, the array will have the dimensions of `time` x `realm`

-   `n_pp` - an array that has numerical plankton abundance for each size class. therMizer will convert these abundances to densities for use within mizer. Can be a vector, if the abundances are constant over time, or an array for a dynamic resource. If you're using time-varying plankton, the array will have the dimensions of `time` x `w`

### Temperature functions

Temperature functions in therMizer overwrite mizer's default rate functions, replacing them with custom functions using the new set of parameters.

-   `therMizerEncounter` affects the encounter rate

-   `therMizerPredRate` affects the predation rate

-   `therMizerEReproAndGrowth` takes care of the maintenance metabolism

These functions can be disabled by settings the arguments `aerobic_effect` and `metabolism_effect` to `FALSE` for encounter and predation rates and for metabolism, respectively.

These functions can also be overwritten by the user using `setRateFunction()`. *See GitHub example.*

*See GitHub example for full sample code. Here I add therMizer into the mini model.*

## Adding therMizer to the North Sea model 

Read in the North Sea model.

```{r}
#Read in mini model
ns_model <- newMultispeciesParams(species_params = NS_species_params,
                                  interaction = NS_interaction)

#Assign thermal tolerance limits to the four species
species_params(ns_model)$temp_min <- seq(3, 6, length.out = dim(species_params(ns_model))[1])
species_params(ns_model)$temp_max <- seq(25, 28, length.out = dim(species_params(ns_model))[1])

# Have a peek at the first few rows
head(species_params(ns_model))

sim = project(ns_model) # run simulation
plotSpectra(sim) # check size spectrum
plotBiomass(sim) # check biomasses are constant

```

Now we will set up the `vertical_migration` array. We'll assume that:
- species with `w_mat` under 500g stay mostly in the top 50m until maturity, after which they move to 50-500m;
- species with `w_mat` between 500g and 1kg stay mostly in the 50-500m range until maturity, after which they move to the bottom;
- species with `w_mat` larger than 1kg have a 30/30/40 distribution for all weights.

```{r}
realm_names <- c("0-50m", "50-500m", "bottom")

species_names <- as.character(ns_model@species_params$species)

sizes <- ns_model@w

#Create the vertical migration array and fill it
vertical_migration_array <- array(0.1, dim = (c(length(realm_names),
                                                length(species_names),
                                                length(sizes))),
                                  dimnames = list(realm = realm_names, sp = species_names, w = signif(sizes, 4))) #realm x species x size

#The vertical migration array we have created is currently filled with 0.1

#Now we will properly fill the vertical migration array

# w_mat < 500g - 0.7/0.2/0.1 before maturity, 0.2/0.7/0.1 after
# 500g < w_mat <= 1kg - 0.2/0.7/0.1 before maturity, 0.1/0.2/0.7 after
# w_mat >= 1kg - 0.3/0.4/0.3 for all weights

for (j in 1:length(species_names)) {
  
  if (ns_model@species_params$w_mat[j] < 500) {
    # w_mat < 500g - 0.7/0.2/0.1 before maturity, 0.2/0.7/0.1 after
    vertical_migration_array [1, j, sizes < ns_model@species_params$w_mat[j]] <- 0.7 # mostly in top layer
    vertical_migration_array [2, j, sizes < ns_model@species_params$w_mat[j]] <- 0.2 # a bit in the middle
    vertical_migration_array [1, j, sizes >= ns_model@species_params$w_mat[j]] <- 0.2 # a bit at the top
    vertical_migration_array [2, j, sizes >= ns_model@species_params$w_mat[j]] <- 0.7 # mostly in the middle
  } else if (ns_model@species_params$w_mat[j] >= 500 & ns_model@species_params$w_mat[j] < 1000) {
    # 500g < w_mat <= 1kg - 0.1/0.7/0.2 before maturity, 0.1/0.2/0.7 after
    vertical_migration_array [2, j, sizes < ns_model@species_params$w_mat[j]] <- 0.7 # mostly in middle layer
    vertical_migration_array [3, j, sizes < ns_model@species_params$w_mat[j]] <- 0.2 # a bit at the bottom
    vertical_migration_array [2, j, sizes >= ns_model@species_params$w_mat[j]] <- 0.2 # a bit in the middle
    vertical_migration_array [3, j, sizes >= ns_model@species_params$w_mat[j]] <- 0.7 # mostly at the bottom
  } else if (ns_model@species_params$w_mat[j] >= 1000) {
    # w_mat >= 1kg - 0.3/0.4/0.3 for all weights
    vertical_migration_array [1, j, ] <- 0.3 # in the top layer
    vertical_migration_array [2, j, ] <- 0.4 # in the middle
    vertical_migration_array [3, j, ] <- 0.3 # at the bottom
  }
}

```

Using the same scenario, everything is present in all layers, so the `exposure` array is all 1s.

```{r}
exposure_array <- array(1, dim = (c(length(realm_names),
                                    length(species_names))), 
                        dimnames = list (realm = realm_names, sp = species_names)) #realm x species

exposure_array
```

Create the temperatures for each realm.

```{r}
#Create temperature array and fill it
times <- 0:100

ocean_temp_array <- array (NA, dim = c(length(times),
                                       length(realm_names)),
                           dimnames = list(time = times, realm = realm_names))

temp_inc <- 0
for (i in 1:length(times)) {
  ocean_temp_array[i, ] <- c(21 + temp_inc, 20 + temp_inc, 19 + temp_inc)
}
```

Create a dynamic, time-varying resource spectra.

```{r}
x <- ns_model@w_full
slope <- -1
intercept <- -5

#Create resource array and fill it
n_pp_array <- array (NA, dim = c(length(times), 
                                 length(x)),
                     dimnames = list(time = times, w = signif(x,3)))



for (i in 1:length(times)) {
  #Add some noise around the slope and intercept as we fill the array
  n_pp_array[i,] <- (slope*runif(1, min = 0.95, max = 1.05)*log10(x)) + (intercept*runif(1, min = 0.95, max = 1.05))
}
```

## Running a scenario

The `upgradeTherParams` function combines a standard `mizerParams` object with the therMizer objects described above.

```{r}
ns_therm_model <- upgradeTherParams(ns_model,
                                    ocean_temp_array = ocean_temp_array,
                                    n_pp_array = n_pp_array,
                                    vertical_migration_array = vertical_migration_array,
                                    exposure_array = exposure_array,
                                    aerobic_effect = TRUE,
                                    metabolism_effect = TRUE)
```

Let's look at the thermal tolerances of all the species in the model.

```{r}
plotThermPerformance(ns_therm_model)
```

Now let us have a look what happens when we run a scenario with these parameters.

```{r}

sim <- therProject(ns_therm_model) 
plotSpectra(sim)
plotBiomass(sim)

```

Stuff is crashing out like nobody's business! This needs sorting.

