---
title: "Exploring the IWC catch database"
author: "Kieran Murphy"
date:  "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

To do:

Add in krill fishing `fish_catch.csv` 

Model domain: 1.474341e+12

Seaaroundus for fish and krill catch

PNAS South China Sea logistic effort curve

Match toothfish fishing effort to match McCormack et al. 2020 as a starting point?

```{r}
library(tidyverse)
library(mizerExperimental)
```


```{r}
# # df_hist_catch <- read.csv("effort_histsoc_1841_2010_regional_models.csv")# historical catch from seaaroundus and fishmip
# 
# df_hist_PB <- df_hist_catch %>% 
#   filter(region == "Prydz.Bay")
# 
# # write.csv(df_hist_PB, "effort_histsoc_1841_2010_regional_models_Prydz.Bay.csv")
# 
# 
# glimpse(df_hist_PB)
# unique(df_hist_PB$FGroup)
# 
# df_hist_PB_krill <- df_hist_PB %>% 
#   filter(FGroup == "krill")
# 
# unique(df_hist_PB_krill$Year)

```

```{r}
df_Prydz.Bay <- read.csv("fish_catch.csv") # seaaroundus/FishMIP fish catch from Prydz Bay model domain

df_krill_raw <- df_Prydz.Bay %>% 
  filter(FGroup == "krill") %>% #Extract only the krill data
  group_by(Year,SAUP,) %>% 
  mutate(catch_comb = sum(Reported,IUU,Discards)) %>%  # Add the three catch columns together to create a total catch variable. Need to determine the unit catch is reported in
  group_by(Year,FGroup) %>% 
  summarise(catch_total = sum(catch_comb))


ggplot(df_krill_raw, aes(x=Year, y=catch_total))+
  geom_line()
  
sum(df_krill_raw$catch_total) # 98579.21

# Time series 1974 - 1996 = 22 years

# 98579.21/22 = 4480.873
```


```{r}
df_catch <- readRDS("ind_catch_weight_BanzareBank_1930_2019.rds")

df_catch <- as.data.frame(df_catch)

df_yield <- dplyr::select(df_catch, Weight_kg, Species, Year, Length)

glimpse(df_yield)
```

```{r}
catch_tidy <- df_yield %>% 
  group_by(Species) %>% 
  mutate(new_bin = cut(Length, breaks= seq(0,3020, by=10), include.lowest = TRUE, labels=F)) %>% 
  group_by(Species, new_bin) %>% 
  summarise(catch = n()) %>% 
  mutate(length = new_bin * 10)

catch_tidy$dl <- 10
```


```{r}
catch_lengths <- catch_tidy %>% 
  filter(!Species == "Bottlenose") %>% 
dplyr::mutate(species = case_when(Species == "Antarctic Minke" ~ "minke whales",
                             Species == "Sperm" ~ "sperm whales",
                             Species == "Killer" ~ "orca",
                             Species == "Baleen" ~ "baleen whales")) %>% 
  dplyr::select(species,catch,length,dl)

glimpse(catch_lengths)
# saveRDS(catch_lengths, "catch_lengths.rds")
```


```{r}
yield_summary <- df_yield %>% 
  group_by(Year, Species) %>% 
  summarise(yield_kg = sum(Weight_kg),
            yield_g = yield_kg * 1000)

glimpse(yield_summary)
```

```{r}
mean_yield_observed <- yield_summary %>% 
  group_by(Species) %>% 
  summarise(yield_observed_v1 = mean(yield_g),
            yield_observed = sum(yield_g)/89) # total values by each species divided by number of years in the time series

glimpse(mean_yield_observed)
```

[1] "euphausiids"              "mesopelagic fishes"       "bathypelagic fishes"      "shelf and coastal fishes" "flying birds"            
 [6] "small divers"             "squids"                   "toothfishes"              "leopard seals"            "medium divers"           
[11] "large divers"             "minke whales"             "orca"                     "sperm whales"             "baleen whales" 


```{r}
# obs_yield <- c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 2175476136, 19923729, 4062177445, 50341696055) # to add a yield_observed column in species_params
obs_yield <- c(NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 2175476136, 19923729, 4062177445, 50341696055) # to add a yield_observed column in species_params
# obs_yield_m2 <- obs_yield/1.474341e+12
```



```{r}
params <-  readRDS("stage1_steady_vXX.rds")

params <- setParams(setFishing(params, initial_effort = 0.2))

species_params(params)$yield_observed <- obs_yield

species_params(params) |> dplyr::select(species, yield_observed)
```

Update max and mat size for orca
```{r}
species_params(params)$w_max[13]
species_params(params)$w_mat[13]

species_params(params)$w_max[13] <- 10628034
species_params(params)$w_mat[13] <- 3198855

species_params(params)$w_max[13]
species_params(params)$w_mat[13]
 
```

Update max size for leopard seal
```{r}
species_params(params)$species
species_params(params)$w_max[9]
species_params(params)$w_mat[9]

species_params(params)$w_max[9] <- 545875.2

species_params(params)$w_max[9]
 
```


```{r}
params <- steady(params)
```


```{r}
params <- params |>
    calibrateYield() |> steady() |> matchBiomasses() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
  calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> matchBiomasses() |> steady()

plotBiomassVsSpecies(params)
```
```{r}
params_tuned_v1 <- tuneParams(params)

params_v2 <- steady(params_tuned_v1)

params_v2@species_params$erepro

params_tuned_v2 <- tuneParams(params_v2)

params_v3 <- steady(params_tuned_v2)

params_v3@species_params$erepro

plotlySpectra(params_v3)
```
Save species params to work with in therMizer workshop
```{r}
# saveRDS(params_v3, "params_16_March_2023.rds")
```


```{r}
gear_params(params)
```

Adjust catchability to only select fishing on species with catch data
```{r}
gear_params(params)$catchability <-  c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1)
```


Find minimum sizes caught to adjust knife_edge_size

Exclude bottlenose for now, something not right Bottlenose whales, not dolphins as I previously thought

```{r}
df_catch_size <- df_yield %>% 
  group_by(Species) %>% 
  summarise(median_size = median(Weight_kg),
            median_size_g = median_size * 1000)

df_catch_size
```

Catchability: 1 value 
Effort changes through time (voyage frequency and length)

```{r}
gear_params(params)$gear <-  c("Main", "Main", "Main", "Main", "Main", "Main",
                                   "Main", "Main", "Main", "Main", "Main", "Main",
                                   "Main", "Main", "Main")

gear_params(params)$l50 <-  c(1, 5, 5, 5, 10, 10, 10, 20, 20, 20, 20, 850, 600, 1500, 2200)

gear_params(params)$l25 <-  c(0.8, 4, 4, 4, 8, 8, 8, 16, 16, 16, 16, 800, 500, 1400, 1500)

gear_params(params)$sel_func <-  c("sigmoid_length", "sigmoid_length", "sigmoid_length", "sigmoid_length", "sigmoid_length", "sigmoid_length",
                                   "sigmoid_length", "sigmoid_length", "sigmoid_length", "sigmoid_length", "sigmoid_length", "sigmoid_length",
                                   "sigmoid_length", "sigmoid_length", "sigmoid_length")

gear_params(params)
# gp <- gear_params(params)

# saveRDS(gp, "gear_params_v1.rds")
```

```{r}
params <- setParams(setFishing(params, initial_effort = 0.2))

params_v1 <- steady(params)

model <- project(params_v1, effort=1)
plot(model)
```


```{r}
plotYieldVsSize(model, species = "baleen whales",
                x_var = "Length")

plotYieldVsSize(model, species = "sperm whales",
                x_var = "Length")

plotYieldVsSize(model, species = "minke whales",
                x_var = "Length")

plotYieldVsSize(model, species = "orca",
                x_var = "Length")
```

```{r}
plotYieldVsSpecies(params_v1)
model_c1 <- calibrateYield(params_v1)
plotYieldVsSpecies(model_c1)

```


```{r}
scaling_factor <- initialN(model_c1)[1, 1] / initialN(model)[1, 1]
scaling_factor
```


```{r}
# Rescale the observed biomass values by that factor
species_params(model_c1)$biomass_observed <- 
    species_params(model_c1)$biomass_observed * scaling_factor
```


```{r}
plotBiomassVsSpecies(model_c1)
model_c1 <- matchBiomasses(model_c1)
model_c1 <- steady(model_c1)
plotBiomassVsSpecies(model_c1)

```

```{r}
model_s1 <- steady(model_c1)
plotBiomassVsSpecies(model_s1)

```


```{r}
plotYieldVsSize(model_s1, species = "baleen whales", catch = catch_lengths, 
                x_var = "Length")
```
```{r}
plotYieldVsSize(model_s1, species = "sperm whales", catch = catch_lengths, 
                x_var = "Length")
```

```{r}
plotYieldVsSize(model_s1, species = "minke whales", catch = catch_lengths, 
                x_var = "Length")
```


```{r}
plotYieldVsSize(model_s1, species = "orca", catch = catch_lengths, 
                x_var = "Length")
```

```{r}

# Modify the l50 and l25 for breams
gear_params(model_s1)["orca, Main", "l25"] <- 550
gear_params(model_s1)["orca, Main", "l50"] <- 650

#check the match between model and observed catch size distributions
plotYieldVsSize(model_s1, species = "orca", catch = catch_lengths, 
                x_var = "Length")


```

```{r}
model_s2 <- steady(model_s1)
plotBiomassVsSpecies(model_s2)
```



```{r}
model_tuned_v1 <- tuneParams(model_s2)

model_tuned_v1 <- tuneParams(model_tuned_v1)

model_tuned_v1 <- tuneParams(model_tuned_v1)

model_tuned_v1 <- tuneParams(model_tuned_v1)
```





```{r}
model_tuned_v2 <- tuneParams(model_tuned_v1)

model_tuned_v2 <- tuneParams(model_tuned_v2)

# saveRDS(model_tuned_v2, "model_tuned_catch_vXX.rds")

```





```{r}
params_cy_v2 <- model_tuned_v2 |>
    calibrateYield() |> steady() |> matchBiomasses() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
  calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> steady() |>
    calibrateYield() |> steady() |> calibrateYield() |> matchBiomasses() |> steady()

plotBiomassVsSpecies(params_cy_v2)
```



```{r}
model_proj_v2 <- project(params_cy_v2, t_max = 200)

plot(model_proj_v2)
```


```{r}
plotYieldVsSize(model_proj_v2, species = "baleen whales", catch = catch_lengths, 
                x_var = "Length")
```


```{r}
plotYieldVsSize(model_proj_v2, species = "sperm whales", catch = catch_lengths, 
                x_var = "Length")
```


Effort through time: sigmoidal size selectivity curve: hand calibrate or fit curve to it with (week 2 of course 1plotYieldvsSize) Size selectivity Approach

Or stylised effort function and test multiple hypothesis of what that looked like to match the catch timeseries: effort evolution through time approach

Progressive depletion of the whale abundance (existing papers)
Caused by collapse?
Shift in effort or did they reach a tipping point or was it a progressive reduction of larger

Does a shift in body size distributions emerge from the inclusion of fishing

Shift from one species to another versus equal effort and then the vulnerability would lead to differences in response.

Dynamics of effort curve and what curve 'shape' matches the catch data the best? Is it a long drawn out max effort or a rapid shift (i.e., steep inflection point) 

To add stochasticity: check github (https://github.com/juliablanchard/fisheries_stability)

```{r}
df_yield %>% 
  filter(!Species == "Bottlenose") %>%
ggplot(aes(x=Weight_kg, fill=Species)) + 
 geom_histogram(position = 'identity') +
  facet_wrap(~Species, scales = "free")

```













