---
title: "R Notebook"
output: html_notebook
---

Prydz Bay model domain 1,433,028 km2

From km2 to m2 1433028 * 1e6 = 1.433028e+12

Check the area for the model domain that gfdl ISIMIP data has been extracted for Prydz Bay

```{r}
# Load surface temperature data
df_gfdl_tos <- read.csv("FishMIP_Plankton_Forcing/gfdl-mom6-cobalt2_obsclim_tos_15arcmin_prydz-bay_monthly_1961_2010.csv") 

head(df_gfdl_tos)
```


```{r}
GFDL_tos <- df_gfdl_tos %>%
  select(!c(lat,lon,area_m2)) %>% # remove lat, lon and area_m2
  melt() %>% # surface temperature of each area for all grid cells
  group_by(variable) %>% # group by month
  summarise(tos = mean(value)) # sum the carbon across all grid cells by each month

glimpse(GFDL_tos)
head(GFDL_tos)
```
The model domain area (1.433028e+12 m^2) from previous notes is a close match to the model domain area used for Prydz Bay data extraction (1.454146e+12 m^2)

Check the size of Phoebe's model domain, as the value for total carbon in moles is two orders of magnitude higher for the Prydz Bay model domain, at least according to my current calculations

```{r}
# Load surface temperature data for Hawai'i model
df_gfdl_tos_hawaii <- read.csv("gfdl-mom6-cobalt2_obsclim_tos_15arcmin_hawaiian-longline_monthly_1961_2010.csv") 

head(df_gfdl_tos_hawaii)
```


```{r}
df_hawaii_area <- df_gfdl_tos_hawaii %>% 
  select(area_m2) %>% 
  melt() %>% 
  group_by(variable) %>% 
  summarise(total_area_m2 = sum(value))

head(df_hawaii_area)
```







