---
title: "Trying out therMizer on the Prydz Bay SO model"
author: "Kieran edited from Samik Datta's original"
date: "`r format(Sys.Date(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

I'm using Hiromi Beran's .Rmd document `Testing therMizer.Rmd` as my insipiration, which itself uses Phoebe Woodworth-Jefcoats' introduction to therMizer (<https://github.com/sizespectrum/therMizer>).

## Load packages

```{r}
# install.packages("remotes")
# remotes::install_github("sizespectrum/therMizer@dev")
remotes::install_github("sizespectrum/therMizer") # if needed
library(therMizer)
library(mizer)
library(tidyverse)
```

## therMizer model parameters required

### Species and temperature parameters

-   The usual mizer species parameters

-   `temp_min` - lower bound of a species' thermal tolerance limit (source from `rfishbase` and the literature e.g. tagging studies)

-   `temp_max` - upper bound of a species' thermal tolerance limit (source from `rfishbase` and the literature e.g. tagging studies)

-   `realm` - depth strata which species inhabit and for which temperatures will be provided. Names need to match the realm names provided in the `ocean_temp` array you're using (examples include *epipelagic*, *mesopelagic*, *surface*, *bottom, diel vertical migration day, diel vertical migration night*)

-   `vertical migration` - simulates the proportion of time that a given size of a given species spends in a given realm (realm being depth stratum)

-   Has the dimensions `realm` x `sp` (species) x `w` (size)

-   Values can range from 0 to 1, and must sum to 1 across all realms for each species and size

-   `exposure` - links `vertical_migration` to `ocean_temp`

-   Has the dimensions of `realm` x `sp` (species)

-   The values are 1 for the realms to which a species is exposed and 0 elsewhere

-   In theory, you could set all values to 1, and as long as `vertical_migration` is constructed correctly, get the same results (because when multiple by `exposure` the result would still be 0 for realms in which species spend no time).

-   Realm names and order have to match those used in `vertical_migration` and `ocean_temp`

-   `ocean_temp` - an array with temperature(s) in degrees Celsius for each `realm`. Can be a vector, if temperature is constant over time, or an array for dynamic temperatures. If you're using time-varying temperatures, the array will have the dimensions of `time` x `realm`

-   `n_pp` - an array that has numerical plankton abundance for each size class. therMizer will convert these abundances to densities for use within mizer. Can be a vector, if the abundances are constant over time, or an array for a dynamic resource. If you're using time-varying plankton, the array will have the dimensions of `time` x `w`

### Temperature functions

Temperature functions in therMizer overwrite mizer's default rate functions, replacing them with custom functions using the new set of parameters.

-   `therMizerEncounter` affects the encounter rate

-   `therMizerPredRate` affects the predation rate

-   `therMizerEReproAndGrowth` takes care of the maintenance metabolism

These functions can be disabled by settings the arguments `aerobic_effect` and `metabolism_effect` to `FALSE` for encounter and predation rates and for metabolism, respectively.

These functions can also be overwritten by the user using `setRateFunction()`. *See GitHub example.*

*See GitHub example for full sample code. Here I add therMizer into the mini model.*

## Adding therMizer to the North Sea model 

Read in the latest SO model.

```{r}
#Read in model

so_params <- readRDS("params_16_March_2023.rds")

```


```{r}
params_tuned_v1 <- tuneParams(so_params)
```


Adding min and max temp for each group

Using sealifebase for all non-fish species/groups. In most cases, there are no listed Temperature values in the 'Environment' section, like there is for more data-rich species, e.g., cod

Instead, there are values at the bottom of the sealifebase pages in 'Estimates of some properties based on models' with 'Preferred temperature' estimates from the AquaMaps model-based approach (https://www.gbif.org/tool/81356/aquamaps-predicted-range-maps-for-aquatic-species)  

[1] "euphausiids":
Antarctic krill - (Preferred temperature (Ref. 115969): -1.5 - 1.6, mean 0.2 (based on 535 cells))
Crystal krill - (Preferred temperature (Ref. 115969): -1.8 - 1.1, mean -1.6 (based on 3588 cells)) 
Thysanoessa macrura - (Preferred temperature (Ref. 115969): -1.2 - 2.8, mean 0.2 (based on 24 cells). Not included due to small number of cells and low contribution to biomass in this model domain
[2]"mesopelagic fishes" - Electrona antarctica(nothing on fishbase)    
[3]"bathypelagic fishes"      
[4]"shelf and coastal fishes":
Marbled rockcod (-1 - 5)
[5]"flying birds":
Southern giant petrel (Preferred temperature (Ref. 115969): 2.1 - 15.6, mean 10.1 (based on 533 cells));
Arctic tern (Sterna paradisaea) (Preferred temperature (Ref. 115969): 0.4 - 13.3, mean 6.8 (based on 2468 cells))
Thalassarche chrysostoma (Grey-headed petrel)  (Preferred temperature (Ref. 115969): 1.3 - 9.1, mean 5.2 (based on 263 cells))
South polar skua (Preferred temperature (Ref. 115969): 0.2 - 18.2, mean 6.4 (based on 1318 cells))
Diomedea exulans (Wandering albatross), (Preferred temperature (Ref. 115969): 0.5 - 12, mean 4.6 (based on 1528 cells)_
[6]"small divers" - no info for adelie, gentoo or crested penguins on sealifebase. Guess based on medium divers       
[7]"squids" - based on Psychroteuthis glacialis (sealifebase)                   
[8]"toothfishes" - no data on fishbase            
[9]"leopard seals" - Preferred temperature (Ref. 115969): -1.9 - 1.4, mean -0.5 (based on 2109 cells).          
[10] "medium divers":
Crabeater Seals - (Preferred temperature (Ref. 115969): -0.4 - 1, mean 0.1 (based on 9397 cells))
Ross seals (Preferred temperature (Ref. 115969): -1.8 - 1.1, mean -1.5 (based on 2110 cells));
Weddell seals (Preferred temperature (Ref. 115969): -1.8 - 1, mean -1.6 (based on 6026 cells));
Hourglass dolphin (Preferred temperature (Ref. 115969): 0.1 - 2, mean 0.9 (based on 11222 cells)) # Not included in the biomass
King penguin (no data on sealifebase)
Emperor penguin (no data on sealifebase)
[11]"large divers" - Southern elephant seals only in this group now that Sperm whales are also seperated         
[12]"minke whales"             
[13]"orca"                     
[14]"sperm whales"             
[15]"baleen whales" 

```{r}
species_params(so_params)$temp_min <- c(-1.8, # euphausiids
                                     -2, # mesopelagic fishes
                                     -2, # bathypelagic fishes
                                     -1, # "shelf and coastal fishes" 
                                     0.2, # "flying birds"
                                     -0.4, # "small divers"             
                                     -0.8, # "squids"                  
                                     -2, # "toothfishes"              
                                     -1.9, # "leopard seals"           
                                     -0.4, # "medium divers"            
                                     0.1, # "large divers"             
                                     0.2, # "minke whales"             
                                     0.3, # "orca"                     
                                     0.3, # "sperm whales"             
                                     0.2) # "baleen whales"

species_params(so_params)$temp_max <- c(1.6, # euphausiids
                                     5, # mesopelagic fishes
                                     5, # bathypelagic fishes
                                     5, # "shelf and coastal fishes" 
                                     18.2, # "flying birds"
                                     1, # "small divers"             
                                     1.1, # "squids"                  
                                     5, # "toothfishes"              
                                     1.4, # "leopard seals"           
                                     1, # "medium divers"            
                                     1.6, # "large divers"             
                                     7, # "minke whales"             
                                     13.1, # "orca"                     
                                     3.8, # "sperm whales"             
                                     10.2) # "baleen whales"
```


```{r}
# Have a peek at the first few rows
head(species_params(so_params))

sim = project(so_params, t_max = 200) # run simulation
plotSpectra(sim) # check size spectrum
plotBiomass(sim) # check biomasses are constant

```
Check the existing depth variable included in `species_params`

```{r}
unique(so_params@species_params$water.column.use)
print("min depth")
so_params@species_params$min_depth
print("max depth")
so_params@species_params$max_depth
so_params@species_params$species
so_params@species_params$water.column.use
```

Now we will set up the `vertical_migration` array. We'll assume that:
- species with `w_mat` under 500g stay mostly in the top 50m until maturity, after which they move to 50-500m;
- species with `w_mat` between 500g and 1kg stay mostly in the 50-500m range until maturity, after which they move to the bottom;
- species with `w_mat` larger than 1kg have a 30/30/40 distribution for all weights.

```{r}
params <- so_params

realm_names <- c("upper50m","bottom","DVM_day","DVM_night","ontogenetic","diving")
species_names <- as.character(params@species_params$species)
sizes <- params@w

# Create the vertical migration array and fill it
vertical_migration_array <- array(0, dim = (c(length(realm_names), 
                                  length(species_names), length(sizes))), 
                                  dimnames = list(realm = realm_names, sp = species_names, 
                                  w = signif(sizes,3))) # realm x species x size

upp <- which(realm_names == "upper50m") # 0 - 50m average
btm <- which(realm_names == "bottom") # sea floor
DVMd <- which(realm_names == "DVM_day") # 200 - 500m average
DVMn <- which(realm_names == "DVM_night") # 0 - 100m average
ont <- which(realm_names == "ontogenetic") # deeper after w_mat
dive <- which(realm_names == "diving") # evenly spread across entire realm

# Set all sizes below w_mat for euphausiids to "upper50m" and all sizes above w_mat to "bottom
spA <- which(species_names == "euphausiids")
vertical_migration_array[upp, spA, sizes < params@species_params$w_mat[spA]] <- 1
vertical_migration_array[btm, spA, sizes >= params@species_params$w_mat[spA]] <- 1

# Have mesopelagic fishes split its time equally using DVM
spB <- which(species_names == "mesopelagic fishes")
vertical_migration_array[DVMd, spB, ] <- 0.5
vertical_migration_array[DVMn, spB, ] <- 0.5

# Set all sizes below w_mat for bathypelagic fishes to "upper50m" and split time equally using DVM for sizes above w_mat
spC <- which(species_names == "bathypelagic fishes")
vertical_migration_array[upp, spC, sizes < params@species_params$w_mat[spC]] <- 1
vertical_migration_array[DVMd, spC, sizes >= params@species_params$w_mat[spC]] <- 0.5
vertical_migration_array[DVMn, spC, sizes >= params@species_params$w_mat[spC]] <- 0.5

# Set all sizes below w_mat for shelf and coastal fishes to "upper50m" and all sizes above w_mat to "bottom"
spD <- which(species_names == "shelf and coastal fishes")
vertical_migration_array[upp, spD, sizes < params@species_params$w_mat[spD]] <- 1
vertical_migration_array[btm, spD, sizes >= params@species_params$w_mat[spD]] <- 1

# Have flying birds spend all their time in "diving"
spE <- which(species_names == "flying birds")
vertical_migration_array[dive, spE, ] <- 1

# Have small divers spend all their time in "diving"
spF <- which(species_names == "small divers")
vertical_migration_array[dive, spF, ] <- 1

# Set all sizes below w_mat for squids to "upper50m" and split time equally using DVM for sizes above w_mat
spG <- which(species_names == "squids")
vertical_migration_array[upp, spG, sizes < params@species_params$w_mat[spG]] <- 1
vertical_migration_array[DVMd, spG, sizes >= params@species_params$w_mat[spG]] <- 0.5
vertical_migration_array[DVMn, spG, sizes >= params@species_params$w_mat[spG]] <- 0.5

# Set all sizes below w_mat for toothfishes to "upper50m" and all sizes above w_mat to "bottom
spH <- which(species_names == "toothfishes")
vertical_migration_array[upp, spH, sizes < params@species_params$w_mat[spH]] <- 1
vertical_migration_array[btm, spH, sizes >= params@species_params$w_mat[spH]] <- 1

# Have leopard seals spend all their time in "diving"
spI <- which(species_names == "leopard seals")
vertical_migration_array[dive, spI, ] <- 1

# Have medium divers spend all their time in "diving"
spJ <- which(species_names == "medium divers")
vertical_migration_array[dive, spJ, ] <- 1

# Have large divers spend all their time in "diving"
spK <- which(species_names == "large divers")
vertical_migration_array[dive, spK, ] <- 1

# Have minke whales spend all their time in "diving"
spL <- which(species_names == "minke whales")
vertical_migration_array[dive, spL, ] <- 1

# Have orca spend all their time in "diving"
spM <- which(species_names == "orca")
vertical_migration_array[dive, spM, ] <- 1

# Have sperm whales spend all their time in "diving"
spN <- which(species_names == "sperm whales")
vertical_migration_array[dive, spN, ] <- 1

# Have baleen whales spend all their time in "diving"
spO <- which(species_names == "baleen whales")
vertical_migration_array[dive, spO, ] <- 1

```





Using the same scenario, everything is present in all layers, so the `exposure` array is all 1s.

```{r}
exposure_array <- array(1, dim = (c(length(realm_names),
                                    length(species_names))), 
                        dimnames = list (realm = realm_names, sp = species_names)) #realm x species

exposure_array
```

Create the temperatures for each realm.


```{r}
# Create temperature array and fill it
times <- 0:500
ocean_temp_array <- array(NA, dim = c(length(times), length(realm_names)), 
                    dimnames = list(time = times, realm = realm_names))
temp_inc <- 0
for (i in 1:501) {
  ocean_temp_array[i,] <- c(2 + temp_inc, -2 + temp_inc, 2 + temp_inc, -1 + temp_inc, -1 + temp_inc, -1 + temp_inc)
  temp_inc <- temp_inc + 0.01
}

```

Create a dynamic, time-varying resource spectra.

```{r}

x <- params@w_full
slope <- -1
intercept <- -5

# Create resource array and fill it
n_pp_array <- array(NA, dim = c(length(times), length(x)), 
                    dimnames = list(time = times, w = signif(x,3)))

for (i in 1:501) {
  # Add some noise around the slope and intercept as we fill the array
  n_pp_array[i,] <- (slope * runif(1, min = 0.95, max = 1.05) * log10(x)) + 
                    (intercept * runif(1, min = 0.95, max = 1.05))
}


```

## Running a scenario

The `upgradeTherParams` function combines a standard `mizerParams` object with the therMizer objects described above.

```{r}
so_therm_model <-  upgradeTherParams(params, 
                            temp_min = temp_min,
                            temp_max = temp_max,
                            ocean_temp_array = ocean_temp_array,
                            n_pp_array = n_pp_array, 
                            vertical_migration_array = vertical_migration_array,
                            exposure_array = exposure_array, 
                            aerobic_effect = TRUE, 
                            metabolism_effect = TRUE)
                                
```

Let's look at the thermal tolerances of all the species in the model.

```{r}
plotThermPerformance(so_therm_model)
```

Now let us have a look what happens when we run a scenario with these parameters.

```{r}
rownames(so_therm_model@other_params$other$ocean_temp) <- 1:501

so_therm_model@other_params$other$ocean_temp


sim <- therProject(so_therm_model) 

sim <- project(params, 
               t_start = as.numeric(1),
               t_max = as.numeric(500))
plotSpectra(sim)
plotBiomass(sim)

```


