---
title: "Getting steady-state in SO trait model"
output: html_notebook
---




Load libraries 
```{r}
remotes::install_github("sizespectrum/mizerExperimental")
```

```{r}
library(mizerExperimental)
library(mizer)
library(tidyverse)
# library(plotly)
```


Load group params

```{r}
groups_raw <- read.csv("group params/trait_groups_params_vCWC.csv")

groups <- groups_raw[,-7]

k_vb_string <- c(0.4,0.3608333,0.1825,0.15,1,0.2,0.5,0.06,0.2,0.2,0.2,0.2,0.2,0.2,0.2) # to use all k_vb values and instead of h

groups$k_vb <- k_vb_string

groups
```



Load interaction matrix
```{r}
theta <- readRDS("interaction matrix/trait_groups_interaction_matrix_vCWC.RDS")
theta
```

Adjust orca interaction with each other
Infanticide has been documented (https://doi.org/10.1038/s41598-018-22714-x), but as a starting point for the model, assuming they are not feeding on one another for maintenance of metabolism, growth and reproduction 

```{r}
# theta[13,13]
# 
# theta[13,13] <- 0
# 
# theta[13,13]
```


Setup params object
```{r}
params <- newMultispeciesParams(species_params = groups, 
                                interaction = theta, 
                                kappa = 2.63907e+13,
                                w_pp_cutoff = 10000,
                                n = 3/4, p = 3/4)
```

```{r}
groups$w_mat
params@species_params$w_mat
params@species_params$species
```
Readjust w_mat for the three mammal groups corrected to 25% of the original w_mat
'For the species leopard seals, large divers, minke whales, sperm whales the value for `w_mat` is not smaller than that of `w_max`'
If w_mat must be smaller than w_max, then use 90% as a compromise value
Correction: co-existence not achievable using 90%. Trying 75%

```{r}
# params_v1 <- params
# 
# params_v1@species_params$w_mat[params_v1@species_params$species == "leopard seals"]  <- 3.480000e+05 * 0.25
# params_v1@species_params$w_mat[params_v1@species_params$species == "large divers"]  <- 2.024000e+06 * 0.25
# params_v1@species_params$w_mat[params_v1@species_params$species == "minke whales"]  <- 6.000000e+06 * 0.25
# params_v1@species_params$w_mat[params_v1@species_params$species == "sperm whales"]  <- 3.650000e+07 * 0.25
# 
# params_v1 <- setParams(params_v1)
```





Check to make sure setParams didn't reagjust the w_mat values
```{r}
groups$w_mat
params_v1@species_params$w_mat
params_v1@species_params$species
```
The values look good, they are at the 90% corrected w_mat value rather than the 25% default

Check what the w_min values were adjusted to for small divers and leopard seals

```{r}
groups$w_min
params_v1@species_params$w_min
params_v1@species_params$species
```
It looks like 25% of w_mat has been used as the w_min correction. Again I'm going to redjust to apply a 90% correction.
Correction: 90% didn't satisfy the requirements of mizer, so I've had to redo with 85%

```{r}
# params_v2 <- params_v1
# 
# params_v2@species_params$w_min[params_v2@species_params$species == "small divers"]  <- params_v2@species_params$w_mat[params_v2@species_params$species == "small divers"] * 0.85
# params_v2@species_params$w_min[params_v2@species_params$species == "leopard seals"]  <- params_v2@species_params$w_mat[params_v2@species_params$species == "leopard seals"] * 0.85
# 
# params_v2 <- setParams(params_v2)
```

Check to make sure setParams didn't change the adjusted w_min values
```{r}
groups$w_min
params_v2@species_params$w_min
params_v2@species_params$w_mat
params_v2@species_params$species
```

Add in box feeding kernel for the baleen whales
max ppmr value are based on w_max baleen whale feeding on w_min euphausiids
min ppmr value based on w_min baleen whale feeding on w_max euphausiids

```{r}
params_v2@species_params$w_min[params_v2@species_params$species == "baleen whales"]
params_v2@species_params$w_max[params_v2@species_params$species == "baleen whales"]

params_v2@species_params$w_min[params_v2@species_params$species == "euphausiids"]
params_v2@species_params$w_max[params_v2@species_params$species == "euphausiids"]
```
```{r}
params_v2 <- params
```


```{r}
# 1.03e+08/6.31e-05 # w_max baleen whale divided by w_min euphausiids
# 2250000/3.162278 # w_min baleen whale divided by w_max euphausiids

# params_v2@species_params$ppmr_min[params_v2@species_params$species == "baleen whales"]  <- 711512.4
# params_v2@species_params$ppmr_max[params_v2@species_params$species == "baleen whales"] <- 1.63233e+12
# params_v2@species_params$ppmr_min[params_v2@species_params$species == "baleen whales"]  <- 711512.4
# params_v2@species_params$ppmr_max[params_v2@species_params$species == "baleen whales"] <- 1e+10
params_v2@species_params$ppmr_min[params_v2@species_params$species == "baleen whales"]  <- 1e5
params_v2@species_params$ppmr_max[params_v2@species_params$species == "baleen whales"] <-5e7
params_v2@species_params$pred_kernel_type[params_v2@species_params$species == "baleen whales"] <- "box"
```


```{r}
params_v2 <- setParams(params_v2)

params_v2@resource_params$w_pp_cutoff
```
Correct R_max for the kappa value

```{r}
params_sim <- params_v2

params_sim@species_params$R_max <- params_sim@resource_params$kappa*params_sim@species_params$w_max^-1.5

params_sim <- setParams(params_sim)

sim <- project(params_sim, effort=0, t_max = 500)

plot(sim)
```

```{r}
params_steady_v1 <- steady(params_sim)
```

Some species don't appear to be getting enough food, especially orca.
Increase gamma to see if this improves things

```{r}
params_sim_v2 <- params_sim
params_sim_v2@species_params$gamma[params_sim_v2@species_params$species == "orca"] <- params_sim@species_params$gamma[params_sim@species_params$species == "orca"] * 50
```


```{r}
sim_v2 <- project(params_sim_v2, effort=0, t_max = 1000)

plot(sim_v2)
```

```{r}
# params_tuned_v1 <- tuneGrowth(params_sim_v2)
params_tuned_v1 <- tuneParams(params_sim_v2)

```

```{r}
sim_tuned_v1 <- project(params_tuned_v1, effort=0, t_max = 500)

plot(sim_tuned_v1)
```


```{r}
params_tuned_v2 <- steady(params_tuned_v1)

params_tuned_v3 <- tuneParams(params_tuned_v2)
```







