---
title: "Calibrating a model within therMizer"
author: "Kieran Murphy"
date:  "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

If working with g m2 for biomass and yield, then I can compare as is with the plankton forcings, provided the depth itegrated plankton is divided by the area m2

## Load libraries

```{r}
# remotes::install_github("sizespectrum/mizerExperimental", ref = "tuneMR") # need to uninstall other mizerExperimental and resinstall this to use plotYieldCurve()

# remotes::install_github("sizespectrum/mizerExperimental")
# remotes::install_github("sizespectrum/therMizer")

# library(mizer)
library(therMizer)
library(tidyverse)
library(mizerExperimental)
# library(mizerMR)
# library(mizerHowTo)
# library(pbapply)
```

## Load therMizer params object

```{r}
# params <- readRDS("therMizer_params_v1.RDS")
```


## Check biomass and yield values currently

```{r}
# species_params(params)$biomass_observed
# species_params(params)$yield_observed
# gear_params(params)$yield_observed
```
Currently g/m2 (same as t/km2), so need to scale biomass using the model domain in m2

Scale this up to biomass values for the entire model domain (1474341km^2)

1474341 * 1e6 for m2 = 1.474341e+12


```{r}
# species_params(params)$biomass_observed <- species_params(params)$biomass_observed * 1.474341e+12
# species_params(params)$yield_observed <- species_params(params)$yield_observed * 1.474341e+12
# gear_params(params)$yield_observed <- gear_params(params)$yield_observed * 1.474341e+12
# 
# species_params(params)$biomass_observed
# species_params(params)$yield_observed
# gear_params(params)$yield_observed
```

Now biomass and yield are expressed as units (grams) per whole model domain

Still steady?

```{r}
# params <- setParams(params)
# params <- steady(params)
```
tune params to new biomasses, make sure steady and save update

```{r}
# params <- tuneParams(params)

# saveRDS(params, "therMizer_params_total_area.RDS")

params_test <- readRDS("therMizer_params_total_area.RDS")
```

```{r}
species_params(params_test)$biomass_observed
species_params(params_test)$yield_observed
gear_params(params_test)$yield_observed
```


```{r}
getReproductionLevel(params_test)
```


```{r}
# pdf(file = "plots/initial_yield_curves.pdf")
# par(mfrow = c(19,1))
plotYieldCurve(params_test, species = "mesozooplankton", F_max = 1)
plotYieldCurve(params_test, species = "other krill", F_max = 1)
plotYieldCurve(params_test, species = "other macrozooplankton", F_max = 1)
plotYieldCurve(params_test, species = "antarctic krill", F_max = 1)
plotYieldCurve(params_test, species = "salps", F_max = 1)
plotYieldCurve(params_test, species = "mesopelagic fishes", F_max = 1)
plotYieldCurve(params_test, species = "bathypelagic fishes", F_max = 1)
plotYieldCurve(params_test, species = "shelf and coastal fishes", F_max = 1)
plotYieldCurve(params_test, species = "flying birds", F_max = 1)
plotYieldCurve(params_test, species = "small divers", F_max = 1)
plotYieldCurve(params_test, species = "squids", F_max = 1)
plotYieldCurve(params_test, species = "toothfishes", F_max = 1)
plotYieldCurve(params_test, species = "leopard seals", F_max = 1)
plotYieldCurve(params_test, species = "medium divers", F_max = 1)
plotYieldCurve(params_test, species = "large divers", F_max = 1)
plotYieldCurve(params_test, species = "minke whales", F_max = 1)
plotYieldCurve(params_test, species = "orca", F_max = 1)
plotYieldCurve(params_test, species = "sperm whales", F_max = 1)
plotYieldCurve(params_test, species = "baleen whales", F_max = 1)
```

```{r}
new_repro_level <- c("mesozooplankton"= 0.4,  "other krill" = 0.8,  "other macrozooplankton" = 0.8, "antarctic krill" = 0.9, salps  = 0.95,  "mesopelagic fishes"  = 0.95, "bathypelagic fishes" = 0.95, "shelf and coastal fishes" = 0.95,"flying birds"= 0.8, "small divers" = 0.8, "squids" = 0.99, "toothfishes" = 0.9, "leopard seals" = 0.8018595, "medium divers" = 0.9, "large divers" = 0.9754472, "minke whales" = 0.9252620, "orca" = 0.9,  "sperm whales" = 0.9918454, "baleen whales" = 0.9)
```



```{r}
params_test_upgrade <- upgradeParams(params_test)

params_test_v2 <- setBevertonHolt(params_test_upgrade, reproduction_level = new_repro_level)

```


```{r}
# pdf(file = "plots/repro_level_v2_yield_curves.pdf")
# par(mfrow = c(19,1))
plotYieldCurve(params_test_v2, species = "mesozooplankton", F_max = 1)
plotYieldCurve(params_test_v2, species = "other krill", F_max = 1)
plotYieldCurve(params_test_v2, species = "other macrozooplankton", F_max = 1)
plotYieldCurve(params_test_v2, species = "antarctic krill", F_max = 1)
plotYieldCurve(params_test_v2, species = "salps", F_max = 1)
plotYieldCurve(params_test_v2, species = "mesopelagic fishes", F_max = 1)
plotYieldCurve(params_test_v2, species = "bathypelagic fishes", F_max = 1)
plotYieldCurve(params_test_v2, species = "shelf and coastal fishes", F_max = 1)
plotYieldCurve(params_test_v2, species = "flying birds", F_max = 1)
plotYieldCurve(params_test_v2, species = "small divers", F_max = 1)
plotYieldCurve(params_test_v2, species = "squids", F_max = 1)
plotYieldCurve(params_test_v2, species = "toothfishes", F_max = 1)
plotYieldCurve(params_test_v2, species = "leopard seals", F_max = 1)
plotYieldCurve(params_test_v2, species = "medium divers", F_max = 1)
plotYieldCurve(params_test_v2, species = "large divers", F_max = 1)
plotYieldCurve(params_test_v2, species = "minke whales", F_max = 1)
plotYieldCurve(params_test_v2, species = "orca", F_max = 1)
plotYieldCurve(params_test_v2, species = "sperm whales", F_max = 1)
plotYieldCurve(params_test_v2, species = "baleen whales", F_max = 1)
```

Visual checks of yield curves look reasonable after the tweaks to reproduction level values


```{r}
params <- tuneParams(params_test_v2)
```

Check erepro

```{r}
params@species_params$erepro

species_params(params)$erepro[species_params(params)$erepro >= 1] <- 0.999

params@species_params$erepro
```

```{r}
species_params(params)$biomass_observed
species_params(params)$yield_observed
gear_params(params)$yield_observed
```

```{r}
# params <- tuneParams(params)
# 
# species_params(params)$erepro
```


```{r}
# saveRDS(params, "therMizer_params_total_area_v2.RDS")
```


```{r}
# effort <- read.csv("effort_array.csv")
effort <- read.csv("FishMIP_fishing_data/effort_array.csv")

years <- effort[,1]

effort <- effort[,-1]
rownames(effort) <- years


# colnames(effort) <- rownames(gear_params(params))
colnames(effort) <- gear_params(params)$gear

head(effort)
glimpse(effort)
dim(effort)
```


Set up time steps
```{r}
time_steps <- 1961:2010
```

### temperature data

#input temperature data 
```{r}
tos <- readRDS("FishMIP_Temperature_Forcing/tos_annual.rds")
t500m <- readRDS("FishMIP_Temperature_Forcing/t500m_annual.rds")
t1000m <- readRDS("FishMIP_Temperature_Forcing/t1000m_annual.rds")
t1500m <- readRDS("FishMIP_Temperature_Forcing/t1500m_annual.rds")
tob <- readRDS("FishMIP_Temperature_Forcing/tob_annual.rds")
ocean_temp <- abind::abind(tos,t500m,t1000m, t1500m, tob, along = 2)

realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")

colnames(ocean_temp) <- realm_names
rownames(ocean_temp) <- time_steps
```


### input npp data

```{r}
isimip_plankton <- read.table("GFDL_resource_spectra_annual.dat") # log10 abundance

# To rescale the plankton abundance to m2, will need to raise N^10, divide area and then log10 again.

isimip_plankton <- as(isimip_plankton, "matrix")

rownames(isimip_plankton) <- time_steps
sizes <- params_test@w_full
# sizes <- params@w_full
colnames(isimip_plankton) <- sizes # signif(sizes, 6)

dim(isimip_plankton)
```

```{r}
n_pp_array <- array(NA, dim = c(length(time_steps), length(sizes)), dimnames = list(time = time_steps, w = sizes)) # setup an empty array for the corrected plankton

#
n_pp_array[1,] <- isimip_plankton[1,] #subtracting anomaly and adding back the mean
#n_pp_array[1,] <- (isimip_plankton[1,] - colMeans(isimip_plankton[,])) + log10(params@initial_n_pp) #subtracting anomaly and adding back the mean
 for (t in seq(1,length(time_steps) - 1,1)) {
   n_pp_array[t + 1,] <- isimip_plankton[t + 1,]
 }

# Introducing zeros which convert to -Inf

head(n_pp_array[1,]) #make it comparible
head(n_pp_array[50,])
# head(log10(initialNResource(params)))
head(log10(initialNResource(params)))
```

Above we changed the array that will be used to force the plankton dynamics through time.
We also need to adjust the initial n_pp
```{r}
params_updated_n_pp <- params

years_to_use<-as.character(2000:2010)

# years_to_use_v2<-as.character(2000-01-01:2010-12-01)

# head(initialNResource(new_params))


# initialNResource(params)[which(w_full(params) < 100)]<-10^colMeans(n_pp_array[years_to_use,])[which(w_full(params) < resource_params(params)$w_pp_cutoff)]
# initialNResource(params_updated_n_pp)[which(w_full(params_updated_n_pp) < 100)]<- 10^colMeans(n_pp_array[years_to_use,])[which(w_full(params_updated_n_pp) < 100)]

initialNResource(params_updated_n_pp)[which(w_full(params_updated_n_pp) < 100)]<- 10^colMeans(n_pp_array[years_to_use,])[which(w_full(params_updated_n_pp) < 100)]/params_updated_n_pp@dw_full # normalise the plankton spectrum to match the therMizer plankton_forcing()
#need to cut resource at 100g
initialNResource(params_updated_n_pp)[which(w_full(params_updated_n_pp) >=resource_params(params_updated_n_pp)$w_pp_cutoff)]<-0.0 # sets all above 100g to 0


head(initialNResource(params))
head(initialNResource(params_updated_n_pp))
```

```{r}
params_updated_n_pp <- steady(params_updated_n_pp)

plotSpectra(params_updated_n_pp)
```

```{r}
params_updated_n_pp_tuned <- tuneParams(params_updated_n_pp)
```


#vertical_migration array set up----
```{r}
realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")
colnames(ocean_temp) <- realm_names
species_names <- as.character(params@species_params$species)
sizes <- params@w

vertical_migration_array <- array(0, dim = (c(length(realm_names),
                                              length(species_names), length(sizes))),
                                  dimnames = list(realm = realm_names, sp = species_names,
                                                  w = signif(sizes, 3)))
```


```{r}
zone01 <- which(realm_names =="tos")
zone02 <- which(realm_names == "t500m")
zone03 <- which(realm_names =="t1000m")
zone04 <- which(realm_names == "t1500m")
zone05 <- which(realm_names == "tob")
```



```{r}
#mesozooplankton
MesZoo <- which(species_names=="mesozooplankton")
vertical_migration_array[zone01, MesZoo, ] <- 0.25
vertical_migration_array[zone02, MesZoo, ] <- 0.25
vertical_migration_array[zone03, MesZoo, ] <- 0.25
vertical_migration_array[zone04, MesZoo, ] <- 0.25
vertical_migration_array[zone05, MesZoo, ] <- 0
#other krill
Oth_krill <- which(species_names=="other krill")
vertical_migration_array[zone01, Oth_krill, ] <- 0.25
vertical_migration_array[zone02, Oth_krill, ] <- 0.25
vertical_migration_array[zone03, Oth_krill, ] <- 0.25
vertical_migration_array[zone04, Oth_krill, ] <- 0.25
vertical_migration_array[zone05, Oth_krill, ] <- 0
#other macrozooplankton
Oth_MacroZoo <- which(species_names=="other macrozooplankton")
vertical_migration_array[zone01, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone02, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone03, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone04, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone05, Oth_MacroZoo, ] <- 0
#antarctic krill
Ant_krill <- which(species_names=="antarctic krill")
vertical_migration_array[zone01, Ant_krill, ] <- 0.25
vertical_migration_array[zone02, Ant_krill, ] <- 0.25
vertical_migration_array[zone03, Ant_krill, ] <- 0.25
vertical_migration_array[zone04, Ant_krill, ] <- 0.25
vertical_migration_array[zone05, Ant_krill, ] <- 0
#salps
salps <- which(species_names=="salps")
vertical_migration_array[zone01, salps, ] <- 0.25
vertical_migration_array[zone02, salps, ] <- 0.25
vertical_migration_array[zone03, salps, ] <- 0.25
vertical_migration_array[zone04, salps, ] <- 0.25
vertical_migration_array[zone05, salps, ] <- 0
#mesopelagic fishes
MesoFish <- which(species_names=="mesopelagic fishes")
vertical_migration_array[zone01, MesoFish, ] <- 0.2
vertical_migration_array[zone02, MesoFish, ] <- 0.4
vertical_migration_array[zone03, MesoFish, ] <- 0.4
vertical_migration_array[zone04, MesoFish, ] <- 0
vertical_migration_array[zone05, MesoFish, ] <- 0
#bathypelagic fishes
BathyFish <- which(species_names=="bathypelagic fishes")
vertical_migration_array[zone01, BathyFish, ] <- 0
vertical_migration_array[zone02, BathyFish, ] <- 0.15
vertical_migration_array[zone03, BathyFish, ] <- 0.15
vertical_migration_array[zone04, BathyFish, ] <- 0.35
vertical_migration_array[zone05, BathyFish, ] <- 0.35
#shelf and coastal fishes
SCFish <- which(species_names=="shelf and coastal fishes")
vertical_migration_array[zone01, SCFish, ] <- 0.2
vertical_migration_array[zone02, SCFish, ] <- 0.2
vertical_migration_array[zone03, SCFish, ] <- 0.2
vertical_migration_array[zone04, SCFish, ] <- 0.2
vertical_migration_array[zone05, SCFish, ] <- 0.2
#flying birds
Birds <- which(species_names=="flying birds")
vertical_migration_array[zone01, Birds, ] <- 1
vertical_migration_array[zone02, Birds, ] <- 0
vertical_migration_array[zone03, Birds, ] <- 0
vertical_migration_array[zone04, Birds, ] <- 0
vertical_migration_array[zone05, Birds, ] <- 0
#small divers
Sm_divers <- which(species_names=="small divers")
vertical_migration_array[zone01, Sm_divers, ] <- 1
vertical_migration_array[zone02, Sm_divers, ] <- 0
vertical_migration_array[zone03, Sm_divers, ] <- 0
vertical_migration_array[zone04, Sm_divers, ] <- 0
vertical_migration_array[zone05, Sm_divers, ] <- 0
#squids
squids <- which(species_names=="squids")
vertical_migration_array[zone01, squids, ] <- 0.2
vertical_migration_array[zone02, squids, ] <- 0.2
vertical_migration_array[zone03, squids, ] <- 0.2
vertical_migration_array[zone04, squids, ] <- 0.2
vertical_migration_array[zone05, squids, ] <- 0.2
#toothfishes
toothfishes <- which(species_names=="toothfishes")
vertical_migration_array[zone01, toothfishes, ] <- 0.2
vertical_migration_array[zone02, toothfishes, ] <- 0.2
vertical_migration_array[zone03, toothfishes, ] <- 0.2
vertical_migration_array[zone04, toothfishes, ] <- 0.2
vertical_migration_array[zone05, toothfishes, ] <- 0.2
#leopard seals
seals <- which(species_names=="leopard seals")
vertical_migration_array[zone01, seals, ] <- 0.5
vertical_migration_array[zone02, seals, ] <- 0.5
vertical_migration_array[zone03, seals, ] <- 0
vertical_migration_array[zone04, seals, ] <- 0
vertical_migration_array[zone05, seals, ] <- 0
#medium divers
Md_divers <- which(species_names=="medium divers")
vertical_migration_array[zone01, Md_divers, ] <- 0.5
vertical_migration_array[zone02, Md_divers, ] <- 0.5
vertical_migration_array[zone03, Md_divers, ] <- 0
vertical_migration_array[zone04, Md_divers, ] <- 0
vertical_migration_array[zone05, Md_divers, ] <- 0
#large diver
Lrg_divers <- which(species_names=="large divers")
vertical_migration_array[zone01, Lrg_divers, ] <- 0.5
vertical_migration_array[zone02, Lrg_divers, ] <- 0.5
vertical_migration_array[zone03, Lrg_divers, ] <- 0
vertical_migration_array[zone04, Lrg_divers, ] <- 0
vertical_migration_array[zone05, Lrg_divers, ] <- 0
#minke whales
minke <- which(species_names=="minke whales")
vertical_migration_array[zone01, minke, ] <- 0.4
vertical_migration_array[zone02, minke, ] <- 0.4
vertical_migration_array[zone03, minke, ] <- 0.1
vertical_migration_array[zone04, minke, ] <- 0.1
vertical_migration_array[zone05, minke, ] <- 0
#orca
orca <- which(species_names=="orca")
vertical_migration_array[zone01, orca, ] <- 0.4
vertical_migration_array[zone02, orca, ] <- 0.4
vertical_migration_array[zone03, orca, ] <- 0.1
vertical_migration_array[zone04, orca, ] <- 0.1
vertical_migration_array[zone05, orca, ] <- 0
#sperm whales
sperm <- which(species_names=="sperm whales")
vertical_migration_array[zone01, sperm, ] <- 0.3
vertical_migration_array[zone02, sperm, ] <- 0.3
vertical_migration_array[zone03, sperm, ] <- 0.2
vertical_migration_array[zone04, sperm, ] <- 0.2
vertical_migration_array[zone05, sperm, ] <- 0
#baleen whales
Baleen <- which(species_names=="baleen whales")
vertical_migration_array[zone01, Baleen, ] <- 0.4
vertical_migration_array[zone02, Baleen, ] <- 0.4
vertical_migration_array[zone03, Baleen, ] <- 0.1
vertical_migration_array[zone04, Baleen, ] <- 0.1
vertical_migration_array[zone05, Baleen, ] <- 0
```



```{r}
paramsT <- upgradeTherParams(params_updated_n_pp,
                             ocean_temp_array = ocean_temp,
                             n_pp_array = n_pp_array, 
                             vertical_migration_array = vertical_migration_array,
                             aerobic_effect = TRUE, metabolism_effect = TRUE)
```


```{r}
params <- paramsT

paramsT@species_params$biomass_observed
```


```{r}
params <- tuneParams(params)
```


```{r}
params <- params %>% 
  steady() %>% matchBiomasses() %>% steady()
```


```{r}
plotTherPerformance(paramsT)
plotTherScalar(paramsT)
plotSpectra(paramsT, total = T, power = 2)
plotGrowthCurves(paramsT, species_panel = T)
```


```{r}
temp_min <- c(-2, # mesozooplankton
                      -2, # other krill
                     -2, # other macrozooplankton
                     -2, # antarctic krill
                                       -2,   # salps
                                     -2, # mesopelagic fishes
                                     -2, # bathypelagic fishes
                                     -2, # "shelf and coastal fishes" 
                                     -1.5, # "flying birds"
                                     -1.5, # "small divers"             
                                     -1.5, # "squids"                  
                                     -1.5, # "toothfishes"              
                                     -1.9, # "leopard seals"           
                                     -1.8, # "medium divers"            
                                     -1.5, # "large divers"             
                                     -1.5, # "minke whales"             
                                     -1.5, # "orca"                     
                                     -1.5, # "sperm whales"             
                                     -1.5 # "baleen whales"
                                      ) 

temp_max <- c(2, # mesozooplankton
                      2, # other krill
                    2,  # other macrozooplankton
                  2, # antarctic krill
                    8.57, # salps
                                     5, # mesopelagic fishes
                                     5, # bathypelagic fishes
                                     5, # "shelf and coastal fishes" 
                                     18.2, # "flying birds"
                                     1, # "small divers"             
                                     1.1, # "squids"                  
                                     8.8, # "toothfishes"              
                                     1.4, # "leopard seals"           
                                     2, # "medium divers"            
                                     1.6, # "large divers"             
                                     7, # "minke whales"             
                                     13.1, # "orca"                     
                                     3.8, # "sperm whales"             
                                     10.2 # "baleen whales"
                                      ) 
```



```{r}
species_params(paramsT)$temp_min <- temp_min
species_params(paramsT)$temp_max <- temp_max


paramsT <- upgradeTherParams(paramsT,
                             ocean_temp_array = ocean_temp,
                             n_pp_array = n_pp_array, 
                             vertical_migration_array = vertical_migration_array,
                             aerobic_effect = TRUE, metabolism_effect = TRUE)
```


```{r}
paramsT <- paramsT %>% 
  steady() %>% matchBiomasses() %>% steady() %>% matchBiomasses() %>% 
  steady() %>% matchBiomasses() %>% steady() %>% matchBiomasses() %>% 
  steady() %>% matchBiomasses() %>% steady() %>% matchBiomasses() %>%
  steady() %>% matchBiomasses() %>% steady() %>% matchBiomasses() %>% 
  steady() %>% matchBiomasses() %>% steady() %>% matchBiomasses() %>% 
  steady() %>% matchBiomasses() %>% steady() %>% matchBiomasses() %>%
  steady() %>% matchBiomasses() %>% steady()


paramsT@species_params$erepro
getReproductionLevel(paramsT)
```

```{r}
# saveRDS(paramsT, "latest_therMizer_params.RDS")
```


```{r}
params <- readRDS("latest_therMizer_params.RDS")

species_params(params)
summary(species_params(params))
```


```{r}
getReproductionLevel(params)
```


```{r}
params <- tuneParams(params)

# saveRDS(params, "therMizer_params_v03.RDS")
```

therMizer_params_v03.RDS
Manually tuned growth in tuneParams() after updating to therMizer params object


Check reproduction parameters after the growth tuning

```{r}
params@species_params$erepro
params@species_params$R_max
getReproductionLevel(params)
```

```{r}
# pdf(file = "plots/therMizer_initial_yield_curves.pdf")
# par(mfrow = c(19,1))
# plotYieldVsF(params, species = "mesozooplankton", F_max = 1)
# plotYieldVsF(params, species = "other krill", F_max = 1)
# plotYieldVsF(params, species = "other macrozooplankton", F_max = 1)
# plotYieldVsF(params, species = "antarctic krill", F_max = 1)
# plotYieldVsF(params, species = "salps", F_max = 1)
# plotYieldVsF(params, species = "mesopelagic fishes", F_max = 1)
# plotYieldVsF(params, species = "bathypelagic fishes", F_max = 1)
# plotYieldVsF(params, species = "shelf and coastal fishes", F_max = 1)
# plotYieldVsF(params, species = "flying birds", F_max = 1)
# plotYieldVsF(params, species = "small divers", F_max = 1)
# plotYieldVsF(params, species = "squids", F_max = 1)
# plotYieldVsF(params, species = "toothfishes", F_max = 1)
# plotYieldVsF(params, species = "leopard seals", F_max = 1)
# plotYieldVsF(params, species = "medium divers", F_max = 1)
# plotYieldVsF(params, species = "large divers", F_max = 1)
# plotYieldVsF(params, species = "minke whales", F_max = 1)
# plotYieldVsF(params, species = "orca", F_max = 1)
# plotYieldVsF(params, species = "sperm whales", F_max = 1)
# plotYieldVsF(params, species = "baleen whales", F_max = 1)
```


```{r}
getReproductionLevel(params)
```



```{r}
new_repro_level <- c("mesozooplankton"= 0.4,  "other krill" = 0.85,  "other macrozooplankton" = 0.85, "antarctic krill" = 0.9, salps  = 0.99,  "mesopelagic fishes"  = 0.99, "bathypelagic fishes" = 0.99, "shelf and coastal fishes" = 0.99,"flying birds"= 0.8213785, "small divers" = 0.7619014, "squids" = 0.999, "toothfishes" = 0.9816799, "leopard seals" = 0.8608031, "medium divers" =  0.9633993, "large divers" = 0.9732486, "minke whales" = 0.8274689, "orca" = 0.99999999,  "sperm whales" = 0.9994376, "baleen whales" = 0.9982165)
```



```{r}
params_v2 <- setBevertonHolt(params, reproduction_level = new_repro_level)
```



```{r}
# pdf(file = "plots/therMizer_repro_level_v2_yield_curves.pdf")
# par(mfrow = c(19,1))
# plotYieldVsF(params_v2, species = "mesozooplankton", F_max = 1)
# plotYieldVsF(params_v2, species = "other krill", F_max = 1)
# plotYieldVsF(params_v2, species = "other macrozooplankton", F_max = 1)
# plotYieldVsF(params_v2, species = "antarctic krill", F_max = 1)
# plotYieldVsF(params_v2, species = "salps", F_max = 1)
# plotYieldVsF(params_v2, species = "mesopelagic fishes", F_max = 1)
# plotYieldVsF(params_v2, species = "bathypelagic fishes", F_max = 1)
# plotYieldVsF(params_v2, species = "shelf and coastal fishes", F_max = 1)
# plotYieldVsF(params_v2, species = "flying birds", F_max = 1)
# plotYieldVsF(params_v2, species = "small divers", F_max = 1)
# plotYieldVsF(params_v2, species = "squids", F_max = 1)
# plotYieldVsF(params_v2, species = "toothfishes", F_max = 1)
# plotYieldVsF(params_v2, species = "leopard seals", F_max = 1)
# plotYieldVsF(params_v2, species = "medium divers", F_max = 1)
# plotYieldVsF(params_v2, species = "large divers", F_max = 1)
# plotYieldVsF(params_v2, species = "minke whales", F_max = 1)
# plotYieldVsF(params_v2, species = "orca", F_max = 1)
# plotYieldVsF(params_v2, species = "sperm whales", F_max = 1)
# plotYieldVsF(params_v2, species = "baleen whales", F_max = 1)
```



```{r}
# saveRDS(params_v2, "therMizer_params_v04.RDS")

```

