---
title: "Calibrating a model within therMizer"
author: "Kieran Murphy"
date:  "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

## Load libraries

```{r}
# library(mizer)
library(therMizer)
library(tidyverse)
library(mizerExperimental)
# library(mizerHowTo)
# library(pbapply)
```

## Load therMizer params object

```{r}
params <- readRDS("therMizer_params_v1.RDS")
```


## Check biomass values currently

```{r}
species_params(params)$biomass_observed
```
Current values are t/km2 (same as g/m2, but need to be specific for next step)

Effort and catch are based on t/km2, so need to scale biomass using the model domain in km2

Scale this up to biomass values for the entire model domain (1474341 km^2)


```{r}
species_params(params)$biomass_observed <- species_params(params)$biomass_observed * 1474341

species_params(params)$biomass_observed
```
Still steady?

```{r}
params <- setParams(params)
params <- steady(params)
```
tune params to new biomasses, make sure steady and save update

```{r}
params <- tuneParams(params)

# saveRDS(params, "therMizer_params_total_area.RDS")
```

Need to scale catch to the whole area too



```{r}
species_params(params)$yield_observed
gear_params(params)$yield_observed
```

```{r}
species_params(params)$yield_observed <- species_params(params)$yield_observed * 1474341
gear_params(params)$yield_observed <- gear_params(params)$yield_observed * 1474341

species_params(params)$yield_observed
gear_params(params)$yield_observed
```

```{r}
# saveRDS(params, "therMizer_params_total_area.RDS")
```


```{r}
effort <- read.csv("effort_array.csv")

years <- effort[,1]

effort <- effort[,-1]
rownames(effort) <- years


# colnames(effort) <- rownames(gear_params(params))
colnames(effort) <- gear_params(params)$gear

head(effort)
glimpse(effort)
dim(effort)
```


Set up time steps
```{r}
time_steps <- 1961:2010
```

### temperature data

#input temperature data 
```{r}
tos <- readRDS("tos_annual.rds")
t500m <- readRDS("t500m_annual.rds")
t1000m <- readRDS("t1000m_annual.rds")
t1500m <- readRDS("t1500m_annual.rds")
tob <- readRDS("tob_annual.rds")
ocean_temp <- abind::abind(tos,t500m,t1000m, t1500m, tob, along = 2)

realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")

colnames(ocean_temp) <- realm_names
rownames(ocean_temp) <- time_steps
```


### input npp data

```{r}
isimip_plankton <- read.table("GFDL_resource_spectra_annual.dat")

isimip_plankton <- as(isimip_plankton, "matrix")

rownames(isimip_plankton) <- time_steps
sizes <- params@w_full
colnames(isimip_plankton) <- sizes # signif(sizes, 6)

dim(isimip_plankton)
```

```{r}
n_pp_array <- array(NA, dim = c(length(time_steps), length(sizes)), dimnames = list(time = time_steps, w = sizes)) # setup an empty array for the corrected plankton

#
n_pp_array[1,] <- (isimip_plankton[1,] - colMeans(isimip_plankton[,])) + log10(params@initial_n_pp) #subtracting anomaly and adding back the mean
 for (t in seq(1,length(time_steps) - 1,1)) {
   n_pp_array[t + 1,] <- (isimip_plankton[t + 1,]  - colMeans(isimip_plankton[,])) + log10(params@initial_n_pp)
 }

# Introducing zeros which convert to -Inf

head(n_pp_array[1,]) #make it comparible
head(n_pp_array[50,])
head(log10(initialNResource(params)))
```

Above we changed the array that will be used to force the plankton dynamics through time.
We also need to adjust the initial n_pp
```{r}
params_updated_n_pp <- params

years_to_use<-as.character(2000:2010)

# years_to_use_v2<-as.character(2000-01-01:2010-12-01)

# head(initialNResource(new_params))


# initialNResource(params)[which(w_full(params) < 100)]<-10^colMeans(n_pp_array[years_to_use,])[which(w_full(params) < resource_params(params)$w_pp_cutoff)]
initialNResource(params_updated_n_pp)[which(w_full(params_updated_n_pp) < 100)]<- 10^colMeans(n_pp_array[years_to_use,])[which(w_full(params_updated_n_pp) < 100)]
#need to cut resource at 100g
initialNResource(params_updated_n_pp)[which(w_full(params_updated_n_pp) >=resource_params(params_updated_n_pp)$w_pp_cutoff)]<-0.0 # sets all above 100g to 0


head(initialNResource(params))
head(initialNResource(params_updated_n_pp))
```

```{r}
params_updated_n_pp <- steady(params_updated_n_pp)

plotSpectra(params_updated_n_pp)
```

#vertical_migration array set up----
```{r}
realm_names <- c("tos","t500m","t1000m", "t1500m", "tob")
colnames(ocean_temp) <- realm_names
species_names <- as.character(params@species_params$species)
sizes <- params@w

vertical_migration_array <- array(0, dim = (c(length(realm_names),
                                              length(species_names), length(sizes))),
                                  dimnames = list(realm = realm_names, sp = species_names,
                                                  w = signif(sizes, 3)))
```


```{r}
zone01 <- which(realm_names =="tos")
zone02 <- which(realm_names == "t500m")
zone03 <- which(realm_names =="t1000m")
zone04 <- which(realm_names == "t1500m")
zone05 <- which(realm_names == "tob")
```



```{r}
#mesozooplankton
MesZoo <- which(species_names=="mesozooplankton")
vertical_migration_array[zone01, MesZoo, ] <- 0.25
vertical_migration_array[zone02, MesZoo, ] <- 0.25
vertical_migration_array[zone03, MesZoo, ] <- 0.25
vertical_migration_array[zone04, MesZoo, ] <- 0.25
vertical_migration_array[zone05, MesZoo, ] <- 0
#other krill
Oth_krill <- which(species_names=="other krill")
vertical_migration_array[zone01, Oth_krill, ] <- 0.25
vertical_migration_array[zone02, Oth_krill, ] <- 0.25
vertical_migration_array[zone03, Oth_krill, ] <- 0.25
vertical_migration_array[zone04, Oth_krill, ] <- 0.25
vertical_migration_array[zone05, Oth_krill, ] <- 0
#other macrozooplankton
Oth_MacroZoo <- which(species_names=="other macrozooplankton")
vertical_migration_array[zone01, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone02, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone03, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone04, Oth_MacroZoo, ] <- 0.25
vertical_migration_array[zone05, Oth_MacroZoo, ] <- 0
#antarctic krill
Ant_krill <- which(species_names=="antarctic krill")
vertical_migration_array[zone01, Ant_krill, ] <- 0.25
vertical_migration_array[zone02, Ant_krill, ] <- 0.25
vertical_migration_array[zone03, Ant_krill, ] <- 0.25
vertical_migration_array[zone04, Ant_krill, ] <- 0.25
vertical_migration_array[zone05, Ant_krill, ] <- 0
#salps
salps <- which(species_names=="salps")
vertical_migration_array[zone01, salps, ] <- 0.25
vertical_migration_array[zone02, salps, ] <- 0.25
vertical_migration_array[zone03, salps, ] <- 0.25
vertical_migration_array[zone04, salps, ] <- 0.25
vertical_migration_array[zone05, salps, ] <- 0
#mesopelagic fishes
MesoFish <- which(species_names=="mesopelagic fishes")
vertical_migration_array[zone01, MesoFish, ] <- 0.2
vertical_migration_array[zone02, MesoFish, ] <- 0.4
vertical_migration_array[zone03, MesoFish, ] <- 0.4
vertical_migration_array[zone04, MesoFish, ] <- 0
vertical_migration_array[zone05, MesoFish, ] <- 0
#bathypelagic fishes
BathyFish <- which(species_names=="bathypelagic fishes")
vertical_migration_array[zone01, BathyFish, ] <- 0
vertical_migration_array[zone02, BathyFish, ] <- 0.15
vertical_migration_array[zone03, BathyFish, ] <- 0.15
vertical_migration_array[zone04, BathyFish, ] <- 0.35
vertical_migration_array[zone05, BathyFish, ] <- 0.35
#shelf and coastal fishes
SCFish <- which(species_names=="shelf and coastal fishes")
vertical_migration_array[zone01, SCFish, ] <- 0.2
vertical_migration_array[zone02, SCFish, ] <- 0.2
vertical_migration_array[zone03, SCFish, ] <- 0.2
vertical_migration_array[zone04, SCFish, ] <- 0.2
vertical_migration_array[zone05, SCFish, ] <- 0.2
#flying birds
Birds <- which(species_names=="flying birds")
vertical_migration_array[zone01, Birds, ] <- 1
vertical_migration_array[zone02, Birds, ] <- 0
vertical_migration_array[zone03, Birds, ] <- 0
vertical_migration_array[zone04, Birds, ] <- 0
vertical_migration_array[zone05, Birds, ] <- 0
#small divers
Sm_divers <- which(species_names=="small divers")
vertical_migration_array[zone01, Sm_divers, ] <- 1
vertical_migration_array[zone02, Sm_divers, ] <- 0
vertical_migration_array[zone03, Sm_divers, ] <- 0
vertical_migration_array[zone04, Sm_divers, ] <- 0
vertical_migration_array[zone05, Sm_divers, ] <- 0
#squids
squids <- which(species_names=="squids")
vertical_migration_array[zone01, squids, ] <- 0.2
vertical_migration_array[zone02, squids, ] <- 0.2
vertical_migration_array[zone03, squids, ] <- 0.2
vertical_migration_array[zone04, squids, ] <- 0.2
vertical_migration_array[zone05, squids, ] <- 0.2
#toothfishes
toothfishes <- which(species_names=="toothfishes")
vertical_migration_array[zone01, toothfishes, ] <- 0.2
vertical_migration_array[zone02, toothfishes, ] <- 0.2
vertical_migration_array[zone03, toothfishes, ] <- 0.2
vertical_migration_array[zone04, toothfishes, ] <- 0.2
vertical_migration_array[zone05, toothfishes, ] <- 0.2
#leopard seals
seals <- which(species_names=="leopard seals")
vertical_migration_array[zone01, seals, ] <- 0.5
vertical_migration_array[zone02, seals, ] <- 0.5
vertical_migration_array[zone03, seals, ] <- 0
vertical_migration_array[zone04, seals, ] <- 0
vertical_migration_array[zone05, seals, ] <- 0
#medium divers
Md_divers <- which(species_names=="medium divers")
vertical_migration_array[zone01, Md_divers, ] <- 0.5
vertical_migration_array[zone02, Md_divers, ] <- 0.5
vertical_migration_array[zone03, Md_divers, ] <- 0
vertical_migration_array[zone04, Md_divers, ] <- 0
vertical_migration_array[zone05, Md_divers, ] <- 0
#large diver
Lrg_divers <- which(species_names=="large divers")
vertical_migration_array[zone01, Lrg_divers, ] <- 0.5
vertical_migration_array[zone02, Lrg_divers, ] <- 0.5
vertical_migration_array[zone03, Lrg_divers, ] <- 0
vertical_migration_array[zone04, Lrg_divers, ] <- 0
vertical_migration_array[zone05, Lrg_divers, ] <- 0
#minke whales
minke <- which(species_names=="minke whales")
vertical_migration_array[zone01, minke, ] <- 0.4
vertical_migration_array[zone02, minke, ] <- 0.4
vertical_migration_array[zone03, minke, ] <- 0.1
vertical_migration_array[zone04, minke, ] <- 0.1
vertical_migration_array[zone05, minke, ] <- 0
#orca
orca <- which(species_names=="orca")
vertical_migration_array[zone01, orca, ] <- 0.4
vertical_migration_array[zone02, orca, ] <- 0.4
vertical_migration_array[zone03, orca, ] <- 0.1
vertical_migration_array[zone04, orca, ] <- 0.1
vertical_migration_array[zone05, orca, ] <- 0
#sperm whales
sperm <- which(species_names=="sperm whales")
vertical_migration_array[zone01, sperm, ] <- 0.3
vertical_migration_array[zone02, sperm, ] <- 0.3
vertical_migration_array[zone03, sperm, ] <- 0.2
vertical_migration_array[zone04, sperm, ] <- 0.2
vertical_migration_array[zone05, sperm, ] <- 0
#baleen whales
Baleen <- which(species_names=="baleen whales")
vertical_migration_array[zone01, Baleen, ] <- 0.4
vertical_migration_array[zone02, Baleen, ] <- 0.4
vertical_migration_array[zone03, Baleen, ] <- 0.1
vertical_migration_array[zone04, Baleen, ] <- 0.1
vertical_migration_array[zone05, Baleen, ] <- 0
```



```{r}
paramsT <- upgradeTherParams(params_updated_n_pp,
                             ocean_temp_array = ocean_temp,
                             n_pp_array = n_pp_array, 
                             vertical_migration_array = vertical_migration_array,
                             aerobic_effect = TRUE, metabolism_effect = TRUE)
```



```{r}
plotTherPerformance(paramsT)
plotTherScalar(paramsT)
plotSpectra(paramsT, total = T, power = 2)
plotGrowthCurves(paramsT, species_panel = T)
```



```{r}
sim <- project(paramsT, t_max = 50)
# plotBiomass(paramsT)
plotlyBiomass(sim)
plotSpectra(sim)
```

```{r}
head(sim@n_pp[1,])
head(sim@n_pp[2,])
head(sim@n_pp[50,])
```


```{r}
paramsT <- tuneParams(paramsT)
```






