---
title: "therMizer"
output: html_notebook
---

The remotes package is needed to install packages hosted on GitHub.

```{r}
install.packages("remotes")

remotes::install_github("sizespectrum/therMizer@dev")
```

Load the newly installed package with

```{r}
library(therMizer)
```



```{r}
params <- readRDS("params_16_March_2023.rds")

sp_params <- params@species_params

sp_params
```

```{r}
plotlySpectra(params)
```


Adding min and max temp for each group

Using sealifebase for all non-fish species/groups. In most cases, there are no listed Temperature values in the 'Environment' section, like there is for more data-rich species, e.g., cod

Instead, there are values at the bottom of the sealifebase pages in 'Estimates of some properties based on models' with 'Preferred temperature' estimates from the AquaMaps model-based approach (https://www.gbif.org/tool/81356/aquamaps-predicted-range-maps-for-aquatic-species)  

[1] "euphausiids":
Antarctic krill - (Preferred temperature (Ref. 115969): -1.5 - 1.6, mean 0.2 (based on 535 cells))
Crystal krill - (Preferred temperature (Ref. 115969): -1.8 - 1.1, mean -1.6 (based on 3588 cells)) 
Thysanoessa macrura - (Preferred temperature (Ref. 115969): -1.2 - 2.8, mean 0.2 (based on 24 cells). Not included due to small number of cells and low contribution to biomass in this model domain
[2]"mesopelagic fishes" - Electrona antarctica(nothing on fishbase)    
[3]"bathypelagic fishes"      
[4]"shelf and coastal fishes":
Marbled rockcod (-1 - 5)
[5]"flying birds":
Southern giant petrel (Preferred temperature (Ref. 115969): 2.1 - 15.6, mean 10.1 (based on 533 cells));
Arctic tern (Sterna paradisaea) (Preferred temperature (Ref. 115969): 0.4 - 13.3, mean 6.8 (based on 2468 cells))
Thalassarche chrysostoma (Grey-headed petrel)  (Preferred temperature (Ref. 115969): 1.3 - 9.1, mean 5.2 (based on 263 cells))
South polar skua (Preferred temperature (Ref. 115969): 0.2 - 18.2, mean 6.4 (based on 1318 cells))
Diomedea exulans (Wandering albatross), (Preferred temperature (Ref. 115969): 0.5 - 12, mean 4.6 (based on 1528 cells)_
[6]"small divers" - no info for adelie, gentoo or crested penguins on sealifebase. Guess based on medium divers       
[7]"squids" - based on Psychroteuthis glacialis (sealifebase)                   
[8]"toothfishes" - no data on fishbase            
[9]"leopard seals" - Preferred temperature (Ref. 115969): -1.9 - 1.4, mean -0.5 (based on 2109 cells).          
[10] "medium divers":
Crabeater Seals - (Preferred temperature (Ref. 115969): -0.4 - 1, mean 0.1 (based on 9397 cells))
Ross seals (Preferred temperature (Ref. 115969): -1.8 - 1.1, mean -1.5 (based on 2110 cells));
Weddell seals (Preferred temperature (Ref. 115969): -1.8 - 1, mean -1.6 (based on 6026 cells));
Hourglass dolphin (Preferred temperature (Ref. 115969): 0.1 - 2, mean 0.9 (based on 11222 cells)) # Not included in the biomass
King penguin (no data on sealifebase)
Emperor penguin (no data on sealifebase)
[11]"large divers" - Southern elephant seals only in this group now that Sperm whales are also seperated         
[12]"minke whales"             
[13]"orca"                     
[14]"sperm whales"             
[15]"baleen whales" 

```{r}
species_params(params)$temp_min <- c(-1.8, # euphausiids
                                     -2, # mesopelagic fishes
                                     -2, # bathypelagic fishes
                                     -1, # "shelf and coastal fishes" 
                                     0.2, # "flying birds"
                                     -0.4, # "small divers"             
                                     -0.8, # "squids"                  
                                     -2, # "toothfishes"              
                                     -1.9, # "leopard seals"           
                                     -0.4, # "medium divers"            
                                     0.1, # "large divers"             
                                     0.2, # "minke whales"             
                                     0.3, # "orca"                     
                                     0.3, # "sperm whales"             
                                     0.2) # "baleen whales"

species_params(params)$temp_max <- c(1.6, # euphausiids
                                     5, # mesopelagic fishes
                                     5, # bathypelagic fishes
                                     5, # "shelf and coastal fishes" 
                                     18.2, # "flying birds"
                                     1, # "small divers"             
                                     1.1, # "squids"                  
                                     5, # "toothfishes"              
                                     1.4, # "leopard seals"           
                                     1, # "medium divers"            
                                     1.6, # "large divers"             
                                     7, # "minke whales"             
                                     13.1, # "orca"                     
                                     3.8, # "sperm whales"             
                                     10.2) # "baleen whales"
```



```{r}
plotThermPerformance(params)
```

