---
title: "IWC data"
output: html_notebook
---

Tidying IWC data relevant for Southern Hemisphere catch from pelagic, land based and Soviet Union corrected numbers

SHL.CSV & SHL.TXT 	Individual data and Header records for catches from the Antarctic land-stations
SHP1.CSV & SHP1.TXT	Part 1 of Individual data and Header records for pelagic catches in the Antarctic
SHP2.CSV & SHP2.TXT	Part 2 of Individual data and Header records for pelagic catches in the Antarctic
SU.CSV and SU.TXT	Revised Individual data and Header records for catches in the Antarctic taken by the Soviet Union

Load required libraries

```{r}
library(tidyverse)
library(lubridate)
```
Whale species coding for variable `Sp`

```{r}
Species.code<-c(1:22) 

# Species<-c("Pilot", "Bottlenose", "Killer", "Blue", "Fin",
#                "Sperm", "Humpback", "Sei", "Common Minke",
#                "Byrdes", "Right", "Gray", "Bairds Beaked",
#                "Baleen", "Pygmy Blue", "Pygmy Right", 
#                "Cuviers Beaked", "Bowhead", "Beaked (unspec)",
#                "Antarctic Minke", "Sei/Brydes", "Dolphin")

Species<-c("Pilot", "Bottlenose", "Killer", "Baleen", "Baleen",
               "Sperm", "Baleen", "Baleen", "Common Minke",
               "Byrdes", "Right", "Gray", "Bairds Beaked",
               "Baleen", "Pygmy Blue", "Pygmy Right", 
               "Cuviers Beaked", "Bowhead", "Beaked (unspec)",
               "Antarctic Minke", "Sei/Brydes", "Dolphin")

whale.codes<-data.frame(Species, Species.code)
```


Read in and inspect raw data frames

```{r}
# First time-series for SH ()
df_1_raw <- read.csv("IWC data/SHP1.csv")

# Second time-series for SH
df_2_raw <- read.csv("IWC data/SHP2.csv")

# Southern hemisphere land catch
df_3_raw <- read.csv("IWC data/SHL.csv")

# Soviet Union Antartic catch
df_4_raw <- read.csv("IWC data/SU.csv")

glimpse(df_1_raw)
glimpse(df_2_raw)
glimpse(df_3_raw)
glimpse(df_4_raw)

summary(df_1_raw)
summary(df_2_raw)
summary(df_3_raw)
summary(df_4_raw)
```

Streamline the dataframes and rename variables as necessary

Lat and Lon have minutes variables `Mn` associated with them, and as such the second one read in is `Mn.1`. The same goes for `AC` and there is an unnamed variable to specify if the Lat or Lon measurement is North-South or East-West respectively.

```{r}
df_1_reduced <- select(df_1_raw, "Sum.Ex", "Day", "Mon", "Year", "Sp", "Len", "L.u", "Sx", "Lat", "Mn", "Lon","Mn.1", "Mat", "Ac", "X", "X.1", "Ac.1")

df_2_reduced <- select(df_2_raw, "Sum.Ex", "Day", "Mon","Year", "Sp", "Len", "L.u", "Sx", "Lat", "Mn", "Lon","Mn.1", "Mat", "Ac", "X", "X.1","Ac.1")

df_3_reduced <- select(df_3_raw, "Sum.Ex", "Day", "Mon","Year", "Sp", "Len", "L.u", "Sx", "Lat", "Mn", "Lon","Mn.1", "Mat", "Ac", "X", "X.1", "Ac.1")

df_4_reduced <- select(df_4_raw, "Sum.Ex", "Day", "Mon","Year", "Sp", "Len", "L.u", "Sx", "Lat", "Mn", "Lon","Mn.1", "Mat", "Ac", "X", "X.1", "Ac.1")


names(df_1_reduced) <- c("Exp.no", "Day", "Month", "Year","Species.code","Length","Length.unit","Sex","Lat","Mn","Lon","Mn.1","Mat","Ac","N.S",    "E.W","Ac.1")

names(df_2_reduced) <- c("Exp.no","Day", "Month", "Year","Species.code","Length","Length.unit","Sex","Lat","Mn","Lon","Mn.1","Mat","Ac","N.S",    "E.W","Ac.1")

names(df_3_reduced) <- c("Exp.no","Day", "Month", "Year","Species.code","Length","Length.unit","Sex","Lat","Mn","Lon","Mn.1","Mat","Ac","N.S",    "E.W","Ac.1")

names(df_4_reduced) <- c("Exp.no","Day", "Month", "Year","Species.code","Length","Length.unit","Sex","Lat","Mn","Lon","Mn.1","Mat","Ac","N.S",    "E.W","Ac.1")

# write.csv(df_1_reduced, "SHP1_v1.csv")
# write.csv(df_2_reduced, "SHP2_v1.csv")
# write.csv(df_3_reduced, "SHP3_v1.csv")
# write.csv(df_4_reduced, "SHP4_v1.csv")
```


```{r}
df_comb <- rbind(df_1_reduced, df_2_reduced, df_3_reduced, df_4_reduced)

df_comb_names <- left_join(df_comb, whale.codes)
```

```{r}
glimpse(df_comb_names)

df_comb_names <- df_comb_names %>% 
  mutate(date = make_date(Year, Month, Day),
         ID = row_number()) 

glimpse(df_comb_names)

```

```{r}
df_effort <- df_comb_names %>% 
  group_by(Exp.no) %>% 
  mutate(Duration = length(unique(date))) %>% 
  group_by(Exp.no,Species) %>% 
  mutate(total_n = n(),
         Ind_CPUE = Duration/total_n)
```


```{r}
df_timeseries <- df_effort %>% 
  group_by(Year, Species) %>% 
  summarise(Catch_individuals = n())

glimpse(df_timeseries)
summary(df_timeseries)
```


```{r}
ggplot(df_timeseries, aes(x=Year, y = Catch_individuals, colour = Species)) +
  geom_line() +
  facet_wrap(~Species)

ggplot(df_timeseries, aes(x=Year, y = Catch_individuals, colour = Species)) +
  geom_line()
```



Editing location data to account for all positive lat/lon values and converting to decimal degrees
Quality control for measurements where location uncertainty present
```{r}
# df_dd <- df_comb_names
df_dd <- df_effort

##correct positive lat and long
df_dd$Lat[which(df_dd$N.S=="S")]<-df_dd$Lat[which(df_dd$N.S=="S")]*-1
df_dd$Lon[which(df_dd$E.W=="W")]<-df_dd$Lon[which(df_dd$E.W=="W")]*-1

##convert to decimal
df_dd$dec.Lat<-df_dd$Lat+(df_dd$Mn/60)
df_dd$dec.Lon<-df_dd$Lon+(df_dd$Mn.1/60)

##NA out the locational data that isnt certain:
df_dd$Lat[which(df_dd$Ac==0)]<-NA;df_dd$Lon[which(df_dd$Ac.1==0)]<-NA
```


```{r}
# world<-map_data("world")

##map the data 
# mp <- NULL
mapWorld <- borders("world", colour="gray50", fill="white") # create a layer of borders
whales<-c("Sperm", "Baleen", "Antarctic Minke", "Killer", "Dolphin")
sub.dd<-df_dd[df_dd$Species%in%whales, ]



 # png("Plots/world map.png", width=1600, height=5000)
  ggplot()+mapWorld +geom_point(data=sub.dd, aes(x=dec.Lon, y=dec.Lat, col=Species), size=0.5)+scale_colour_brewer(palette = "Set2", name = "Species")+coord_equal(ratio=1)+theme_bw()+theme(legend.position="top")+ xlab("")+ylab("")+theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.text.y = element_blank())+facet_wrap(~Species, ncol=1)
 # dev.off()
```

```{r}
library(sf)

df_sf <- st_as_sf(x = df_dd,
                  coords = c("dec.Lon","dec.Lat"),
                  crs = "+proj=longlat +datum=WGS84")

model_domain <- readRDS("model_domains/Stacey/BanzareBank.rds")

df_sf_update <- df_sf %>% st_set_crs(4326)

mod_domain_update <- model_domain %>% st_crs(4326)

df_overlap <- st_intersection(df_sf_update, model_domain)

df_BB_map <- sf_transform_xy(df_overlap, 4326,3347)

##convert to decimal
df_BB_map$dec.Lat<-df_BB_map$Lat+(df_BB_map$Mn/60)
df_BB_map$dec.Lon<-df_BB_map$Lon+(df_BB_map$Mn.1/60)

# png("Plots/BanzareBank map.png", width=1600, height=5000)
  ggplot()+mapWorld +geom_point(data=df_BB_map, aes(x=dec.Lon, y=dec.Lat, col=Species), size=0.5)+scale_colour_brewer(palette = "Set2", name = "Species")+coord_equal(ratio=1)+theme_bw()+theme(legend.position="top")+ xlab("")+ylab("")+theme(axis.ticks = element_blank(), axis.text.x = element_blank(),axis.text.y = element_blank())+facet_wrap(~Species, ncol=1)
 # dev.off()
```



```{r}
sub.dd.ts <- sub.dd %>% 
  group_by(Year, Species) %>% 
  summarise(Catch_individuals = n()) %>% 
  filter(!Catch_individuals == 0) %>% 
  filter(!is.na(Catch_individuals))

ggplot(sub.dd.ts, aes(x=Year, y = Catch_individuals, colour = Species)) +
  geom_line() +
  facet_wrap(~Species, scales = "free_y")

# tiff(file="Plots/SH_Whale_Landings.tiff", width=6, height=4, units="in", res=500)
ggplot(sub.dd.ts, aes(x=Year, y = Catch_individuals, colour = Species)) +
  geom_line(size=0.5) +
  ylab("Landings (no. of whales)") +
  # scale_y_log10() +
  theme_classic()
# dev.off()
```

```{r}
sub_df_BB_map <- df_BB_map %>% 
  group_by(Year, Species) %>% 
  summarise(Catch_individuals = n()) %>% 
  filter(!Catch_individuals == 0) %>% 
  filter(!is.na(Catch_individuals))

whales<-c("Sperm", "Baleen", "Antarctic Minke", "Killer", "Dolphin")
sub.dd.BB <- sub_df_BB_map[sub_df_BB_map$Species%in%whales, ]

ggplot(sub.dd.BB, aes(x=Year, y = Catch_individuals, colour = Species)) +
  geom_line() +
  facet_wrap(~Species, scales = "free_y")

# tiff(file="Plots/BanzareBank_Whale_Landings.tiff", width=6, height=4, units="in", res=500)
ggplot(sub.dd.BB, aes(x=Year, y = Catch_individuals, colour = Species)) +
  geom_line(size=0.5) +
  ylab("Landings (no. of whales)") +
  # scale_y_log10() +
  theme_classic()
# dev.off()
```

```{r}
library(SOmap)
# tiff(file="Plots/BanzareBank_Whale_Landings_Map.tiff", width=6, height=4, units="in", res=500)
SOmap_auto(df_BB_map$dec.Lon, df_BB_map$dec.Lat, input_lines = F, pcol = 4)
# dev.off()


# df_map <- df_BB_map %>% 
#   rename(lat = dec.Lat,
#          lon = dec.Lon)
# 
# ## first transform our track to polar-stereographic coordinates
# library(sp)
# library(raster)
# ele_sp <- df_BB_map
# coordinates(ele_sp) <- c("lon", "lat")
# projection(ele_sp) <- "+proj=longlat +ellps=WGS84"
# ele_sp <- spTransform(ele_sp,
#             CRS("+proj=stere +lat_0=-90 +lat_ts=-71 +lon_0=0 +k=1 +x_0=0 +y_0=0 
#             +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"))
# 
# ## plot the base map with ocean fronts shown
# SOmap(fronts = TRUE)
# ## add current marine protected areas
# SOmanagement(mpa = TRUE, mpa_col = "darkblue")
# ## add our track
# plot(ele_sp, col = "darkgreen", add = TRUE)
```


Body size stuff

```{r}
##function to convert ft and inches to cm
ft.in.2cm<-function(x){
	return(((as.numeric(substring(as.character(x), 1, 2))*12)+as.numeric(substring(as.character(x), 3, 4)))*2.54)
}

ft2cm<-function(x){
	return((as.numeric(substring(as.character(x), 1, 2))*12)*2.54)
}
```

```{r}
library(reshape2)
##remove spp with hardly any data:
# omit<-c("Cuviers Beaked", "Baleen", "Pygmy Blue")

##convert the ft and inches to cm

dd <- df_BB_map
# dd <- sub.dd

d1<-(melt(tapply(dd$Species, list(dd$Year, dd$Species), length)))
names(d1)<-c("Year", "Species", "Landings")

dd$Length[which(dd$Length==0000)]<-NA
dd$Length[which(dd$Length.unit==4)]<-ft.in.2cm(dd$Length[which(dd$Length.unit==4)])
dd$Length[which(dd$Length.unit==2)]<-ft2cm(dd$Length[which(dd$Length.unit==2)])

##remove shitty data (eg. massive individuals)
dd$Length[which(dd$Length>3400)]<-NA
```


W = a × L^b

Blue whale: https://www.sealifebase.ca/PopDyn/LWRelationshipList.php?ID=69007&GenusName=Balaenoptera&SpeciesName=musculus&fc=1174

Best values:
a = 0.0061
b = 3

Fin whale: https://www.sealifebase.ca/PopDyn/LWRelationshipList.php?ID=69006&GenusName=Balaenoptera&SpeciesName=physalus&fc=1174

Best values:
a = 0.00555
b = 2.95 (average of male and female from most reliable sources for Antarctic)

```{r}
df_LW_Baleen <- dd %>% 
  filter(Species == "Baleen") %>% 
  filter(!is.na(Length)) %>% 
  mutate(Weight_kg = (0.0061*(Length)^3)/1000)
```

```{r}
df_Baleen_Catch <- df_LW_Baleen %>% 
  group_by(Year) %>% 
  summarise(Catch_kg = sum(Weight_kg))
```

```{r}
ggplot(df_Baleen_Catch, aes(x=Year, y= (Catch_kg/(1000)/1e6))) +
  geom_line() +
  ggtitle("Baleen Whale Harvest in Prydz Bay/Banzare Bank") +
  ylab("Catch (Tonnes)")
```


Fin whale: https://www.sealifebase.ca/PopDyn/LWRelationshipList.php?ID=69006&GenusName=Balaenoptera&SpeciesName=physalus&fc=1174

https://www.sealifebase.ca/PopDyn/LWRelationshipList.php?ID=70932&GenusName=Balaenoptera&SpeciesName=bonaerensis&fc=1174

Best values:
a = 0.0115
b = 3 (average of male and female from most reliable sources for Antarctic)
```{r}
df_LW_Minke <- dd %>% 
  filter(Species == "Antarctic Minke") %>% 
  filter(!is.na(Length)) %>% 
  mutate(Weight_kg = (0.0115*(Length)^3)/1000)
```

```{r}
df_Ant_Minke_Catch <- df_LW_Minke %>% 
  group_by(Year) %>% 
  summarise(Catch_kg = sum(Weight_kg))
```

```{r}
ggplot(df_Ant_Minke_Catch, aes(x=Year, y= (Catch_kg/(1000)/1000000))) +
  geom_line() +
  ggtitle("Antarctic Minke Whale Harvest in Prydz Bay/BB") +
  ylab("Catch (Million Tonnes)")
```


Sperm whale: https://www.sealifebase.ca/PopDyn/LWRelationshipList.php?ID=68882&GenusName=Physeter&SpeciesName=macrocephalus&fc=1182

Best values:
a = 0.0109
b = 3 (unsexed)
```{r}
df_LW_Sperm <- dd %>% 
  filter(Species == "Sperm") %>% 
  filter(!is.na(Length)) %>% 
  mutate(Weight_kg = (0.0109*(Length)^3)/1000)
```

```{r}
df_Sperm_Catch <- df_LW_Sperm %>% 
  group_by(Year) %>% 
  summarise(Catch_kg = sum(Weight_kg))
```

```{r}
ggplot(df_Sperm_Catch, aes(x=Year, y= (Catch_kg/(1000)/1000000))) +
  geom_line() +
  ggtitle("Sperm Whale Harvest in Prydz Bay/BB") +
  ylab("Catch (Million Tonnes)")
```

Killer whale: https://www.sealifebase.ca/PopDyn/LWRelationshipList.php?ID=69400&GenusName=Orcinus&SpeciesName=orca&fc=1175

```{r}
 # max length = 980

# 0.2080*(Length)^2.577)/1000


(0.2080*(980)^2.577) # max length expressed in grams

(0.2080*(615)^2.577) # mat length expressed in grams (used midpoint of the Lm range)
```


Best values:
a = 0.2080
b = 2.577 (mixed sex)
```{r}
df_LW_Orca <- dd %>% 
  filter(Species == "Killer") %>% 
  filter(!is.na(Length)) %>% 
  mutate(Weight_kg = (0.2080*(Length)^2.577)/1000)
```

```{r}
df_Orca_Catch <- df_LW_Orca %>% 
  group_by(Year) %>% 
  summarise(Catch_kg = sum(Weight_kg))
```

```{r}
ggplot(df_Orca_Catch, aes(x=Year, y= (Catch_kg/(1000)/1000000))) +
  geom_line() +
  ggtitle("Killer Whale Harvest in Prydz Bay/BB") +
  ylab("Catch (Million Tonnes)")
```


Dolphin: https://www.sealifebase.ca/PopDyn/LWRelationshipList.php?ID=69376&GenusName=Tursiops&SpeciesName=truncatus&fc=1175

Best values:
a = 0.2754
b = 2.73 (mixed sex)
```{r}
df_LW_Dolphin <- dd %>% 
  filter(Species == "Dolphin") %>% 
  filter(!is.na(Length)) %>% 
  mutate(Weight_kg = (0.2754*(Length)^2.73)/1000)
```

```{r}
df_Dolphin_Catch <- df_LW_Dolphin %>% 
  group_by(Year) %>% 
  summarise(Catch_kg = sum(Weight_kg))
```

```{r}
ggplot(df_Dolphin_Catch, aes(x=Year, y= (Catch_kg/(1000)/1000000))) +
  geom_line() +
  ggtitle("Dolphin Harvest in Prydz Bay/BB") +
  ylab("Catch (Million Tonnes)")
```

Combined dataframe of individuals

```{r}

df_ind_weight <- rbind(df_LW_Sperm,df_LW_Baleen,df_LW_Minke, df_LW_Orca)

glimpse(df_ind_weight)
summary(df_ind_weight)

# saveRDS(df_ind_weight, "ind_catch_weight_BanzareBank_1930_2019.rds")
df_ind <- readRDS("ind_catch_weight_BanzareBank_1930_2019.rds")
summary(df_ind)

df_ind_CPUE <- as.data.frame(df_ind) %>% 
  dplyr::select(ID, Exp.no, Year, Species, Length, Length.unit, Weight_kg, date) %>% 
  group_by(Exp.no, Species) %>% 
  mutate(Duration = length(unique(date)))

glimpse(df_ind_CPUE)

df_CPUE_kg_day <- df_ind_CPUE %>% 
  group_by(Exp.no, Species) %>% 
  mutate(Sp_total_kg = sum(Weight_kg)) %>% 
  ungroup() %>% 
  mutate(Ind_UnitEffort_day = Weight_kg/(Sp_total_kg/Duration)) %>% 
  group_by(Year, Species) %>% 
  summarise(total_catch_kg = sum(Weight_kg),
            effort_days = sum(Ind_UnitEffort_day),
            CPUE = total_catch_kg/effort_days)

glimpse(df_CPUE_kg_day)

# saveRDS(df_ind_CPUE, "ind_catch_weight_BanzareBank_1930_2019_CPUE.rds")
# saveRDS(df_CPUE_kg_day, "catch_timeseries_BanzareBank_1930_2019_CPUE.rds")
  
```

```{r}
ggplot(df_CPUE_kg_day, aes(x = Year, y = CPUE, colour = Species)) +
    geom_line() +
   # scale_y_log10() +
    facet_wrap(~Species)

ggplot(df_CPUE_kg_day, aes(x = Year, y = effort_days, colour = Species)) +
    geom_line() +
    facet_wrap(~Species,scales = "free")

ggplot(df_CPUE_kg_day, aes(x = Year, y = total_catch_kg, colour = Species)) +
    geom_line() +
    facet_wrap(~Species, scales = "free")
```


Combine harvest data

```{r}
df_Sperm_Catch$Species <- "Sperm"
df_Baleen_Catch$Species <- "Baleen"
df_Ant_Minke_Catch$Species <- "Antarctic Minke"
df_Orca_Catch$Species <- "Killer"
# df_Dolphin_Catch$Species <- "Dolphin"

df_harvest <- rbind(df_Sperm_Catch, df_Baleen_Catch, df_Ant_Minke_Catch, df_Orca_Catch) %>% 
  select(Year, Catch_kg, Species)

# write.csv(df_catch_timeseries, "catch_timeseries_BanzareBank_1930_2019.csv")
```


```{r}
df_harvest <- read.csv("catch_timeseries_BanzareBank_1930_2019.csv")

ggplot(df_harvest, aes(x=Year, y= (Catch_kg/(1000)/1e6), colour = Species)) +
  geom_line() +
  ggtitle("Prydz Bay Whale Harvest") +
  # facet_wrap(~Species) +
  # scale_y_log10() +
  ylab("Catch (Million Tonnes)")

```


```{r}
# ##calculate means per spp per yr
# szs<-(melt(tapply(dd$Length, list(dd$Year, dd$Species), mean, na.rm=T)))
# szs<-szs[!szs$Var2%in%omit,]
# d1$Mean.size<-szs$value
# 
# ##calculate size sd per yr
# szs.sd<-(melt(tapply(dd$Length, list(dd$Year, dd$Species), sd, na.rm=T)))
# szs.sd<-szs.sd[!szs.sd$Var2%in%omit,]
# d1$SD.size<-szs.sd$value
# 
# ##calculate the 95 percentile 
# q<-function(x){quantile(na.omit(x), 0.95)}
# szs.95<-(melt(tapply(dd$Length, list(dd$Year, dd$Species), q)))
# szs.95<-szs.95[!szs.95$Var2%in%omit,]
# d1$size.95<-szs.95$value
# 
# ##calculate L-MAX
# lmx<-function(x){max(na.omit(x))}
# lmax<-(melt(tapply(dd$Length, list(dd$Year, dd$Species), lmx)))
# lmax<-lmax[!lmax$Var2%in%omit,]
# d1$lmax<-lmax$value
```


```{r}
# ggplot(na.omit(d1), aes(x=Year, y=size.95))+geom_point(col="maroon")+geom_line(col="red")+theme_bw()+facet_wrap(~Species)+theme(legend.position="none")+geom_vline(xintercept=1985)+stat_smooth(col="black")+xlab("Year")+ylab("Mean size (cm)")+thm+xlim(1900, 1985)

```
Length-weight Relationships

Guide for using fishbase/sealifebase value for L-W relationships.

https://www.fishbase.de/manual/fishbasethe_length_weight_table.htm#:~:text=The%20LENGTH%2DWEIGHT%20table%20presents,about%20over%202%2C000%20fish%20species.

We have length for individuals above, so need to apply it to the individual data on a species by species basis


```{r}

```


## Leopard seal

```{r}

 # max length = 300


(0.0117*(360)^3) # max length expressed in grams

# (0.2080*()^2.577) # mat length expressed in grams (used midpoint of the Lm range)
```
