---
title: "Figure: Predator-prey size relations"
author: "Shane A. Richards + Kieran Murphy"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: show
    highlight: tango
    theme: cerulean
    toc: yes
    toc_float: no
  word_document:
    toc: yes
  pdf_document:
    toc: yes
---

This is a run through of extracting the size structured diet output from one variant of the Feeding + Growth Model. 
After extraction, these data are reformatted to be suitable for a plot of mean prey size vs predator size to establish emergent PPMR values for each functional group

## Setup workspace
```{r}
# install.packages('tidyverse')
# install.packages('reshape2')
library(tidyverse)
library(reshape2)
```

## Import Model

```{r}
# model <-readRDS("FG_Model_v5.rds")

# saveRDS(model$param, "FG_Model_v5_params.RDS")
# model$param
```

```{r}
##### PHYTO DIET edited by Ryan and Kieran July 2020 to save abundance of phyto at each phyto size class for each pred size class
##### PHYTO DIET DIM1 = SAVE STEP, DIM2 = PRED GROUP, DIM3 = PRED SIZE, DIM4 = Phyto size
##### DYNAM DIET DIM1 = SAVE STEP, DIM2 = PRED GROUP, DIM3 = PRED SIZE, DIM 4 = PREY GROUP, DIM 5 = PREY SIZE

 ## Average diet array for each predator size class for the final half of the projection
  dynamic_diet = apply(model$dynam_diet_full[c(floor(0.5*dim(model$dynam_diet_full)[1]):dim(model$dynam_diet_full)[1]),,,,], c(2,3,4,5), mean)
  phyto_diet = apply(model$phyto_diet_full[c(floor(0.5*dim(model$phyto_diet_full)[1]):dim(model$phyto_diet_full)[1]),,,], c(2,3,4), mean)
  
```


```{r}
saveRDS(dynamic_diet, "dynamic_diet.RDS")
saveRDS(phyto_diet, "phyto_diet.RDS")


dynamic_diet <- readRDS("dynamic_diet.RDS")
phyto_diet <- readRDS("phyto_diet.RDS")

```


## Phytoplankton Diet

```{r}
phytodiet <- melt(phyto_diet)
glimpse(phytodiet)
```


```{r}
  ## Assign names to all variables in phytodiet
names(phytodiet) <- c("Predator_Species", "Predator_Size_Class", "Prey_Size_Class", "Prey_Abundance")
glimpse(phytodiet)
```

```{r}
## make predators and prey factor variables
phytodiet$Predator_Species <- as.factor(phytodiet$Predator_Species)

## Confirm the variables are correct
summary(phytodiet)
```


```{r}
  
  
  df_size_classes <- model$w # Create new object with list of size classes in log10 g
  df_size_classes <- as.data.frame(df_size_classes) # Convert to a dataframe
  names(df_size_classes) <- c("Predator_Size") # Create new name Size for the variable
  df_size_classes$Predator_Size_Class <- 1:length(model$w) # Create Size_class variable to join with phytodiet by
  
  glimpse(df_size_classes)
```

```{r}
  df_species_predator <- model$param$groups$species # Create new object with species names from model
  df_species_predator <- as.data.frame((df_species_predator)) # Convert to a dataframe
  names(df_species_predator) <- c("Predator") # Rename variable
  df_species_predator$Predator_Species <- 1:length(model$param$groups$species) # Create new integer variable in ascending order with 'Species'
  df_species_predator$Predator_Species <- as.factor(df_species_predator$Predator_Species) # Convert to factor
  
  glimpse(df_species_predator)
```

```{r}
  df_phyto_size_classes <- model$w_phyto # Create new object with list of size classes in log10 g
  df_phyto_size_classes <- as.data.frame(df_phyto_size_classes) # Convert to a dataframe
  names(df_phyto_size_classes) <- c("Prey_Size") # Create new name Size for the variable
  df_phyto_size_classes$Prey_Size_Class <- 1:length(model$w_phyto)
 
glimpse(df_phyto_size_classes) 
```

```{r}
## join phytodiet with df_size_class data in order to create log10 size classes in addition to the size class integer 1:184
  df_combined <- left_join(phytodiet, df_size_classes,
                           by="Predator_Size_Class")

glimpse(df_combined)
```

```{r}
## join phytodiet with df_species_predator data in order to create predator name variable in addition to the integer 1:4
df_combined <- df_combined %>%
    left_join(df_species_predator,
              by="Predator_Species")

glimpse(df_combined)
```

```{r}
df_combined <- df_combined %>%
    left_join(df_phyto_size_classes,
              by="Prey_Size_Class")

df_combined <- df_combined %>%
    mutate(Prey_Biomass = Prey_Abundance*Prey_Size)

df_combined$Prey <- c("Phyto")

df_combined$Prey <- as.factor(df_combined$Prey)

glimpse(df_combined)
```

```{r}
summary(df_combined)
```

```{r}
# df_combined <- df_combined %>%
#   filter(Prey_Abundance > 0)

summary(df_combined)
```

## Dynamic Diet

```{r}
 ### Dynamic diet dataframe
dynamdiet <- melt(dynamic_diet)
  
glimpse(dynamdiet)
```

```{r}
  ## Assign names to all variables in df_diet
names(dynamdiet) <- c("Predator_Species", "Predator_Size_Class", "Prey_Species", "Prey_Size_Class", "Prey_Abundance")

# make predators and prey factor variables
dynamdiet$Predator_Species <- as.factor(dynamdiet$Predator_Species)
dynamdiet$Prey_Species <- as.factor(dynamdiet$Prey_Species)
  
glimpse(dynamdiet)
```
  
```{r}
  # df_size_classes <- model$w # Create new object with list of size classes in log10 g
  # df_size_classes <- as.data.frame(df_size_classes) # Convert to a dataframe
  # names(df_size_classes) <- c("Predator_Size") # Create new name Size for the variable
  # df_size_classes$Predator_Size_Class <- 1:length(model$w)
```


```{r}
  df_species_prey <- model$param$groups$species # Create new object with species names from model
  df_species_prey <- as.data.frame((df_species_prey)) # Convert to a dataframe
  names(df_species_prey) <- c("Prey") # Rename variable
  df_species_prey$Prey_Species <- 1:length(model$param$groups$species) # Create new integer variable in ascending order with 'Species'
  df_species_prey$Prey_Species <- as.factor(df_species_prey$Prey_Species) # Convert to factor

glimpse(df_species_prey)
```

```{r}
  df_prey_size_classes <- model$w # Create new object with list of size classes in log10 g
  df_prey_size_classes <- as.data.frame(df_prey_size_classes) # Convert to a dataframe
  names(df_prey_size_classes) <- c("Prey_Size") # Create new name Size for the variable
  df_prey_size_classes$Prey_Size_Class <- 1:length(model$w)

glimpse(df_prey_size_classes) 
```

```{r}
  ## join dynamdiet with df_size_class data in order to create log10 size classes in addition to size class integer 1:184 
  df_combined_dynam <- left_join(dynamdiet, df_size_classes,
                           by="Predator_Size_Class")

glimpse(df_combined_dynam)
```

```{r}
df_combined_dynam <- df_combined_dynam %>%
    left_join(df_species_predator,
              by="Predator_Species")

glimpse(df_combined_dynam)
```

```{r}
df_combined_dynam <- df_combined_dynam %>%
  left_join(df_prey_size_classes,
              by="Prey_Size_Class")

glimpse(df_combined_dynam)
```


```{r}
df_combined_dynam <- df_combined_dynam %>%
  left_join(df_species_prey,
              by="Prey_Species")

df_combined_dynam <- df_combined_dynam %>%
    mutate(Prey_Biomass = Prey_Abundance*Prey_Size)
  
glimpse(df_combined_dynam)
```

### Combine reformatted dynamic and phyto diet output into single dataframe

```{r}

## Remove integer variables and row bind the two dataframes

  df_d <- df_combined_dynam %>%
    select(Predator, Prey, Predator_Size, Prey_Size, Prey_Abundance, Prey_Biomass)
  df_pd <- df_combined %>%
    select(Predator, Prey, Predator_Size, Prey_Size, Prey_Abundance, Prey_Biomass)
  
  df <- rbind(df_d, df_pd) # join data frames, one on top of the other

glimpse(df)
summary(df)
```



```{r}
# reorder factors
df$Predator <- factor(df$Predator, 
  levels = c("Zooplankton", "Fish", "Inactive Cephs", "Active Cephs"))

df$Prey <- factor(df$Prey, 
  levels = c("Phyto", "Zooplankton", "Fish", "Inactive Cephs", "Active Cephs"))

unique_sizes <- sort(unique(c(df$Prey_Size, df$Predator_Size)))

df_sizes <- tibble(
  Prey_Size = unique_sizes, 
  Prey_size_class = 1:length(unique_sizes)
)

df <- left_join(df, df_sizes, by = "Prey_Size") %>%
  arrange(Predator, Prey, Predator_Size, Prey_Size)

df_sizes <- tibble(
  Predator_Size = unique_sizes, 
  Predator_size_class = 1:length(unique_sizes)
)

df <- left_join(df, df_sizes, by = "Predator_Size") %>%
  arrange(Predator, Prey, Predator_Size, Prey_Size)

glimpse(df)
summary(df)
```

# Prey size distributions

```{r fig.width=9, fig.height=3}
# examine prey size distribution for a specified predator-prey
df_test <- df %>%
  filter(Predator == "Inactive Cephs", Prey == "Zooplankton")
  # filter(Predator == "Inactive Cephs")

max_size_class <- max(df_test$Predator_size_class)

class_plot <- max_size_class - 40 # predator size class to focus on

df_test <- df_test %>%
  filter(Predator_size_class == class_plot)

# calculate cumulative distribution for calculating quantiles
df_test$cumfreq <- cumsum(df_test$Prey_Abundance) /
  sum(df_test$Prey_Abundance)
# calculate the approximate quantiles
q05 <- df_test$Prey_Size[df_test$cumfreq > 0.05][1]
q50 <- df_test$Prey_Size[df_test$cumfreq > 0.5][1]
q95 <- df_test$Prey_Size[df_test$cumfreq > 0.95][1]

ggplot(df_test, aes(x = Prey_Size, y = Prey_Abundance)) +
  geom_area(fill = "salmon") +
  geom_point(color = "red") +
  geom_vline(xintercept = df_sizes$Predator_Size[class_plot], 
    color = "red") +
  geom_vline(xintercept = q05, linetype = "dashed") +
  geom_vline(xintercept = q50) +
  geom_vline(xintercept = q95, linetype = "dashed") +
  scale_x_log10() +
  labs(x = "Prey size", y = "Prey abundance") +
  theme_bw()
```

X-axis is prey (`Zooplankton`) size. Vertical red line is the predator (`Inactive Cephs`) size  (only one predator size class is considered for this plot). Dashed lines are the 90% prey size interval and solid black line is the median prey size.


```{r fig.width=9, fig.height=3}
library(viridis)
# examine prey size distribution for a specified predator-prey
df_test <- df %>%
  filter(Predator == "Inactive Cephs", Prey == "Zooplankton")
  # filter(Predator == "Inactive Cephs")

df_total <- df_test %>%
  group_by(Predator_size_class) %>%
  summarise(
    sum_prey_abundance = sum(Prey_Abundance)
  )

df_test <- left_join(df_test, df_total, by = "Predator_size_class") %>%
  mutate(norm_freq = Prey_Abundance / sum_prey_abundance)

ggplot(df_test, 
  aes(x = Predator_size_class, y = Prey_size_class, fill = norm_freq)) +
  geom_tile() +
  scale_fill_viridis(option="magma") +
  labs(x = "Predator size class", y = "Prey size class") +
  theme_bw()
```


```{r}
# examine prey size distribution for a specified predator-prey
df_test <- df

df_total <- df_test %>%
  group_by(Predator_size_class, Predator, Prey) %>%
  summarise(
    sum_prey_abundance = sum(Prey_Abundance)
  )

df_test <- left_join(df_test, df_total, 
    by = c("Predator_size_class", "Predator","Prey")) %>%
  mutate(norm_freq = Prey_Abundance / sum_prey_abundance)

```


```{r}
ggplot(df_test, 
  aes(x = Predator_size_class, y = Prey_size_class, fill = norm_freq)) +
  geom_tile() +
  scale_fill_viridis(option="magma", direction = -1, na.value ="pink") +
  labs(x = "Predator size class", y = "Prey size class") +
  facet_grid(Prey ~ Predator) +
  theme_bw()
```


```{r}
# create a data frame for plotting
df_plot <- df %>%
  mutate(PySzPyAb = Prey_Size*Prey_Abundance) %>%  # total size
  group_by(Predator, Prey, Predator_Size) %>% # groups of prey distributions
  summarise(
    sum_prey_abundance = sum(Prey_Abundance), # total prey numbers
    sum_PySzPyAb       = sum(PySzPyAb), # total size
    mean_prey_size     = sum_PySzPyAb / sum_prey_abundance # mean size
  )

# data for dashed line where expected prey size is 10% of predator size
df_ref <- tibble(
  Predator_Size = c(min(df$Predator_Size), max(df$Predator_Size))
)
df_ref$Expected_Prey_Size = 0.1*df_ref$Predator_Size
```



```{r}
ggplot() +
  geom_line(data = df_ref, aes(x = Predator_Size, y = Expected_Prey_Size),
    linetype = "dashed") +
  geom_line(data = df_plot,
    aes(x = Predator_Size, y = mean_prey_size, color = Prey)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap( ~ Predator) +
  labs(x = "Predator size", y = "Mean prey size") +
  theme_bw()
```


```{r}
# calculate total prey abundance for each predator size
df_total_intake <- df %>%
  mutate(PySzPyAb = Prey_Size*Prey_Abundance) %>% # total size
  group_by(Predator, Predator_Size) %>%
  summarise(
    total_prey_size = sum(PySzPyAb)
  )

df_plot <- left_join(df_plot, df_total_intake,
  by = c("Predator", "Predator_Size")) %>%
  mutate(frac_total = sum_PySzPyAb / total_prey_size) # prey fraction

```

```{r}
frac_keep <- 0.001 # plot prey if contributes at least this fraction of diet

ggplot() +
  geom_line(data = df_ref, aes(x = Predator_Size, y = Expected_Prey_Size),
    linetype = "dashed") +
  geom_line(data = filter(df_plot, frac_total > frac_keep), 
    aes(x = Predator_Size, y = mean_prey_size, color = Prey)) +
  scale_x_log10() +
  scale_y_log10() +
  facet_wrap( ~ Predator) +
  labs(x = "Predator size", y = "Mean prey size") +
  theme_bw()
```

The horizontal lines don't appear to actually contribute to the diet , as they are less than 0.1 % for any given prey and size class.

I would like to redo the combined dataframe below after removing this fraction but I haven't quite worked out how to do that yet when `df_plot` is created by using only `Predator` and `Predator_Size` in group_by below.


## Combined Predator Diet

```{r}
# create a data frame for plotting
df_plot <- df %>%
  mutate(PySzPyAb = Prey_Size*Prey_Abundance) %>%  # total size
  group_by(Predator, Predator_Size) %>% # groups of prey distributions
  summarise(
    sum_prey_abundance = sum(Prey_Abundance), # total prey numbers
    sum_PySzPyAb       = sum(PySzPyAb), # total size
    mean_prey_size     = sum_PySzPyAb / sum_prey_abundance # mean size
  )

# data for dashed line where expected prey size is 10% of predator size
df_ref <- tibble(
  Predator_Size = c(min(df$Predator_Size), max(df$Predator_Size))
)
df_ref$Expected_Prey_Size_10000 = 0.0001*df_ref$Predator_Size
df_ref$Expected_Prey_Size_1000 = 0.001*df_ref$Predator_Size
df_ref$Expected_Prey_Size_100 = 0.01*df_ref$Predator_Size
df_ref$Expected_Prey_Size_10 = 0.1*df_ref$Predator_Size
df_ref$Expected_Prey_Size_1 = 1*df_ref$Predator_Size
```



```{r fig.width=5, fig.height=5}
ggplot() +
  geom_line(data = df_ref, aes(x = Predator_Size, y = Expected_Prey_Size_1),
    linetype = "solid", size = .75) +
  geom_line(data = df_ref, aes(x = Predator_Size, y = Expected_Prey_Size_10),
    linetype = "longdash", size = .75) +
  geom_line(data = df_ref, aes(x = Predator_Size, y = Expected_Prey_Size_100),
    linetype = "dashed", size = .75) +
  geom_line(data = df_ref, aes(x = Predator_Size, y = Expected_Prey_Size_1000),
    linetype = "dotted", size = .75) +
  geom_line(data = df_ref, aes(x = Predator_Size, y = Expected_Prey_Size_10000),
    linetype = "dotdash", size = .75) +
  geom_line(data = df_plot,
    aes(x = Predator_Size, y = mean_prey_size, color = Predator), size = 1.5) +
  scale_x_log10() +
  scale_y_log10() +
  # facet_wrap( ~ Predator) +
  labs(x = "Predator size", y = "Mean prey size") +
  theme_bw() +
  theme(legend.key.size = unit(1, 'lines'),
          legend.position=c(0.25,0.8),
          legend.background = element_blank())
```

```{r}
# ggsave("Feeding+Growth_v5_PPMR.png",
#        units="mm",
#        dpi = 600,
#        width = 90,
#        height = 90)
```




