---
title: "1f_simplified_groups_params_KJM.R"
output: html_notebook
---

https://github.com/juliablanchard/nsea_calibration/blob/main/from-steady-state-to-time-series.Rmd

Determining traits for simplified groups

# Species Groupings

## Zooplankton

### Original groups:

 - Zooplankton
 - Euphausiids
 - salps
 - other macrozooplankton
 - mesozooplankton
 - microzooplankton

### Simplified groups:

 - macrozooplankton
 - mesozooplankton
 - microzooplankton
 
### Further reduced

 - Krill
 
The level of complexity involved in resolving the smallest organisms as part of the dynamic size spectrum has proved difficult enough that we will delay the inclusion until a later date in the project. Main emphasis is on how changes to the structure and function will effect krill and as such we will prioritise their inclusion in the model

## Fishes

### Original groups:

 - small myctophids (epipelagic)
 - large myctophids (migrating)
 - gonostomatiids (mid and lower mesopelagic)
 - bathylagiids (migrating bathypelagics)
 - other mesopelagics
 - other bathypelagics (non-migrating bathypelagis)
 - shelf and coast demersal (icefish, silverfish)
 - toothfish (large pelagic-foraging demersal; shelf and slope)
	
### Simplified groups:

 - mesopelagic fishes
 - bathypelagic fishes
 - shelf and coastal fishes
 - toothfishes

## Squids

### Original groups:
 - small nektonic
 - pelagic
	
### Simplified groups:

 - squids

## Bird and mammal "top predators"

### Minimum set of:
 - baleen whales
 - toothed whales and large seals
 - small seals and large penguins
 - small penguins and other birds
 - apex predators (orca and leopard seals)
 
### Simplified groups:
 - flying birds
 - small divers
 - medium divers
 - apex predators
 - large divers
 - baleen whales

# for each group, initially need
	# group name
	# min and max size
	# depth
	# water column use
	# preferred PPMR
	
# Functional Groups

Order may vary, as will be ordered by asymptotic size

1)  krill

2)  mesopelagic fishes
3)  bathypelagic fishes
4)  shelf and coastal fishes
5)  toothfishes

6)  squids

7)  flying birds
8)  small divers
9)  medium divers
10) apex predators
11) large divers
12) baleen whales

Rowan's notes
- some growth and consumption values were collated for the size spectrum modelling workshop
- Q/B and P/B estimates have also been collated by Stacey McCormack on SOKI for the Prydz Bay Ecopath model
- grazing rates have also been collated on SOKI for the Atlantis model initialization
- daily maximum rations for krill in %C/day, along with lots of detail on feeding ecology, are in Schmidt and Atkinson chapter in Siegel 2016 book.


```{r}
library(readxl)
library(tidyverse)
# remotes::install_github("SCAR/solong")
# library(solong)
# remotes::install_github("ropensci/rfishbase")
# library(rfishbase)
```

Use `solong` package (https://github.com/SCAR/solong) to estimate weights where data is available

```{r}
sol_properties()
x <- tibble(TL = c(0.7, 61), species = "Euphausia superba")
x

```

```{r}
x$TL <- sol_set_property(x$TL, "total length")
```


```{r}
# sol_equations() %>% dplyr::filter(taxon_name == "Euphausia superba") %>% summary # Antarctic krill
# sol_equations() %>% dplyr::filter(taxon_name == "Euphausia crystallorophias") %>% summary # Crystal krill
# sol_equations() %>% dplyr::filter(taxon_name == "Thysanoessa macrura") %>% summary 
# sol_equations() %>% dplyr::filter(taxon_name == "Leptonychotes weddellii") %>% summary # Weddell seal

```


```{r}
sol_allometry(x, "236217_WW~TL_Farb1994")
```


# Krill

https://www.ccamlr.org/en/system/files/science_journal_papers/09-Siegel.pdf

https://link.springer.com/article/10.1007/s00300-021-02984-4 (highlights the range of lenght-weight relationships and the uncertianty in size parameters). Can we model this uncertainty?

Initial zooplankton dataframe with minimum and maximum sizes in mm and length-weight coefficients `a` and `b`

```{r}
# simplified zooplankton groups
zsp <- read_xlsx("./excel data/zoop_parameters_v1.xlsx", sheet="zoop simple groups", skip = 2)
zsp

df_z <- zsp %>% 
  filter(group == "euphausiids") %>% 
  mutate(max_size_mm = as.numeric(max_size_mm),
         min_size_mm = as.numeric(min_size_mm),
         lw_a = as.numeric(lw_a),
         lw_b = as.numeric(lw_b),
         w_max = lw_a*max_size_mm^lw_b/1000, 
         w_min = lw_a*min_size_mm^lw_b/1000)
  
df_z
```

```{r}
# sizes to weights
z.prm.s<- df_z %>% 
	select(group, max_depth, min_depth, beta, Wmax, Wmat, W0) %>% rename(species=group)

z.prm.s$h <- c(33)
# - max for krill: 33 (from Surma et al 2014; Meyer estimate as high as 102 for juvs)
# - max for mesozoop: 39 (Stacey compilation)


z.prm.s
```


There are a lot of different param files and within them a lot of different size metrics, particularly for max size

I need to inspect the original values, compare them to what we have been using and correct any obvious issues.

`fish_params_wide.RDS` is the original param file used to calculate the fish group parameters

```{r}
fish_params_wide <- readRDS("fish_params_wide.RDS")
glimpse(fish_params_wide)
fish_params_wide
```

I want to reorder the species for ease of comparison with the pasted parameter list in `1a_fish_trait_calc.R` starting on line 53

Here they are for ease of use
 
	# 		              species wmax.calc     wmax    w_mat  p_mat_max p_calc_max p_calc_max_mat
	#                       <chr>     <dbl>    <dbl>   <dbl>      <dbl>      <dbl>          <dbl>
	# 1     Benthalbella elongata     494.3       NA   107.7         NA         NA     0.21781820		#  looks OK, wmax calc maybe a bit big
	# 2   Benthalbella macropinna      79.8       NA    18.8         NA         NA     0.23613383		#  looks OK
	# 3    Bathylagus antarcticus      24.9     51.0     7.2 0.14095284  0.4888120     0.28835795		#  PROBLEM: wmax calc is 1/2 wmax
	# 4       Cynomacrurus piriei     603.7       NA   110.3         NA         NA     0.18276597		#  looks ok
	# 5   Champsocephalus gunnari    2422.2   2000.0    76.2 0.03807680  1.2110952     0.03143997		#  PROBLEM: w_mat tiny relative to wmax; but checked and correct
	# 6     Gymnoscopelus braueri      15.2     21.0     9.7 0.46027982  0.7229740     0.63664779		#  
	# 7    Gymnoscopelus nicholsi      60.4     51.0     8.7 0.16978336  1.1847070     0.14331253		#  
	# 8    Nannobrachium achirus       28.5       NA     8.0         NA         NA     0.27873775		#  
	# 9       Cyclothone microdon       1.6       NA     0.1         NA         NA     0.08699037		#  PROBLEM: w_mat looks pretty tiny; checked, and driven purely by lw params
	# 10       Melanonus gracilis      36.1       NA     9.5         NA         NA     0.26381340		#  
	# 11         Notolepis coatsi     237.8     24.0    49.3 2.05481675  9.9098538     0.20735086		#  PROBLEM: w_mat is 2x wmax; wmax.calc huge
	# 12 Paradiplospinus gracilis     161.2    189.0    33.8 0.17870200  0.8530738     0.20948013		#  looks good
	# 13 Pleuragramma antarcticum     156.7    200.0    17.3 0.08647293  0.7837103     0.11033787		#  CHECK: w_mat looks a bit small
	# 14     Electrona antarctica      15.1     15.0     5.4 0.35751671  1.0097588     0.35406151		#  looks good
	# 15     Electrona carlsbergi       4.5     14.0     2.3 0.16205880  0.3224837     0.50253327		#  PROBLEM: wmax calc is 1/3 wmax
	# 16    Gymnoscopelus fraseri        NA       NA  2944.5         NA         NA             NA		#  PROBLEM: w_mat looks much to big; why no wmax?? ### Was a typo; 73 cm for w_mat rather than 7.3
	# 17 Krefftichthys anderssoni       1.8      4.0     0.7 0.18748138  0.4403679     0.42573807		#  PROBLEM: wmax calc is 1/2 observed
	# 18    Protomyctophum bolini       3.8      3.9     1.6 0.41625003  0.9725315     0.42800674		#  looks good
	# 19 Dissostichus eleginoides  157569.3 135000.0 11734.9 0.08692498  1.1671800     0.07447436		#  looks OK
	# 20     Dissostichus mawsoni  121611.0  70000.0 13795.5 0.19707908  1.7372996     0.11343989		#  looks OK			

This is to reorder the `fish_params_wide.rds` data so I can compare more easily with Rowan's trouble shooting notes pasted above.
```{r}
full_name_order <- c("Benthalbella elongata","Benthalbella macropinna","Bathylagus antarcticus","Cynomacrurus piriei","Champsocephalus gunnari",
           "Gymnoscopelus braueri","Gymnoscopelus nicholsi","Nannobrachium achirus","Cyclothone microdon","Melanonus gracilis","Notolepis coatsi",
           "Paradiplospinus gracilis","Pleuragramma antarcticum","Electrona antarctica","Electrona carlsbergi","Gymnoscopelus fraseri","Krefftichthys anderssoni","Protomyctophum bolini","Dissostichus eleginoides","Dissostichus mawsoni")

abbrev_order <- c("B.elongata","B.macropinna","B.antarcticus","C.piriei","C.gunnari",
           "G.braueri","G.nicholsi","N.achirus","C.microdon","M.gracilis","N.coatsi",
           "P.gracilis","P.antarcticum","E.antarctica","E.carlsbergi","G.fraseri","K.anderssoni","P.bolini","D.eleginoides","D.mawsoni")
```


Duplicate original
```{r}
df_fish <- fish_params_wide
```

Rearrange df_fish to double check with Rowan's original values and see what has been adjusted
```{r}
df <- df_fish %>% 
  arrange(factor(species, levels = abbrev_order))

df
```

Reduce variables for ease of comparison 
```{r}
# df_f <- df %>% 
# 	select(species, lw_a, lw_b, egg_mass_g, w_inf,wmax, w_mat, wmax.calc)
# 
# df_f
```

```{r}
fish_w_max_updated <- c(
494.3,
79.8,
51,
603.7,
2422.2,
21,
52.5,
28.5,
1.6,
36.1,
240,
189,
200,
15,
14,
6.1,
4,
3.9,
157569.3,
141744.8
)

df$w_max <- fish_w_max_updated

df
```

```{r}

species_order <- c("B.antarcticus","B.elongata","B.macropinna","C.piriei","C.gunnari","G.braueri","G.nicholsi","N.achirus","C.microdon","M.gracilis","N.coatsi","P.gracilis","P.antarcticum","E.antarctica","K.anderssoni","P.bolini","E.carlsbergi","G.fraseri","D.mawsoni","D.eleginoides")

f.df <- df %>% 
  arrange(factor(species, levels = species_order))

f.df
```


```{r}
# # initial simplified groups
# f.sgroups<- data.frame(
  
# species=c("B.antarcticus","B.elongata","B.macropinna","C.piriei","C.gunnari","G.braueri","G.nicholsi","N.achirus","C.microdon","M.gracilis","N.coatsi","P.gracilis","P.antarcticum","E.antarctica","K.anderssoni","P.bolini","E.carlsbergi","G.fraseri","D.mawsoni","D.eleginoides"), 
# 
# simple_group=c("bathylagiids","other bathypelagics","other bathypelagics","other bathypelagics","shelf and coastal fishes","large myctophids","large myctophids","large myctophids","gonostomatiids","other mesopelagics","other mesopelagics","other mesopelagics","shelf and coastal fishes","small myctophids","small myctophids","small myctophids","small myctophids","small myctophids","toothfish","toothfish"))

# even simpler groups
f.sgroups<- data.frame(

species=c("B.antarcticus","B.elongata","B.macropinna","C.piriei","C.gunnari","G.braueri","G.nicholsi","N.achirus","C.microdon","M.gracilis","N.coatsi","P.gracilis","P.antarcticum","E.antarctica","K.anderssoni","P.bolini","E.carlsbergi","G.fraseri","D.mawsoni","D.eleginoides"),

simple_group=c("bathypelagic fishes","bathypelagic fishes","bathypelagic fishes","bathypelagic fishes","shelf and coastal fishes","mesopelagic fishes","mesopelagic fishes","mesopelagic fishes","mesopelagic fishes","mesopelagic fishes","mesopelagic fishes","mesopelagic fishes","shelf and coastal fishes","mesopelagic fishes","mesopelagic fishes","mesopelagic fishes","mesopelagic fishes","mesopelagic fishes","toothfishes","toothfishes"))

f.prm.w <- inner_join(f.sgroups,f.df)

f.prm.w
```

Tidy up parameter names and add depth specifications that I think are required for the interaction matrix

```{r}
f.prm.s <- f.prm.w %>% group_by(simple_group) %>% summarize(
	 W0 = min(egg_mass_g),
	 Wmax = max(w_max),
	 k_vb=mean(k_vb),
	 beta = mean(beta),
	 max_depth = max(max_depth),
	 min_depth = min(min_depth),
	 Wmat.mizer=mean(w_mat)
	) %>% rename(species=simple_group)

wcu <- data.frame(species=c("bathypelagic fishes","shelf and coastal fishes","mesopelagic fishes","toothfishes"),
                  water.column.use=c("DVM","ontogenetic","DVM","non DVM"))

# wcu <- data.frame(species=c("bathylagiids","gonostomatiids","large myctophids","other bathypelagics","other mesopelagics","shelf and coastal fishes","small myctophids", "toothfish"), water.column.use=c("DVM","non DVM","DVM","non DVM","non DVM","ontogenetic","DVM","non DVM"))

f.prm.s<- inner_join(f.prm.s, wcu)

# f.prm.s$max_depth <- 2000 # set all max depths to 2000m 

f.prm.s
```
## Squid

```{r}
# simplified squids groups
s.prm.w<- readRDS("squid_params_wide.RDS")

s.prm.w

# s.prm.w$simple_group <- c("pelagic squids","pelagic squids", "small nektonic squids")

# s.prm.s <- s.prm.w %>% group_by(simple_group) %>% summarize(
# 	 W0 = min(egg_mass_g),
# 	 Wmax = max(wmax),
# 	 beta = mean(beta),
# 	 max_depth = max(max_depth),
# 	 min_depth = min(min_depth) 
# 	) %>% rename(species=simple_group) 

s.prm.s <- s.prm.w %>% 
	 mutate(w_mat=lw_a*(lmat_cm*10)^lw_b, wmax=lw_a*(lmax_cm*10)^lw_b ) %>%
summarize(
	 W0 = min(egg_mass_g),
	 Wmax = max(wmax),
	 Wmat.mizer = mean(w_mat),
	 beta = mean(beta),
	 max_depth = 2000,
	 min_depth = 0 
	) %>% mutate(species="squids") 

s.prm.s$water.column.use <- "non DVM"

s.prm.s
```



```{r}
# simplified pred groups
p.if <- read_excel("./excel data/predator_parameter_collating_v1.xlsx", sheet="interaction_factors")

# 	  - baleen whales
#     - toothed whales and large seals: large divers
#     - small seals and large penguins: medium divers
#     - small penguins: small divers
# 	  - flying birds
# 	  - apex predators

p.prm.w <- readRDS("predator_params_wide.RDS")
# p.names <- read_csv("./csvs/predator_group_names.csv")
p.names <- read_csv("./csvs/predator_group_names_catch_calibration.csv")


# p.prm.s<- p.prm.w %>% inner_join(p.names) %>% group_by(simple_group) %>% summarize(
# 		Wmax = max(wmax),
# 		W0	 = min(w_indep),
# 		Wmat = min(w_indep),
# 		beta = 1/mean(beta),
# 	)


df_pred <- p.prm.w %>% inner_join(p.names)
```

Check the group means

```{r}
df_group_mean <- df_pred %>% 
  # group_by(simple_group) %>% 
  group_by(group_names_catch) %>% 
  summarise(Wmin_group = min(w_birth),
            Wind_group = min(w_indep),
            Wmat_mean = mean(wmat),
            Wmax_group = max(wmax))

df_group_mean
```




```{r}
p.prm.s<- p.prm.w %>% inner_join(p.names) %>% group_by(group_names_catch) %>% summarize(
		Wmax = max(wmax),
		W0	 = min(w_indep),
		Wmat.mizer = mean(wmat),
		beta = 1/mean(beta),
	)

p.prm.s
```


```{r}
p.prm.s$water.column.use<- "diving"
# note that flying birds are classified as "diving"

# depth ranges
p.prm.s <- p.if %>% inner_join(p.names, by=c("taxon"="group")) %>% group_by(group_names_catch) %>%
	summarize(
		max_depth = max(`max diving depth`),
		p_time_prydz = mean(`p time spent in area (prydz)`)
		) %>% inner_join(p.prm.s) %>% rename(species=group_names_catch)

# per-capita offspring 

# 60 % success for emperor penguins (Robinson 1992)
# assume 60% for small and medium divers
# 50 % for large divers, apex predators, baleen whales and flying birds

# p.prm.s$pc_annual_offspring <- c(rep(0.5,4),rep(0.6,2))
p.prm.s$pc_annual_offspring <- c(rep(0.5,4),rep(0.6,1),rep(0.5,2),rep(0.6,1),rep(0.5,1)) # Edit to match expanded groups

# # intake rates
pQB <- read_excel("./excel data/predator_parameter_collating_v1.xlsx", sheet="QBs")
p.prm.s <- pQB %>% group_by(group_names_catch) %>% summarize(h=mean(QonBmax)) %>% ungroup() %>% rename(species=group_names_catch) %>% full_join(p.prm.s)

# Not using empirical intake rates for now, let the model calculate these based on size and kvb
p.prm.s
```

```{r}
z.prm.s_save <- z.prm.s

z.prm.s_save$Wmat.mizer <- 10^(z.prm.s$Wmat)  
z.prm.s_save$Wmax <- 10^(z.prm.s$Wmax)
z.prm.s_save$W0 <- 10^(z.prm.s$W0)

z.prm.s_save
```


weight in grams
```{r}
z.prm.s_save
p.prm.s
f.prm.s
s.prm.s

all.prm <- full_join(z.prm.s_save, p.prm.s) %>% full_join(f.prm.s) %>% full_join(s.prm.s)

all.prm
```



```{r}

# fill gaps
all.prm[is.na(all.prm$min_depth),]$min_depth<-0
all.prm[is.na(all.prm$p_time_prydz),]$p_time_prydz<-1
all.prm[is.na(all.prm$water.column.use),]$water.column.use<-"non DVM"

all.prm
```


```{r}

all.prm <- select(all.prm, species,W0, Wmat.mizer, Wmax, beta, k_vb, h, min_depth, max_depth, water.column.use, p_time_prydz,pc_annual_offspring)

all.prm

```

```{r}

all.prm.tidy <- all.prm %>% rename(
		w_inf = Wmax,
		w_min	 = W0,
		w_mat = Wmat.mizer
	)

all.prm.tidy
```

```{r}
biomass_observed <- c(5.9, # euphausiids 
                      0.127, # baleen whales
                      0.003, # flying birds
                      0.022, # large divers
                      0.002, # leopard seals
                      0.265,  # medium divers
                      0.014, # minke whales
                      0.006, # orca
                      0.016,  # small divers
                      0.011, # sperm whales
                      1.93, # bathypelagic fishes
                      1.2,   # mesopelagic fishes
                      0.802, # shelf and coastal fishes
                      0.75, # toothfishes
                      0.15 # squids
                      )

# biomass_observed <- c(5.9, # Euphausiids 
#                       0.008, # apex predators
#                       0.141, # baleen whales
#                       0.003, # flying birds
#                       0.022, # large divers
#                       0.265,  # medium divers
#                       0.016,  # small divers
#                       1.93, # bathypelagic fishes
#                       1.2,   # mesopelagic fishes
#                       0.802, # shelf and coastal fishes
#                       0.75, # toothfishes
#                       0.15 # squids
#                       )



all.prm.tidy$biomass_observed <- biomass_observed

all.prm.tidy

all.prm.tidy.ordered <- all.prm.tidy %>%  
  arrange(w_inf)

all.prm.tidy.ordered
```

```{r}
# write.csv(all.prm.tidy.ordered, "trait_groups_params_vCWC.csv") #version Commercial Whaling Calibration
# write_rds(all.prm.tidy.ordered, "trait_groups_params_vCWC.RDS")

# all.prm.tidy.ordered <- readRDS("trait_groups_params.RDS")
```

```{r}
# all.prm %>% saveRDS("all_groups_params_distributions.rds")
# all.prm.tidy.ordered %>% saveRDS("trait_groups_params_distributions_vCWC.rds")
```


```{r}
ggplot(all.prm.tidy.ordered, aes(log10(w_inf), log10(biomass_observed), colour = species)) +
  geom_point()
```


log10 weights
```{r}
# weights into common units before binding

# p.prm.s_save <- p.prm.s
# f.prm.s_save <- f.prm.s
# s.prm.s_save <- s.prm.s
# 
# p.prm.s_save$Wmat.mizer <- log10(p.prm.s$Wmat.mizer)  
# p.prm.s_save$Wmax <- log10(p.prm.s$Wmax)
# p.prm.s_save$W0 <- log10(p.prm.s$W0)
# 
# f.prm.s_save$Wmat.mizer <- log10(f.prm.s$Wmat.mizer)
# f.prm.s_save$Wmax <- log10(f.prm.s$Wmax)
# f.prm.s_save$W0 <- log10(f.prm.s$W0)
# 
# s.prm.s_save$Wmat.mizer <- log10(s.prm.s$Wmat.mizer)
# s.prm.s_save$Wmax <- log10(s.prm.s$Wmax)
# s.prm.s_save$W0 <- log10(s.prm.s$W0)
# 
# p.prm.s_save
# f.prm.s_save
# s.prm.s_save
```









